### 2020-02-20 docs/papers/osr/body.tex
An initial outline.

\newcommand{ironstar}{Iron{\star}}
\newcommand{bepsilon}{B^{\epsilon}}

Systems verification is picking up momentum.
The promise is that we might someday replace a methodology of
write-test-debug with a comprehensive compile-time confirmation
of correctness. In particular, critical software could remove
``ship'' from the bug-fix development cycle.

The chief obstacle to the practical application of verification for
systems-scale software development is the sheer tedium of involved
in reducing a proof to the level of mechanically verifiability.
Automated verifiers based on SAT solvers have made tremendous progress
grinding through this logical tedium, brinigng systems-scale
verification into the realm of feasibility.

\section{{\ironstar}: an automation-based approach}

The authors have been involved in a few projects~\cite{}
over which we have developed strategies that exploit
SAT-solver automation to improve the scalability of verification.

\section{What makes systems verification challenging?}

\subsection{Stating a system specification}

\subsection{Structuring a proof for maintainability}

\subsection{Proof tedium}

\subsection{Heap reasoning}

\subsection{Concurrency}

\subsection{Liveness and performance properties}

\section{How {\ironstar} addresses these challenges}

\subsection{Stating a system specification}\label{sec:specification}

{ironstar} adopts Lamport's TLA+ model for specifying the
allowed behaviors of an evolving system~\cite{specifying-systems}.
TLA+ emphasizes
\begin{itemize}
\item Mathematical abstraction to elide every detail that isn't
  essential to a specification
\item Nondeterminism to capture environmental concurrency,
  nondeterminism, adversarial behavior, and abstract 
\end{itemize}

TLA+ is built on untyped set theory, but we find using a
decidable type system dramatically more convenient,
as it dispatches many tedious proof obligations in the type checker before
reaching the theorem prover.
Dafny's type system offers numeric and logical primitive types,
sequences, sets and maps, user-defined
algebraic (struct/tagged-union) recursive types, and generics.

We define behaviors as TLA+-style state machines:
a struct provides the \textemph{state} type, and
then \textsc{Init} and \textsc{Next} predicates define
the allowed nondeterministic state machine relations.

A state machine is a natural way to define the goalposts for a system.
How should a replicated state machine behave?
Just like a instance of single state machine.
How should a physically decentralized filesystem behave?
Just like a logically-centralized filesystem~\cite{Farsite}.
How should a {\bepsilon}tree store behave?
Just like a simple map.

A state-machine goal avoids a common verification trap of specifying
an implementation's correctness by its invariants. The invariants are
means to an end; a reference state machine is often the most straightforward
way to describe that end.

\subsection{Modeling the environment}

A distributed system consists of several processes running
the program we wish to verify,
a network,
and our assumption that those the processes may only interact through
the network.

A storage system consists of a process running the program
we wish to verify,
a storage device,
our assumption about the asynchronous bus that connects the
device to the process,
and our assumption that the system may sponatenously \texttt{crash},
zeroing the process memory but preserving the device memory.
\footnote{
It may be easier to see how general the model is if you imagine
the storage system as a strangely-shaped distributed system:
one process that runs the program,
another that runs a trusted storage state machine,
a network (the SATA bus) that connects them,
and a failure model for the processes.
The same sorts of nondeterminism that reorders network messages and fails
hosts in a distributed system
can reorder I/O requests and fail the filesystem process.
}

In both cases, the ultimate theorem claims that the
\textit{system} behavior---process(es) and environment---refines
to that of the logically-centralized specification.
What does mechanical verification of the program mean?
It means that, if the program is instantiated as processes in an
environment matching the \textit{environment model},
its behavior will match the \textit{application model}.
That is, one must read the text defining not only the application
but also the environment to know what the verification promises;
in exchange, one needn't read the program implementation, as the
mechanical verifier has done that for you.

\subsection{Structuring a proof for maintainability}

A systems verification problem, ultimately, is simply a really complex
proof goal: an inductive statement of refinement between an implementation
state machine and a specification state machine, each of which is a
big tree of definitions ultimately rooted in some first-order logic or
basic datatypes.

In early interactive theorem provers, the operator stated a goal theorem.
The proof engine offered a bag of tools that transform the proof rules.
The operator repeatedly applied tools until the proof goal has been
transformed into \texttt{true}.
These systems would record the operator's effort as a ``proof script'';
one could replay the script to verify the veracity of the top-level goal.

Proof scripts are brittle; a small change in the system definitions perturb
the proof, and necessary repairs are often non-local~\cite{}.
Chlipala has made remarkable progress in organizing proof scripts into
tactic modules that mimic the structure of the system under
verification. His approach improves the modularity and locality
of maintenance of the proof artifact.
It requires the operator to simultaneously maintain the system
and the tactic library that supports it.

\comment{Diagram here.}

An {\ironstar} proof is a refinement hierarchy of state machines.

In {\ironstar}, the operator works as follows:
First, she defines the state machine (and its data structures) that
specify the application~(\S\ref{sec:specification}).
She uses abstraction to omit as many details as possible while capturing the
desired application behavior.
She also models the finished \texttt{system} as a compound state machine that
models the environment and takes a program as a parameter.
Her goal is to write a program that, when instatiated in the \texttt{system},
causes the system to satisify the application spec.

The program will ultimately be expressed as
imperative code over an immutable heap,
so that the programmer has sufficient control to achieve the
desired performance.
The imperative code is event-driven, so that we can view each
trip through the main loop or invocation of a handler as an
atomic step in a state machine,
and thus can be plugged in as a component into the \texttt{system} definition.

In any system of reasonable size, the semantic gap between the abstract
application definition and the pointer-diddling implementation will be
substantial.
Why is any given line of C++ code correct? There are surely many reasons.
Perhaps the \texttt{->} operator isn't a null pointer dereference because
the pointer was freshly allocated six lines ago; an automatic tool may
be able to confirm such local or simple properties.
But the arguments to the referenced method might be the right ones because
they were the results of another method call earlier in the function.
And the side-effect of the method call might be correct
because it establishes a property that will be used when another I/O
event completes in the future.
Most of these higher-level arguments appear at best in comments,
or perhaps in a distant and outdated English design document,
or they can be inferred from the penumbras and emanations of the
call graph or object hierarchy.
Most likely, they appear nowhere other than an author's mental model.

In a verified system, every such correctness argument must be explicitly
stated and the collection chained together to connect the implementation
code all the way up to the abstract application specification.
In an {\ironstar} verified system, these arguments are organized into
a refinement hierarchy.
In the refinement hierarchy, state machine models act as waypoints
along the route from abstract to concrete. Each refinement proof that
connects a state machine to the abstract one above it captures some
aspect of the system's correctness reasoning. 
The refinement hierarchy is a mechanism to \textit{organize the
proof arguments} into a maintainable structure.

The idea of refinement as a proof strategy is far from new with {\ironstar}.
Note that we do \textit{not} aim to use refinement
as a way to break a proof down into small-enough pieces that a
mechanical verifier can leap the layers completely automatically
(as is done in~\cite{cspec,ironarmada}).
Instead, the refinement layers are meant to be ``human-sized'' concepts:
just as a skilled engineer decomposes a function or organizes a
class hierchy at a scale that's designed to communicate to the next
human reader,
refinement layers in {\ironstar} are designed to organize the proof
artifact into human-brain-sized conceptual chunks.

\comment{Discuss the tree hierarchy in veribetrfs}
Figure~{diagram} shows the richest refinement hierarchy in the
{\veribetrfs} system.
The top refinement explains how queries searching a {\bepsilon}tree
structure give it behavior equivalent to a map,
while abstracting the internal structure of tree nodes as infinite
maps.
The next layer explains how nodes can be implemented in finite space
with pivots; note that this layer disregards the {\bepsilon}tree structure,
since those correctness arguments are already dispatched in the
layer above.
The pivot layer assumes that nodes are abstract datatypes that can
be passed through disk storage, which of course is an abstraction of
the real disk. The next layer shows how marshalling can provide the
same abstract data types while passing byte strings through disk storage.

Each of the layers above are written in with immutable data types
(that is, in a functional, side-effect-free style), abstracting away
the complexity of in-place updates. They also make widespread use of
mathematical integers, wherein $i+1$ is always bigger than $i$.
The bottom layer -- the concrete implementation -- uses machine integers
and exploits in-place updates to achieve excellent performance.
Its refinement proof shows that both transformations are correct,
but again, it disregards any of the reasoning at higher levels of
abstraction.

The developer uses

\subsection{Proof tedium}

\subsection{Heap reasoning}

\subsection{Concurrency}

\subsection{Liveness and performance properties}

\section{Managing automation}

### 2020-02-20 framework/Benchmarks.cpp
avoid calling ComputeSize for indirection table - small optimization

@@ -198,6 +198,8 @@ public:
       }
 
       app.Insert(key, value);
+
+      //if (i % 50000 == 0) app.Sync();
     }
 
     app.Sync();

### 2020-02-20 framework/Benchmarks.cpp
marshall faster without constructing a V

@@ -199,7 +199,7 @@ public:
 
       app.Insert(key, value);
 
-      //if (i % 50000 == 0) app.Sync();
+      if (i % 50000 == 0) app.Sync();
     }
 
     app.Sync();

### 2020-02-20 tools/run-veri-config-experiment.py
have run script print command it uses

@@ -155,7 +155,9 @@ def main():
 
   clear_page_cache()
 
-  os.system("cgexec -g memory:VeribetrfsExp ./build/VeribetrfsYcsb " + wl + " " + loc + " --veribetrkv")
+  command = "cgexec -g memory:VeribetrfsExp ./build/VeribetrfsYcsb " + wl + " " + loc + " --veribetrkv"
+  print(command)
+  os.system(command)
   assert ret == 0
 
 if __name__ == "__main__":

### 2020-02-20 docs/papers/osr/body.tex
more text

@@
 \newcommand{bepsilon}{B^{\epsilon}}
+\newcommand{veribetrfs}{\texttt{\textit{veri}betrfs}}
 
@@
 \comment{Discuss the tree hierarchy in veribetrfs}
-Figure~{diagram} shows the richest refinement hierarchy in the
+Figure~\ref{betreerefinement} shows the richest refinement hierarchy in the
 {\veribetrfs} system.
@@
 
-The developer uses
+The developer creates the intermediate state machines as
+``interface boundaries'' that separate concerns.
+She organizes the bulk of the proof artifact
+by slicing it across the axis of abstraction.
+
+Of course, the proof artifact can also be orthogonally organized
+by module.
+In {\veribetrfs},
+the {\bepsilon}tree refinement stack of Figure~\ref{betreerefinement}
+has a (simpler) sibling that reasons about a disk block caching module.
 
 \subsection{Proof tedium}
+
+\subsubsection{A day in the life of a verification engineer.}
+What is it like to develop software in this style?
+
+Define a state machine.
+- it's immutable datatypes; get comfortable with abstraction.
+State a refinement.
+- it's a boilerplate induction pattern.
+Assert its proof.
+- Start with {} -- see how far automation takes you.
+- Probably a simple case breakdown. Write an if or case statement!
+Diagnose
+- bisection debugging! This is an exciting part of the story;
+  forward-reference it. Contrast it with proof-goal manipulation
+  (Coq/Isabelle/TLAPS) or counterexample deconstruction.
+Develop an invariant
+- I bet you're missing an invariant
+Patch up a proof
+- bugs in definitions (off by one errors)
+- witnesses
+- automation syntactic triggering
+
+recursion vs. quantification story
 

### 2020-02-21 framework/Framework.h
don't copy when taking subseq and fix a few things that get slow as a result

@@ -46,6 +46,31 @@ namespace NativeArrays_Compile {
       }
     }
 
+    static int32_t ByteSeqSliceCmpByteSeqSlice(
+        DafnySequence<uint8> b1,
+        uint64 lo1, uint64 hi1,
+        DafnySequence<uint8> b2,
+        uint64 lo2, uint64 hi2)
+    {
+      size_t size1 = hi1 - lo1;
+      size_t size2 = hi2 - lo2;
+      int result = memcmp(
+          b1.ptr() + lo1,
+          b2.ptr() + lo2,
+          size1 < size2 ? size1 : size2);
+      if (result == 0) {
+        if (size1 == size2) {
+          return 0;
+        } else if (size1 > size2) {
+          return 1;
+        } else {
+          return -1;
+        }
+      } else {
+        return result;
+      }
+    }
+
     template <typename T>
     static DafnyArray<T> newArrayFill(uint64 len, T val)
     {

### 2020-02-21 tools/checkout-dafny-commit.sh

@@ -20,4 +20,4 @@ set -x
 
 # https://github.com/secure-foundations/dafny.git
 # cpp branch
-git checkout aac89fb539a0173a7862da35428a6394d75b900e
+git checkout 735dfcac773ff5df1635cb6f75a45fa0d1227723

### 2020-02-21 framework/Benchmarks.cpp
benchmark syncing

@@ -620,6 +620,10 @@ void benchmark_append(string const& name, long long ns) {
   sw[name].append(ns);
 }
 
+void benchmark_clear() {
+  sw.clear();
+}
+
 void dump() {
   for (auto& p : sw) {
     string name = p.first;

### 2020-02-21 framework/Framework.cpp

@@ -418,6 +418,8 @@ void Application::crash() {
 }
 
 void Application::Sync() {
+  benchmark_start("Application::Sync");
+
   LOG("Sync");
 
   uint64 id = handle_PushSync(k, hs, io);
@@ -428,16 +430,21 @@ void Application::Sync() {
 
   for (int i = 0; i < 500000; i++) {
     while (this->maybeDoResponse()) { }
+
     auto tup2 = handle_PopSync(k, hs, io, id);
     bool wait = tup2.first;
     bool success = tup2.second;
     if (success) {
       LOG("doing sync... success!");
       LOG("");
+
+      benchmark_end("Application::Sync");
       return;
     } else if (wait) {
       LOG("doing wait...");
+      benchmark_start("waitForOne");
       io->waitForOne();
+      benchmark_end("waitForOne");
     } else {
       LOG("doing sync...");
     }
@@ -562,12 +569,16 @@ void Application::QueryAndExpect(ByteString key, ByteString expected_val)
 bool Application::maybeDoResponse()
 {
   if (io->prepareReadResponse()) {
+    benchmark_start("handle_ReadResponse");
     handle_ReadResponse(k, hs, io);
+    benchmark_end("handle_ReadResponse");
     LOG("doing read response...");
     return true;
   }
   else if (io->prepareWriteResponse()) {
+    benchmark_start("handle_WriteResponse");
     handle_WriteResponse(k, hs, io);
+    benchmark_end("handle_WriteResponse");
     LOG("doing write response...");
     return true;
   }

### 2020-02-21 ycsb/YcsbMain.cpp

@@ -214,6 +214,9 @@ void ycsbRun(
     }
 }
 
+void benchmark_clear();
+void dump();
+
 #ifdef _YCSB_VERIBETRFS
 class VeribetrkvFacade {
 protected:
@@ -237,7 +240,11 @@ public:
     }
 
     inline void sync() {
+        benchmark_clear();
         app.Sync();
+        cout << "========================================" << endl;
+        dump();
+        cout << "========================================" << endl;
     }
 };
 

### 2020-02-21 tools/run-veri-config-experiment.py
support rocks in the experiment script

@@ -89,6 +89,8 @@ def main():
   value_updates = []
   config = None
 
+  rocks = None
+
   for arg in sys.argv[1:]:
     if arg.startswith("ram="):
       ram = arg[len("ram=") : ]
@@ -106,6 +108,8 @@ def main():
       config = "64kb"
     elif arg == "config-8mb":
       config = "8mb"
+    elif arg == "rocks"
+      rocks = True
     else:
       assert False, "unrecognized argument: " + arg
 
@@ -113,6 +117,7 @@ def main():
   assert device != None
 
   if config != None:
+    assert not rocks
     assert ram != None
     value_updates = autoconfig(config, ram) + value_updates
 
@@ -128,11 +133,19 @@ def main():
   assert ret == 0
 
   for (name, value) in value_updates:
+    assert not rocks
     print("setting " + name + " to " + value)
     splice_value_into_bundle(name, value)
 
+  if rocks:
+    exe = "build/RocksYcsb"
+    cmdoption = "--rocks"
+  else:
+    exe = "build/VeribetrfsYcsb"
+    cmdoption = "--veribetrkv"
+
   print("Building executable...")
-  ret = os.system("make build/VeribetrfsYcsb -s -j4 > /dev/null 2> /dev/null")
+  ret = os.system("make " + exe + " -s -j4 > /dev/null 2> /dev/null")
   assert ret == 0
 
   wl = "ycsb/workload" + workload + "-onefield.spec"
@@ -155,8 +168,13 @@ def main():
 
   clear_page_cache()
 
-  command = "cgexec -g memory:VeribetrfsExp ./build/VeribetrfsYcsb " + wl + " " + loc + " --veribetrkv"
+  # bitmask indicating which CPUs we can use
+  # See https://linux.die.net/man/1/taskset
+  taskset_cmd = "taskset 4 "
+
+  command = taskset_cmd + "cgexec -g memory:VeribetrfsExp ./" + exe + " " + wl + " " + loc + " " + cmdoption
   print(command)
+
   os.system(command)
   assert ret == 0
 

### 2020-02-21 ycsb/YcsbMain.cpp
ycsb always sync at the end

@@ -194,6 +194,13 @@ void ycsbRun(
         }
     }
 
+    auto sync_started = chrono::steady_clock::now();
+    db.sync();
+    auto sync_completed = chrono::steady_clock::now();
+    auto sync_duration = std::chrono::duration_cast<std::chrono::nanoseconds>(
+        sync_completed - sync_started).count();
+    sync_latency_hist.add_value(sync_duration);
+
     auto clock_end = chrono::steady_clock::now();
     long long bench_ns = std::chrono::duration_cast<std::chrono::nanoseconds>(clock_end - clock_start).count();
 

### 2020-02-21 tools/run-veri-config-experiment.py
fix bugs in experiment script

@@ -108,7 +108,7 @@ def main():
       config = "64kb"
     elif arg == "config-8mb":
       config = "8mb"
-    elif arg == "rocks"
+    elif arg == "rocks":
       rocks = True
     else:
       assert False, "unrecognized argument: " + arg
@@ -128,9 +128,10 @@ def main():
   ret = os.system("rm -rf build/")
   assert ret == 0
 
-  print("Building Bundle.cpp...")
-  ret = os.system("make build/Bundle.cpp > /dev/null 2> /dev/null")
-  assert ret == 0
+  if not rocks:
+    print("Building Bundle.cpp...")
+    ret = os.system("make build/Bundle.cpp > /dev/null 2> /dev/null")
+    assert ret == 0
 
   for (name, value) in value_updates:
     assert not rocks

### 2020-02-21 framework/Benchmarks.cpp
add log_stats option to run-veri-config-experiment.py

@@ -620,13 +620,17 @@ void benchmark_append(string const& name, long long ns) {
   sw[name].append(ns);
 }
 
-void dump() {
+void benchmark_dump() {
   for (auto& p : sw) {
     string name = p.first;
     int count = p.second.count;
     long long ms = p.second.ns / 1000000;
     cout << name << " " << ms << " ms, " << count << " ticks" << endl;
   }
+}
+
+void benchmark_clear() {
+  sw.clear();
 }
 
 namespace NativeBenchmarking_Compile {
@@ -641,5 +645,5 @@ namespace NativeBenchmarking_Compile {
 
 void RunBenchmark(string filename, string const& name, bool skipPrepare) {
   benchmark_by_name(name)->run(filename, skipPrepare);
-  dump();
+  benchmark_dump();
 }

### 2020-02-21 framework/Framework.h

@@ -208,5 +208,7 @@ namespace NativeBenchmarking_Compile {
 void benchmark_start(std::string const&);
 void benchmark_end(std::string const&);
 void benchmark_append(std::string const&, long long ns);
+void benchmark_clear();
+void benchmark_dump();
 
 void Mkfs(std::string filename = ".veribetrfs.img");

### 2020-02-21 tools/run-veri-config-experiment.py

@@ -88,6 +88,7 @@ def main():
   mem = None
   value_updates = []
   config = None
+  log_stats = False
 
   rocks = None
 
@@ -110,6 +111,8 @@ def main():
       config = "8mb"
     elif arg == "rocks":
       rocks = True
+    elif arg == "log_stats":
+      log_stats = True
     else:
       assert False, "unrecognized argument: " + arg
 
@@ -145,8 +148,13 @@ def main():
     exe = "build/VeribetrfsYcsb"
     cmdoption = "--veribetrkv"
 
+  make_options = ""
+  if log_stats:
+    make_options = "LOG_QUERY_STATS=1 "
+
   print("Building executable...")
-  ret = os.system("make " + exe + " -s -j4 > /dev/null 2> /dev/null")
+  ret = os.system(make_options + "make " +
+      exe + " -s -j4 > /dev/null 2> /dev/null")
   assert ret == 0
 
   wl = "ycsb/workload" + workload + "-onefield.spec"

### 2020-02-21 ycsb/YcsbMain.cpp

@@ -183,6 +183,15 @@ void ycsbRun(
 
             cerr << db.name << " [op] sync (completed " << i << " ops)" << endl;
 
+            #ifdef _YCSB_VERIBETRFS
+            #ifdef LOG_QUERY_STATS
+            cout << "=========================================" << endl;
+            benchmark_dump();
+            benchmark_clear();
+            cout << "=========================================" << endl;
+            #endif
+            #endif
+
             auto sync_duration = std::chrono::duration_cast<std::chrono::nanoseconds>(
                 sync_completed - clock_op_completed).count();
             sync_latency_hist.add_value(sync_duration);

### 2020-02-25 framework/BundleWrapper.cpp
journals

@@ -7,8 +7,8 @@ std::pair<Constants, Variables> handle_InitState()
 {
   auto tup2 = __default::InitState();
   Constants k;
-  k.k = std::shared_ptr<BetreeGraphBlockCache_Compile::Constants>(
-      new BetreeGraphBlockCache_Compile::Constants(tup2.t0));
+  k.k = std::shared_ptr<BlockJournalCache_Compile::Constants>(
+      new BlockJournalCache_Compile::Constants(tup2.t0));
   Variables hs;
   hs.hs = tup2.t1;
   return std::make_pair(k, hs);
@@ -16,7 +16,7 @@ std::pair<Constants, Variables> handle_InitState()
 
 DafnyMap<uint64, DafnySequence<uint8>> handle_Mkfs()
 {
-  return MkfsImpl_Compile::__default::Mkfs();
+  return __default::Mkfs();
 }
 
 uint64 handle_PushSync(Constants k, Variables hs, std::shared_ptr<MainDiskIOHandler_Compile::DiskIOHandler> io)
@@ -26,7 +26,7 @@ uint64 handle_PushSync(Constants k, Variables hs, std::shared_ptr<MainDiskIOHand
 
 std::pair<bool, bool> handle_PopSync(Constants k, Variables hs, std::shared_ptr<MainDiskIOHandler_Compile::DiskIOHandler> io, uint64 id)
 {
-  auto p = __default::handlePopSync(*k.k, hs.hs, io, id);
+  auto p = __default::handlePopSync(*k.k, hs.hs, io, id, true /* graphSync */);
   return std::make_pair(p.t0, p.t1);
 }
 

### 2020-02-25 framework/BundleWrapper.h

@@ -3,7 +3,7 @@
 #include <optional>
 #include "Framework.h"
 
-namespace BetreeGraphBlockCache_Compile {
+namespace BlockJournalCache_Compile {
   struct Constants;
 }
 namespace MainHandlers_Compile {
@@ -11,7 +11,7 @@ namespace MainHandlers_Compile {
 }
 
 struct Constants {
-  std::shared_ptr<BetreeGraphBlockCache_Compile::Constants> k;
+  std::shared_ptr<BlockJournalCache_Compile::Constants> k;
 };
 
 struct Variables {

### 2020-02-27 docs/papers/osr/body.tex
a draft

@@
 
+For half a century, we have known that a program is constructed by
+writing code, testing, and debugging. Program verification promises
+an alternative process in which no bug escapes the compiler.
+
+Convincing a mechanical checker that a program is correct requires
+resolving a fantastic degree of detail.
+The SMART system offered an extension to Paxos~\cite{smart}.
+Its authors provided a hand-written proof at a level of detail intended
+to satisfy a mechanical checker, although it was never checked.
+Just the proof of the SMART algorithm, without an imperative implementation,
+is 70 pages of text~\cite{smart-proof}.
+If we are to apply verification as a practical engineering tool,
+we must rely heavily on automation to retire tedious logic obligations,
+focusing humans on just the subtle invariants that reflect the insights
+of the system design.
+
+The authors have verified several systems using highly-automated
+verification tools. This paper conveys the experience of learning and using
+such tools. We argue that strong automation is the path forward for
+large-scale systems verification, and we highlight open challenges.
+
+\section{A verified codebase}
+
+The code base of a verified system has three parts:
+the specification, the implementation, and the proof.
+
+The \textit{specification} defines the application spec
+(what does the program do?) and the environment spec
+(what can the program assume about its environment?).
+
+The \textit{implementation} is the source code that will compile
+into the final program. It is imperative (has in-place updates) and
+uses machine-supported datatypes (uint64) to give the programmer
+enough control to achieve great performance.
+
+The \textit{proof} files illuminate the invariants and structure
+the arguments that show that the implementation does what the specification
+says.
+
+A typical specification is smaller than 1/10^{th} the size of
+the implementation it constrains.
+In {\ironstar}, the spec files carry a distinguished file extension,
+and spec files may only include other spec files.
+If the developer reads all of the spec files, she need read no other
+files to understand what the verified program does.
+
+An {\ironstar} proof text is around five times
+as big as the implementation code.
+The effort spent constructing and maintaining it is in lieu of
+maintaining a conventional test suite.
+
+\section{A day in the life of a verification engineer}
+
+At the large scale, the team's task is to write a fast implementation,
+and show that its behavior matches the spec. This task breaks down into
+making smaller arguments; a typical argument might say ``splitting a node
+in a tree creates a new tree that represents all the same key-value pairs.''
+
+When using an automation-centric tools, the engineer begins each such
+argument the same way: She states the argument, and then offers an empty
+proof:
+\begin{verbatim} % TODO get an environment that numbers lines.
+lemma TODOPlausibleLemmaName(before:Tree, after:Tree, pivot:Key)
+  requires TreeInv(a)
+  requires SplitNodeRelation(before, after, pivot)
+  ensures TreeInv(b)
+{
+}
+\end{verbatim}
+Then she asks the verifier to try to verify the lemma.
+This is analogous to testing every possible argument assignment,
+but of course the mechanism is not actually model checking, since
+there are infinite such assignments.
+
+Sometimes, automation
+successfully opens the tall tree of definitions under
+\texttt{SplitNodeRelation} and \texttt{TreeInv}, and confirms the validity
+of the lemma.
+More commonly, the verifier cannot convince itself, and it reports
+\texttt{TODO real text at line 4}.
+Maybe the lemma is invalid: The definitions don't actually
+do what the engineer hoped, either because of a deep misunderstanding
+or due to a minor off-by-one error.
+This is why we're using automation in the first place; it's a failed test
+pointing to a flaw in the designer's understanding.
+Or perhaps the lemma is valid, but the automation can't see why.
+This is by far the more common case; it's the tax we pay to enjoy
+exhaustive testing.
+
+However, the engineer's action in either case is the same: she must
+elaborate the proof until either the verifier admits its validity
+or she discovers an underlying error.
+
+The currency of definitions is are quantified logical statments.
+\verb{TreeInv(t)} is a list of conjuncts; one says that the tree
+is acyclic:
+\begin{verbatim}
+predicate Acyclic(t) {
+  forall p, i, j :: Path(p, t) && i!=j ==> p[i] != p[i]
+}
+\end{verbatim}
+This says that no path through the tree visits the same node twice.
+
+Suppose the engineer wonders if acyclicity is the thing tripping up the
+verifier. So she replaces the lemma proof with:
+\begin{verbatim}
+{
+  assert Acyclic(b);
+}
+\end{verbatim}
+The verifier will respond in one of the following ways:
+\begin{itemize}
+  \item line 6 assertion failed: Yep, acyclicity is the difficult part,
+  \textit{and if we assume it's true, the postcondition verifies}.
+  \item line 4 postcondition failed: acyclicity was valid; some other
+  conjunct in TreeInv(b) is the problem.
+  \item \textit{both} errors: The verified needs help seeing more than one
+  conjunct in \texttt{TreeInv}.
+  \item neither error: the assertion itself has given the verifier a
+  sufficient hint to verify the lemma (see \S\ref{observes}).
+  \item \textit{Timeout}. See \S\ref{timeouts}.
+\end{itemize}
+The first two cases are by far the most common: the engineer has successfully
+"bisected" the problem, either implicating or vindicating the \verb{Acyclic}
+conjunct.
+
+She applies this bisection technique recursively, just as she would use
+printf bisection debugging to back a deterministic runtime test failure into a
+corner.
+If \verb{Acyclic(b)} is unproven, she drills down:
+\begin{verbatim}
+{
+  forall p, i, j | Path(p, t) && i!=j
+    ensures p[i] != p[i]
+  {
+  }
+  assert Acyclic(b);
+}
+\end{verbatim}
+The forall brings new variables \verb{p, i, j} into scope for use in its
+proof body which, again, begins as \verb{\{\}}.
+The verifier complains about the postcondition on line 7.
+It accepts the assert on line 10, predicated on the assumption that the
+engineer will ultimately correct the failing forall statement.
+
+Now that \verb{p, i, j} have been introduced, the engineer may drill
+further down into definitions to argue why \verb{p[i] != p[i]}.
+In this exmaple, she would begin a proof by contradiction:
+If there exists a cyclic path \verb{pb} in \verb{b},
+then construct a witness path \verb{pa} in tree \verb{a} that is
+cyclic, which contradicts the given \verb{Acyclic(a)}.
+
+This iterative, user-guided debugging approach stands in contrast to other
+common verification development methodologies.
+The engineer isn't confronted with a half-transformed proof goal
+with a hundred subterms, forcing her to
+think in the verifier's internal representation.
+Nor does she dig through a counterexample with a value assigned for each
+of hundreds of in-scope variables, only two of which are actually relevant.
+Both of those approaches present information to the user that scales up with
+the system size, and both ask the user to interpret the system's model.
+By contrast, bisection proof debugging lets the engineer develop her own
+mental model by probing the verification Oracle in a game of twenty-questions.
+The benefit of letting the engineer steer the story and maintain her own
+mental model outweighs the fact that she must make a logarithmic number of
+inquiries to reach the root cause, especially because each inquiry is answered
+rapidly.
+
+Returning to the possible verifier responses: in case 3, the verifier
+complains about both the assertion and the postcondition; the engineer
+has failed to bisect the problem. It could be as simple as the assertion
+being incorrect, or it could be that there is more than one problem
+(error in definition, or missing proof) to identify. The engineer proceeds
+the same as above, but has more ground to cover before she's done.
+
+Case 4 is a victory: by introducing the assertion, the engineer has
+triggered the automation to exploit a quantifier it hadn't considered
+automatically, and it has completed the proof.
+
+Case 5, a timeout, is a meta-problem: not only can the verifier not
+see the proof, but the automation strategy has broken down, and requires
+human intervention to fix. These issues are rare: we bisect-debug all day,
+but have to resolve a timeout perhaps once a week.
+We discuss automation control and timeout mitigation in Section~\ref{timeout}.
+
+Other than timeout management, note that the nitty-gritty work of verification
+engineering feels a lot like printf-debugging a deterministic test failure.
+The process is familiar, easy to learn, and it generally requires deep insight
+only about the system being debugged, not about the verification tool itself.
+
+TODO the glory of quantified vs recursive definitions.
+
+\subsection{Structuring the proof}
+
+While the low-level mechanics are simple, the verification engineer does
+need to adopt new disciplines to organize her proof text for scale and
+maintainability. The goal is the same as organizing a big software system
+into modules for maintainability, but the techniques for arranging a
+proof argument are different.
+
+A single line of an imperative program may be correct for many reasons,
+at multiple levels of abstraction, all of which must hold to support the
+program's correctness. For example, in this line:
+\begin{verbatim}
+  node->writeToBlock(lba)
+\end{verbatim}
+Dereferencing \verb{node} is safe because \verb{node} was allocated earlier
+in the method and not freed. The write to address \verb{lba} is safe
+because \verb{lba} was returned from a library function managing free space
+on the device. The content written, a marshalled representation of \verb{node},
+is appropriate because later in the program we'll store a pointer to \verb{lba}
+that expects a representation of \verb{node} to appear there.
+
+Common static analysis tools can confirm non-null properties like the first
+one. The other arguments are at higher levels of abstraction, interact with
+distant lines of code in the program, and may even interect with other
+instances of the process, such as before or after a crash event that resets
+RAM. In a conventional program, each informal argument might appear in a
+comment on the code line, or might be inferrable from the name of
+the methods or the methods it calls, or may appear in an Engilsh design
+document from two years ago, or may appear nowhere but in an engineer's
+memory.
+
+Mechanical verification requires that each such argument be written down.
+Trying to cram all of these arguments into the ground-level source code
+is overwhelming; the more abstract arguments are weighed down with the
+tedium of implementation. In {\ironstar}, we employ state machine refinement
+to organize proof text%
+\footnote{We contrast this use with other verification efforts that also
+use refinement, but in service to the verification system, to divide the
+proof up into layers so fine that automation can leap from one to the next
+without assistance. In {\ironstar}, refinement layers are introduced at
+a granularity that humans find natural, just as developers choose appropriate
+granularity boundaries for functional abstraction or object encapsulation.
+}
+by separating it along the axis of abstraction.
+
+% TODO discuss app specification in more detail somewhere?
+The app spec is a TLA+-style state machine.
+In {\veribetrfs}, the spec is a key-value dictionary extended with
+a \verb{flush} operation and a \verb{crash} event that interleaves
+at the whim of the environment. These specs are all written with immutable
+data structures and mathematical conveniences (unbounded integers, infinite
+sets and sequences) that let us focus on the behavior while ignoring
+implementation details.
+
+The first layer of refinement introduces the structure of a {\bepsilon}
+tree, while abstracting the nodes as infinite sets mapping every key to
+a child node.
+Proof text at this layer argues that the tree state machine maintains the
+necessary invariants such that the tree acts exactly like a dictionary.
+
+The second layer refines the infinite nodes to ``pivot'' nodes, which
+map keys from A---G to one child, H---N to the next, and so on. Proof
+text at this layer argues that pivot nodes are a (finite!) shorthand
+for infinite maps, and hence that {\bepsilon}{\langle}pivot{\rangle} is
+still a dictionary.
+
+The previous layer supposed a disk that accepts the Node abstract data
+type as arguments to \verb{write}; the third layer refines the Pivot Node to
+have a marshalled representation. Proof text at this layer argues
+that \verb{unmarshall(marshall(n) == n}, and thus that, when
+a {\bepsilon} tree made of marshallable nodes is connected to a disk
+that does I/O on byte strings, the system behaves the same as the prior layer.
+
+Finally, the bottom-most layer replaces immutable datatypes with mutable
+ones (to exploit the performance of in-place updates), and replaces
+mathematical integers with machine \verb{uint64}s. Proof text at this
+layer shows that aliasing is controlled enough that mutable data structures
+mimic their immutable models above. It also shows that bounds are respected
+so that \verb{uint64}s don't overflow.
+
+Each of the arguments is necessary to stitch the proof together from the bottom
+layer implementation to the top level application spec.
+The important observation from this tour is that the refinement layers
+act as narrow interfaces along the axis of abstraction: arguments about
+high-level concepts like {\bepsilon} tree structure may omit details
+about marshalling and register bounds; conversely, low-level arguments
+may attend only to detailed reasoning, as the abstract properties were
+dispensed with elsewhere.
+
+Module design in an imperative program creates conceptual boundaries,
+so that an engineer can study and modify part of the program while being
+insulated from interactions with most of the other lines of code in the
+system.
+Likewise, we have found this organization enables us to study and modify
+parts of our verified system while insulating our changes from other
+parts. Most importantly, it means that we can introduce a new feature
+(like sub-node reads) while leaving most of the proof text undisturbed.
+Because proof is 80\% of the text of our verified system,
+incremental and modular maintenance is as important for the proof as
+for the imperaive code.
+
+------------------------------------------------------------------------------
 Systems verification is picking up momentum.

### 2020-03-16 docs/papers/osdi2020/intro.tex

\begin{abstract}
For verification to make the leap from the laboratory to
practical use,
it must become performant, maintainable, and accessible.

\end{abstract}

\section)Introduction}
Systems researchers have begun to demonstrate that substantial
systems software components can be constructed using verification
as the primary means of establishing functional correctness.
Compared to ordinary test-driven software development,
verfication promises effectily exhaustive testing.

What stands in the way of verification replacing testing?
Verification must produce software that offers competitive \textit{performance}.
The languages and techniques must be \textit{accessible}
to mainstream developers.
The burden of creating verified software and its associated
proof-guiding annotations must be \textit{affordable}.
The proof artifact must be incrementally \textit{maintainable} to evolve
as the specification and implementation code evolve.

We present {\veribetrkv}, a high-performance key-value store
with the explicit goal of developing a maintainable code artifact
with a verification methodology.
{\veribetrkv} implements the persistent key-value storage substrate of
BetrFS\cite{}.
We build {\veribetrkv} as the first step in building a verified BetrFS,
with the explicit goal of making an adoptoble filesystem:
one \texit{faster} than ext4, free of corruption and crash-fault bugs,
and maintainable.
On realistic workloads it outperforms RocksDB over disk-like devices
and is competitive within 30\% even over emerging nonvolatile storage.

The methodology, called {\ironstar},
builds on and improves the accessibility and automation of
the tools and techniques used in prior projects.

The executable code in {\ironstar} is specified as imperative, heap-mutating
code, giving the developer a high degree of control over memory behavior
and hence system perforamnce.
{\ironstar} prescribes a refinement-based plan for organizing proof code
into localized modules that enhance maintainability.

To be affordable, a verification methodology must heavily exploit automation.
{\ironstar} improves on its methodological predecessor by offering a path
exploiting maximal automation while avoiding the timeouts that have plagued
prior highly-automated approaches.

The evaluation demonstrates that {\veribetrkv} performs comparably to
state-of-the-art unverified key-value storage, and dramatically outperforms
prior verified systems when considering realistic workloads that stress
the filesystems' CPU activity.

We show that changes to {\veribetrkv} implementation create isolated
changes to its proof obligations, showing that {\ironstar} code is
refactorable and maintainable.

{\veribetrkv} demonstrates these attributes while maintaining an affordable
burden as compared with prior approaches, as measured by the proof:code
ratio.

\endsection{Introduction}

### 2020-09-01 tools/mutablemap-cook.sh
assembling the pdf...

@@ -1,0 +1,32 @@
+#!/bin/bash
+
+logs="$@"
+
+function generate_output_line {
+    operation="$1"
+    if [ -z "$logs" ]; then
+        return
+    fi
+    throughputs=`sed -n "s,^$operation	\(.*\)$,\1,p" $logs`
+    if [ -z "$throughputs" ]; then
+        return
+    fi
+    throughputs=$(for t in $throughputs; do python3 -c "print(64000000 / $t * 1000)"; done)
+    sorted_throughputs=`echo "$throughputs" | sort -n`
+    min=`echo "$sorted_throughputs" | head -1`
+    max=`echo "$sorted_throughputs" | tail -1`
+    nthroughputs=`echo "$sorted_throughputs" | wc -l`
+    median_index=$[ $nthroughputs / 2 + 1 ]
+    median=`echo "$sorted_throughputs" | head -$median_index | tail -1`
+    echo $operation  $median $min $max
+}
+
+function generate_output_file {
+    echo "phase  med min max"
+    operations=("insert" "readpositive" "readnegative" "remove")
+    for operation in ${operations[@]}; do
+        generate_output_line $operation 
+    done
+}
+
+generate_output_file

### 2020-09-01 osdi20-artifact/eval.tex
hey, add the paper, too

\section{Evaluation}\label{sec:eval}

Our evaluation is designed to answer two main questions:
\begin{enumerate}
  \item Do our automation-control techniques (\S\ref{sec:discipline}) and language
    improvements (\S\ref{sec:lang}) improve the developer experience? 
  \item Can our verification methodology scale to the complexity of a
    modern key-value-store data structure, and can we deliver the
    performance gains of write optimization?
\end{enumerate}

\todo{We should measure how much (and what percentage) of code we added/modified to create linear Dafny}

\subsection{Developer Experience}\label{sec:eval:effort}

\paragraph{Measuring Tedium}\label{eval:tedium}
We estimate the amount of tedium (or conversely, the efficacy of automation)
by the ratio of the lines of proof 
(e.g., lemmas, pre-/post-conditions, loop invariants)
to the lines of executable implementation code.
This is not a precise model, since it measures a completely verified artifact,
where the developers may have cleaned up temporary lines of tedium that
were typed in the course of resolving verification failures.
However, the proof text in the ``cleaned up'' code at least
reflects the tedium needed to bridge what automation
could not manage by itself.

\begin{figure}
\begin{center}
\input{data/line-counts}
\end{center}
  \vspace{-3mm}
\caption{Line counts~\cite{sloc} by major components in Figure~\ref{fig:full}.}
\label{line-counts}
\end{figure}

\begin{figure}
\begin{center}
\input{data/line-counts-micro}
\end{center}
  \vspace{-3mm}
\caption{Line counts~\cite{sloc} of two subcomponents, comparing dynamic-frame implementations with  our linear type system.\protect\footnotemark~
        Linear typing reduces the proof burden by 31-37\%}
\label{line-counts-micro}
  \vspace{-2mm}
\end{figure}

%\todo{It would be nice to have Disk, IoSystem and such like be charged to a ``framework'' category}

Figure~\ref{line-counts} gives line counts for \name, 
organized by the major components shown in Figure~\ref{fig:full}.
We see that the proof ratio for the implementation code
is 4:1, which grows to 7:1 %for the system as a whole
%(i.e., when proving crash safety).
when including all system refinement proofs (``total'').
This is comparable to the distributed, in-memory key-value store
verified in previous work~\cite{ironfleet},
which also reports a 7:1 ratio.
However, \name's implementation is 3$\times$ larger,
indicating that automated verification techniques can scale
to larger systems without super-linear effort.

The results also show that \name's implementation is 
more than 5$\times$ larger than its specification,
giving us reason to hope that the specification
is less likely to contain bugs than an unverified implementation.

Figure~\ref{line-counts-micro}\footnotetext{The implementation of the hash table (with dynamic frames and the linear type system) and search tree (with dynamic frames) coincidentally have an identical line count, despite having substantial differences.} 
compares two \name components that we wrote using
both dynamic frames and linear reasoning.
The results show that
switching to linear reasoning saves tedium,
reducing proof overhead by 31-37\%. \todo{Overhead on verification time?}
%on the hash table
%and search tree subcomponents.

\paragraph{Measuring Proof-Attempt Latency}\label{eval:discipline}
To assess the success
of our discipline for keeping automation under control (\S\ref{sec:discipline}),
we measure the time to verify individual proof
units (e.g., definitions, methods, or lemmas) 
where developers spend most of their time waiting for verification results.
This approximately measures the developer's perception of the 
latency of the verification-development cycle.

\begin{figure}
\includegraphics[width=1\columnwidth]{figures/verification-times.pdf}
  \vspace{-5mm}
\caption{Cumulative distribution of verification times of function
  definitions, implementation methods, and proof lemmas. Most
  definitions--{\twentySecFasterPctile}--verify in less than {\twentySecSlowThresholdSecs}s,
  and {\tenSecFasterPctile}--verify in less than {\tenSecSlowThresholdSecs}s.
  }
  \vspace{-4mm}
\label{verification-times}
\end{figure}

Figure~\ref{verification-times} demonstrates that
{\name} almost always exhibits interactive verification times,
with {\tenSecFasterPctile} completing in under {\tenSecSlowThresholdSecs} seconds,
and {\twentySecFasterPctile} in under {\twentySecSlowThresholdSecs} seconds.
This suggests that our timeout-averse
development policy is effective.
The figure reveals that
we did not apply our policy with perfect consistency,
as {\twentySecNumSlowVerifications} proofs do take longer than
{\twentySecSlowThresholdSecs}s to verify.
Of those, {\twentySecPctOfSlowVerifsInvolvingHeap} involve dynamic frames.

It is difficult to capture every place we applied the policy. As a proxy,
we can count places where we explicitly hid a definition.
This both overestimates the cost of the policy 
(because sometimes we hid a definition based on intuition,
before observing a timeout) and underestimates it (because automation can
be controlled with other techniques).

\begin{figure}
\includegraphics[width=1\columnwidth]{figures/automation-figure.pdf}
  \vspace{-4mm}
  \caption{{\System} manually hides {\autoOpaqueCount} definitions ({\autoOpaquePct} of all system definitions),  revealing
them {\autoRevealCount} times.
This metric reflects the effort involved in explicitly managing automation.
}
  \vspace{-3mm}
\label{automation-histogram}
\end{figure}

These approximations aside, Figure~\ref{automation-histogram}
gives an idea of the burden of controlling automation.
We hid {\autoOpaqueCount} definitions,
{\autoOpaquePct} of all definitions in the system.
These hidden definitions are manually revealed {\autoRevealCount} times.
Of the {\autoOpaqueCount}, {\autoOpaqueZeroReveals} were never revealed:
the salient properties of the definition could be exported as a postcondition
without causing timeouts.
{\autoFivePct} of hidden definitions are revealed no more than five
times; their essential features are captured in lemmas or wrapped into
higher-level definitions.
%
%\cut{The most revealed definition is IsStrictlySorted, revealed {\autoRevealMax} times.}

Developers are typically less sensitive to the latency of continuous-integration
builds that check that the system as a whole still verifies.
For \name, these take 1.8 hours of CPU time,
but thanks to the inherent parallelism of modular verification,
complete in 11 minutes on 32 cloud machines.

\subsection{Performance}\label{sec:eval:perf}

Our performance evaluation is designed for two questions:
\begin{enumerate}
  \item Does {\name} demonstrate
    the insertion-performance gains of write-optimized data
    structures?
  \item Does our linear extension produce code with
    performance comparable to hand-written code using dynamic frames?
\end{enumerate}

All experiments were run on cloud instances with directly attached physical storage.
The HDD experiments and subcomponent microbenchmarks are run on AWS EC2 d2.xlarge instances, with 4x hardware hyperthreads on a 2.4 GHz Intel Xeon E5-2676 v3.
The SSD experiments are run on AWS EC2 i2.xlarge instances, with 4x hardware hyperthreads on a 2.5 GHz Intel E5-2670 v2 and a SATA SSD.

\newcommand{\ycsbHddDataDir}{data/ycsb-hdd/golden}
\newcommand{\ycsbSsdDataDir}{data/ycsb-ssd/golden}
\newcommand{\mutablemapDataDir}{data/mutablemap/boldgolden}
\newcommand{\mutablebtreeDataDir}{data/mutablebtree/silver}

\pgfplotsset{
  GenericBarplot/.style={
    ybar=1pt,
    bar width=5.5pt,
    height = 1.75in,
    enlarge x limits=0.1,
    major x tick style = transparent,
    scaled y ticks=false,
    ylabel near ticks,
    point meta=rawy,
    nodes near coords,
    every node near coord/.append style={rotate=90, anchor=west, font=\tiny},
    nodes near coords={\pgfmathprintnumber[fixed, precision=0]{\pgfplotspointmeta}},
    area legend,
    legend columns=-1,
    label style={font=\small, align=center},
  },
  MinMidMaxFile/.style n args={3}{
    /pgfplots/error bars/y dir=both,
    /pgfplots/error bars/y explicit,
    table/y error plus expr = (\thisrow{#3} - \thisrow{#2}),
    table/y error minus expr = (\thisrow{#2} - \thisrow{#1}),
  }
}

\pgfplotsset{
  YcsbThroughput/.style={
    GenericBarplot,
    symbolic x coords = {Load, A, B, C, D, F, Cuniform},
    ylabel={Operations/Second},
    MinMidMaxFile={MinThroughput}{MedianThroughput}{MaxThroughput},
  },
  YcsbLoadThroughput/.style={
    YcsbThroughput,
    width = 1.1in,
    restrict x to domain=0:0,
    xtick = {Load},
    xlabel={(from YCSB A)},
  },
  YcsbRunThroughput/.style={
    YcsbThroughput,
    width = 3.1in,
    xmin = A,
    xmax = Cuniform,
    restrict x to domain=1:10,
    xtick = {A, B, C, D, F, Cuniform},
    xticklabels = {A, B, C, D, F, U},
    xlabel={YCSB Workload},
  },
}

\pgfplotsset{
    LinearStyle/.style={teal, fill=teal!10!white, postaction={pattern=north west lines, pattern color=teal!50!white}},
    HeapStyle/.style={violet!50!black, fill=violet!50!black!10!white, postaction={pattern=north east lines, pattern color=violet!50!black!50!white}},
}

\pgfplotsset{
    BerkeleyStyle/.style={blue, fill=blue!10!white, postaction={pattern=crosshatch dots, pattern color=blue!50!white}},
    VeriHeapStyle/.style={HeapStyle},
    VeriLinearStyle/.style={LinearStyle},
    RocksStyle/.style={red, fill=red!05!white, postaction={pattern=crosshatch, pattern color=red!50!white}},
}

\begin{figure*}
\centering
%\ref{ycsb-legend}
\begin{subfigure}[c]{\textwidth}
  \centering
  \parbox[c]{0.1\textwidth}{\subcaption{HDD}} ~
  \parbox[c]{0.7\textwidth}{
  \begin{tikzpicture}
    \begin{axis}[
        YcsbLoadThroughput,
        ymode = log,
        ymax = 100000,
        ytick = { 100, 1000, 10000, 100000, 1000000 },
        legend columns=1,
        legend to name={ycsb-legend},
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbHddDataDir/berkeleydb.csv};
      \addlegendentry{BerkeleyDB}
      \addplot+ [VeriHeapStyle] table {\ycsbHddDataDir/veribetrkv.csv};
      \addlegendentry{\name-DF}
      \addplot+ [VeriLinearStyle] table {\ycsbHddDataDir/veribetrkv-linear.csv};
      \addlegendentry{\name}
      \addplot+ [RocksStyle] table {\ycsbHddDataDir/rocksdb.csv};
      \addlegendentry{RocksDB}
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{axis}[
        YcsbRunThroughput,
        ymin = 0,
        ymax = 800,
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbHddDataDir/berkeleydb.csv};
      \addplot+ [VeriHeapStyle] table {\ycsbHddDataDir/veribetrkv.csv};
      \addplot+ [VeriLinearStyle] table {\ycsbHddDataDir/veribetrkv-linear.csv};
      \addplot+ [RocksStyle] table {\ycsbHddDataDir/rocksdb.csv};
    \end{axis}
  \end{tikzpicture}
  }%
  \hspace{-0.6in}%
  \ref{ycsb-legend}
  \label{subfig:ycsb-hdd}
\end{subfigure}

\begin{subfigure}{\textwidth}
  \centering
  \parbox[c]{0.1\textwidth}{\subcaption{SSD}} ~
  \parbox[c]{0.7\textwidth}{
  \begin{tikzpicture}
    \begin{axis}[
        YcsbLoadThroughput,
        ymode = log,
        ymax = 200000,
        ytick = { 100, 1000, 10000, 100000, 1000000 },
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbSsdDataDir/berkeleydb.csv};
      % \addlegendentry{BerkeleyDB}
      \addplot+ [VeriHeapStyle] table {\ycsbSsdDataDir/veribetrkv-128k.csv};
      % \addlegendentry{\name-Dynamic Frames}
      \addplot+ [VeriLinearStyle] table {\ycsbSsdDataDir/veribetrkv-linear-128k.csv};
      % \addlegendentry{\name}
      \addplot+ [RocksStyle] table {\ycsbSsdDataDir/rocksdb.csv};
      % \addlegendentry{RocksDB}
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}
    \begin{axis}[
        YcsbRunThroughput,
        ymin = 0,
        ymax = 20000,
        ytick = { 5000, 10000, 15000, 20000, 1000000 },
      ]
      \addplot+ [BerkeleyStyle] table {\ycsbSsdDataDir/berkeleydb.csv};
      \addplot+ [VeriHeapStyle] table {\ycsbSsdDataDir/veribetrkv-128k.csv};
      \addplot+ [VeriLinearStyle] table {\ycsbSsdDataDir/veribetrkv-linear-128k.csv};
      \addplot+ [RocksStyle] table {\ycsbSsdDataDir/rocksdb.csv};
    \end{axis}
  \end{tikzpicture}
  }
  \hspace{0.9in}
  \label{subfig:ycsb-ssd}
\end{subfigure}
% %   \caption{
% %     Median throughput of YCSB workloads with 24B keys and 512B
% %     values running on a HDD and 2GB of RAM.  Load is 10M operations
% %     ($\approx$ 5GB of data) and runs are 10000 operations each.
% %     Workload ``U'' is a workload of uniformly random queries.
% %     Error
% %     bars indicate min/max of 6 runs.  Higher is
% %     better.}\label{fig:ycsb-hdd-512B}
\caption{
Median throughput of YCSB workloads values running on an HDD (\subref{subfig:ycsb-hdd}) and SSD (\subref{subfig:ycsb-ssd}) and 2GB of RAM. \name-DF is a version of our system with hand-tuned code using dynamic frames reasoning.  Load is 10M operations
($\approx$ 5GB of data) and runs are 10000 operations each.
Error bars indicate min/max of 6 runs.  Higher is better.
On an HDD (\subref{subfig:ycsb-hdd}), \name insertions are over $25\times$ faster than in
BerkeleyDB, but lag RocksDB, by about $9\times$.
\name queries are are about $4\times$ slower than both.
On an SSD (\subref{subfig:ycsb-ssd}), \name still beats BerkleyDB on random
insertions, but \name queries are slower than BerkeleyDB.
}
\label{fig:ycsb}
\end{figure*}

\subsubsection{YCSB}

\Cref{fig:ycsb} show throughput for \name,
BerkeleyDB, and RocksDB on the YCSB benchmarks~\cite{ycsb} on hard drive and SSD,
including the load phase for workload A, and a uniformly random query
workload (labeled as workload ``U'').

There are three main take-aways from these measurements.  First, \name
demonstrates the performance gains of using a write-optimized data
structure.  For the load phase on hard drive, which consists of a pure
random insertion workload, \name is over $25\times$ faster than
BerkeleyDB.  Even on SSD, where random I/O is much cheaper, \name
modestly outperforms BerkeleyDB on insertions.

In contrast, \name is roughly $8\times$ slower than RocksDB on hard
disk and $4\times$ slower on SSD.  We think this gap is due to two main
factors.  First, our implementation performs a sync whenever its
on-disk log fills up, whereas RocksDB performs sync only when the
application requests them.  Second, due to memory fragmentation and
other overheads, our effective cache size is approximately $4\times$
smaller than Rocks.  Ths is because Rocks primarily uses the OS page
cache, which can pack 4KB pages in memory, whereas we use an internal
(and, hence, verified!) cache of nodes.

The second main take-away is that queries in \name are a constant
factor slower than in a \btree: roughly $4\times$ slower than
BerkeleyDB on both hard drive and SSD.  It is well known that
write-optimized data structures need additional optimizations to match
\btree point-query performance.  For example, RocksDB is
also slower than BerkeleyDB on SSD and has relatively mixed
performance on hard disk.  The primary bottleneck for \name query
performance is that each cache miss loads an entire node, and nodes
are large (2MB on hard disk and 128K on SSD) in order to amortize I/O
overheads.  BerkeleyDB nodes, on the other hand, are only 16KB.
Industrial-grade \beptrees solve this problem with
``sub-nodes'', but we have not yet implemented this optimization.

The final take-away is that at a macro-level our linear implementation has essentially
the same performance as the version with hand-tuned code
using dynamic frames reasoning.  
%This demonstrates that our linear type system extension to
%Dafny enables us to write algorithms and data structures as if we are
%using immutable variables, but compile down to efficient,
%update-in-place code.

Overall, we conclude that \name demonstrates that a verified system
can achieve the performance gains of a write-optimized storage system,
but it needs further optimization to match highly-tuned commercial
implementations.

\begin{figure}
\centering
\ref{micro-legend}
\begin{subfigure}[t]{.25\textwidth}
  \begin{tikzpicture}
    \begin{axis}[
        GenericBarplot,
        enlarge x limits=0.2,
        width = 1.8in,
        symbolic x coords = {insert, readpositive, readnegative, remove},
        xtick = {insert, readpositive, readnegative, remove},
        xticklabels = {Insert, Pos. Query, Neg. Query, Delete},
        x tick label style={rotate=90},
        ymin = 0,
        ymax = 8300000,
        ytick = {0, 2000000, 4000000, 6000000, 8000000},
        yticklabels={0, 2M, 4M, 6M, 8M},
        ylabel={Operations/Second},
        legend to name={micro-legend},
        MinMidMaxFile={min}{med}{max},
      ]
      \addplot+ [HeapStyle] table {\mutablemapDataDir/master.csv};
      \addlegendentry{Dynamic Frames}
      \addplot+ [LinearStyle] table {\mutablemapDataDir/linear.csv};
      \addlegendentry{Linear}
    \end{axis}
  \end{tikzpicture}
  \subcaption{Hash table}
        \label{subfig:micro-hashtable}
\end{subfigure}~%
\begin{subfigure}[t]{.2\textwidth}
  \centering
  \parbox{1.7\textwidth}{
    \vspace{-2.475in}
  \begin{tikzpicture}
    \begin{axis}[
        GenericBarplot,
        enlarge x limits=0.5,
        width = 1.3in,
        symbolic x coords = {write,read},
        xtick = {write, read},
        xticklabels = {Insert, Query},
        x tick label style={rotate=90},
        ymin = 0,
        ymax = 300000,
        ylabel={Operations/Second},
        MinMidMaxFile={min}{med}{max},
      ]
      \addplot+ [HeapStyle] table {\mutablebtreeDataDir/8000000-master.csv};
      \addplot+ [LinearStyle] table {\mutablebtreeDataDir/8000000-linear.csv};
    \end{axis}
  \end{tikzpicture}}\\
        \subcaption{Search tree}
        \label{subfig:micro-btree}
\end{subfigure}
        \caption{Median throughput of subcomponent microbenchmarks.
Higher is better. Error bars are min/max of 6 runs.
%Our linear hash table (\subref{subfig:micro-hashtable}) is slightly faster,
%or at least comparable, than our hand-optimized implementation using
%dynamic frames reasoning. Throughput of successively inserting
%64 million key-value pairs, performing 64 million positive and
%negative queries, and emptying the hash table by removing each
%key. 
%Our linear in-memory search tree (\subref{subfig:micro-btree}) is comparable with our
%hand-optimized update-in-place version using dynamic frames reasoning.
%Throughput of inserting 8 million key-value pairs, and performing 8 million negative queries.
}\label{fig:datastructures-micro}
\end{figure}

\subsubsection{Linear Data Structures}

\Cref{fig:datastructures-micro} shows the performance of
our linear and dynamic-frames-based hash-table and search-tree implementations.
%
The main take-away from both experiments is that, even in
microbenchmarks, the linear and non-linear implementations have very
comparable performance.

The hash table benchmark inserts 64 million key-value
pairs, then performs as many positive and negative queries, and then
deletes everything in the hash table.  The linear version is slightly
faster than the non-linear version, except for deletes, which are
slightly slower.  We suspect the speedup comes primarily from the lack
of shared pointer overhead.

The search-tree benchmark measures the time to insert 8 million key-value
pairs and then query them all.  Performance for the linear version is
close to the non-linear version, but slightly faster for queries and
slightly slower for inserts.  We believe queries are faster due to the
elimination of shared pointer overheads, and the insertions are slower
due to the overheads of destructing nodes on the way down the tree and
reconstructing them on the way back up.  
%We could eliminate this
%overhead by adding support for mutably borrowing references to our
%linear type system.  The non-linear version was also written to be
%tail recursive.  This is not possible in the linear version, since it
%needs to reconstruct nodes as it unwinds the recursion.  Mutable
%borrow support would also eliminate this overhead.

Overall, we conclude that our linear type system enables us to
construct performant code without the challenges of dynamic frame
reasoning.

%% \begin{figure}
%% \includegraphics[width=1\columnwidth]{figures/btree-perf.pdf}
%% \caption{linear/repr perf}
%% \label{btree-perf}
%% \end{figure}

### 2020-09-01 tools/mutablebtree-cook-2.sh
add btree benchmark

@@ -1,34 +1,34 @@
 #!/bin/bash
 
+count=$1
+shift
 logs="$@"
-
-operations=("insert" "readpositive" "readnegative" "remove")
 
 function generate_output_line {
     operation="$1"
     if [ -z "$logs" ]; then
         return
     fi
-    throughputs=`sed -n "s,^$operation	\(.*\)$,\1,p" $logs`
+    throughputs=`sed -n "s,^$count\s*$operation\s*\(.*\)$,\1,p" $logs`
     if [ -z "$throughputs" ]; then
         return
     fi
-    throughputs=$(for t in $throughputs; do python3 -c "print(64000000 / $t * 1000)"; done)
+    throughputs=$(for t in $throughputs; do python3 -c "print($count / $t * 1000)"; done)
     sorted_throughputs=`echo "$throughputs" | sort -n`
     min=`echo "$sorted_throughputs" | head -1`
     max=`echo "$sorted_throughputs" | tail -1`
     nthroughputs=`echo "$sorted_throughputs" | wc -l`
     median_index=$[ $nthroughputs / 2 + 1 ]
     median=`echo "$sorted_throughputs" | head -$median_index | tail -1`
-    echo $operation  $median $min $max
+    echo $operation $median $min $max
 }
 
 function generate_output_file {
-    echo "phase  med min max"
+    echo "operation  med min max"
+    operations=(write read)
     for operation in ${operations[@]}; do
         generate_output_line $operation 
     done
 }
 
 generate_output_file
-

### 2020-09-01 tools/mutablebtree-cook-2.sh
add btree benchmarks to pdf

@@ -9,7 +9,7 @@ function generate_output_line {
     if [ -z "$logs" ]; then
         return
     fi
-    throughputs=`sed -n "s,^$count\s*$operation\s*\(.*\)$,\1,p" $logs`
+    throughputs=`sed -n "s,^$count[ \t]*$operation[ \t]*\(.*\)$,\1,p" $logs`
     if [ -z "$throughputs" ]; then
         return
     fi

### 2020-09-02 tools/plot/plot-verification-time.py
fix verification-constants.tex bug

@@ -299,7 +299,7 @@ def main():
     detect_bogus_workers(pile)
     report_slowest_symbols(pile)
     report_sequential_time(pile)
-    emit_constants(CDFStuff(pile).defs())
+    #emit_constants(CDFStuff(pile).defs())
     plot_all(pile)
 
 main()

### 2020-09-02 tools/ycsb-cook.sh
add ycsb to pdf

@@ -1,0 +1,41 @@
+#!/bin/bash
+
+logs="$@"
+
+function generate_output_line {
+    workload="$1"
+    mode="$2"
+    if [ -z "$logs" ]; then
+        return
+    fi
+    throughputs=`sed -n "s,^\[step\] $workload $mode throughput \(.*\) ops/sec$,\1,p" $logs`
+    if [ -z "$throughputs" ]; then
+        return
+    fi
+    sorted_throughputs=`echo "$throughputs" | sort -n`
+    min=`echo "$sorted_throughputs" | head -1`
+    max=`echo "$sorted_throughputs" | tail -1`
+    nthroughputs=`echo "$sorted_throughputs" | wc -l`
+    median_index=$[ $nthroughputs / 2 + 1 ]
+    median=`echo "$sorted_throughputs" | head -$median_index | tail -1`
+    if [ $mode == "load" ]; then
+        echo Load $median $min $max
+    else
+        echo $workload $median $min $max
+    fi
+    
+}
+
+workloads=("A" "Cuniform" "B" "C" "D" "F")
+
+function generate_output_file {
+    echo Workload  MedianThroughput  MinThroughput  MaxThroughput
+    for workload in ${workloads[@]}; do
+        if [ "$workload" == "A" ]; then
+            generate_output_line "$workload" "load"
+        fi
+        generate_output_line "$workload" "run"
+    done
+}
+
+generate_output_file

### 2020-09-02 tla-tree/build.sh
remove abandoned junk/, tla-tree/, and immutable-btree/ dirss

@@ -1,7 +1,0 @@
-#!/bin/bash
-dafny KVTypes.dfy
-dafny Disk.dfy
-dafny CrashableMap.dfy
-dafny CrashSafeLog.dfy
-dafny CrashSafeLogInv.dfy
-dafny CrashSafeLogRefinement.dfy

### 2020-09-02 tla-tree/fullbuild.sh
jonh's cleanup commits (squashed)
remove abandoned junk/, tla-tree/, and immutable-btree/ dirss

improve first-time dev docs

replace scary-sounding comments with coherent explanations

fix a couple offensive comments; rip out a lot of dead code

@@ -1,2 +1,0 @@
-#!/bin/bash
-time dafny KVTypes.dfy Disk.dfy MissingLibrary.dfy ImmutableDiskTree.dfy ImmutableDiskTreeInv.dfy ImmutableDiskTreeHeight.dfy ImmutableDiskTreeInterpretation.dfy ImmutableDiskTreeRefinement.dfy

### 2020-09-02 tools/prep-environment.sh
improve first-time dev docs

@@ -1,0 +1,18 @@
+#!/bin/bash
+#
+# Run this script as root (sudo tools/prep-environment) to install a recent
+# version of mono and a few other necessary tools.
+
+apt-get update
+apt-get install -y ca-certificates gnupg2
+apt-key adv --keyserver keyserver.ubuntu.com --recv-keys A6A19B38D3D831EF
+apt-get install make wget
+dafny_server/mono-official-stable.list 
+cat > /etc/apt/sources.list.d/mono-official-stable.list <<__EOF__
+deb https://download.mono-project.com/repo/ubuntu stable-bionic main
+__EOF__
+apt-get update
+apt-get install -y mono-runtime mono-mcs mono-devel make wget unzip
+pip3 install toposort
+
+tools/install-dafny.sh

### 2020-09-03 tools/linear_line_counts.py
remove git invocations from some of the benchmarking scripts

@@ -3,6 +3,7 @@
 A hand-tuned script for doing head-to-head line counts of linear component across branches.
 """
 
+import os
 import json
 import collections
 import subprocess
@@ -14,15 +15,12 @@ def do_cmd(cmd):
     subprocess.call(cmd)
 
 class Package:
-    def __init__(self, label, branch, sources):
+    def __init__(self, label, sources):
         self.label = label
-        self.branch = branch
         self.sources = sources
         self.counter = collections.Counter()
 
     def count(self):
-        do_cmd(["git", "checkout", self.branch])
-        print("NOW ON", cur_branch())
         for source in self.sources:
             self.count_one(source)
 
@@ -41,46 +39,37 @@ class Package:
     def jsondict(self):
         d = dict(self.counter)
         d["label"] = self.label
-        d["branch"] = self.branch
         d["sources"] = self.sources
         return d
 
 packages = [
-    # Package("Hashtable-linear", branch="osdi20-artifact-linear", sources=
-    #         ["lib/DataStructures/LinearMutableMap.i.dfy"]),
     # Package("BTree-linear", branch="osdi20-artifact-linear", sources=
     #         ["lib/DataStructures/BtreeModel.i.dfy",
     #             "lib/DataStructures/MutableBtree.i.dfy"]),
-    Package("Hashtable-master", branch="osdi20-artifact-dynamic-frames", sources=
-            ["lib/DataStructures/MutableMapModel.i.dfy",
-                "lib/DataStructures/MutableMapImpl.i.dfy"]),
-    Package("BTree-master", branch="osdi20-artifact-dynamic-frames", sources=
+    Package("BTree", sources=
             ["lib/DataStructures/BtreeModel.i.dfy",
-                "lib/DataStructures/MutableBtree.i.dfy"]),
+             "lib/DataStructures/MutableBtree.i.dfy"]),
+    Package("Hashtable-master", sources=
+            ["lib/DataStructures/MutableMapModel.i.dfy",
+             "lib/DataStructures/MutableMapImpl.i.dfy"]),
     ]
 
-def cur_branch():
-    return subprocess.run("git rev-parse --abbrev-ref HEAD".split(), stdout=subprocess.PIPE).stdout.decode("utf-8").strip()
+if os.path.exists("lib/DataStructures/LinearMutableMap.i.dfy"):
[…]

### 2020-09-03 bench/run-mutable-map.cpp
import mutable map harness from eval-mutablemap-linear

@@ -1,0 +1,13 @@
+#include "bench/LinearMutableMap.h"
+
+int main(int argc, char**argv) {
+    if (argc != 3) {
+        printf("invalid number of arguments\n");
+        return -1;
+    }
+    uint64 seed = atol(argv[1]);
+    bool dry = (strcmp(argv[2], "true") == 0);
+    printf("METADATA title running with seed %llu dry run %d\n", seed, dry);
+    MutableMapBench_Compile::__default::Run(seed, dry);
+    return 0;
+}

### 2020-09-03 bench/run-mutable-btree.cpp
import btree harness from eval-mutablemap-linear

@@ -1,0 +1,15 @@
+#include "lib/Lang/LinearExtern.h"
+#include "lib/DataStructures/MutableBtree.i.h"
+
+int main(int argc, char**argv) {
+    if (argc != 4) {
+        printf("invalid number of arguments\n");
+        return -1;
+    }
+    uint64 seed = atol(argv[1]);
+    uint64 ops = atol(argv[2]);
+    bool dry = (strcmp(argv[3], "true") == 0);
+    printf("METADATA title running %llu ops with seed %llu dry run %d\n", ops, seed, dry);
+    MainModule_Compile::__default::Run(seed, ops, dry);
+    return 0;
+}

### 2020-09-03 bench/run-mutable-btree.cpp
fix warning about printf types

@@ -9,7 +9,7 @@ int main(int argc, char**argv) {
     uint64 seed = atol(argv[1]);
     uint64 ops = atol(argv[2]);
     bool dry = (strcmp(argv[3], "true") == 0);
-    printf("METADATA title running %llu ops with seed %llu dry run %d\n", ops, seed, dry);
+    printf("METADATA title running %llu ops with seed %llu dry run %d\n", (unsigned long long)ops, (unsigned long long)seed, dry);
     MainModule_Compile::__default::Run(seed, ops, dry);
     return 0;
 }

### 2020-09-03 bench/run-mutable-map.cpp

@@ -7,7 +7,7 @@ int main(int argc, char**argv) {
     }
     uint64 seed = atol(argv[1]);
     bool dry = (strcmp(argv[2], "true") == 0);
-    printf("METADATA title running with seed %llu dry run %d\n", seed, dry);
+    printf("METADATA title running with seed %llu dry run %d\n", (unsigned long long)seed, dry);
     MutableMapBench_Compile::__default::Run(seed, dry);
     return 0;
 }

### 2020-09-03 bench/run-mutable-btree.cpp
fix warning about printf types

@@ -8,7 +8,7 @@ int main(int argc, char**argv) {
     uint64 seed = atol(argv[1]);
     uint64 ops = atol(argv[2]);
     bool dry = (strcmp(argv[3], "true") == 0);
-    printf("METADATA title running %llu ops with seed %llu dry run %d\n", ops, seed, dry);
+    printf("METADATA title running %llu ops with seed %llu dry run %d\n", (unsigned long long)ops, (unsigned long long)seed, dry);
     MainModule_Compile::__default::Run(seed, ops, dry);
     return 0;
 }

### 2020-09-03 tools/automation-study.py
incorporate more of Make-ified build system

@@ -74,10 +74,11 @@ def opaqueInstances():
     for iref in opaqueRefs:
         symbolName = symbolNameForOpaqueIref(iref)
         #print(symbolName)
-        reveal_count = len(grepAll("reveal_%s" % symbolName))
+        reveal_lines = grepAll("reveal_%s\\b" % symbolName)
+        reveal_count = len(reveal_lines)
         record = {"file": iref.normPath, "line": iref.line_num, "symbol":symbolName, "reveal_count":reveal_count}
         records.append(record)
-        #print("revealed %2d times: %s in %s at %s" % (reveal_count, symbolName, iref.normPath, iref.line_num))
+        print("revealed %2d times: %s in %s at %s" % (reveal_count, symbolName, iref.normPath, iref.line_num))
         #sys.stdout.flush()
     return records
 
@@ -88,20 +89,21 @@ def scrape():
     open(INTERMEDIATE, "w").write(json.dumps(records, indent = 4))
 
 def plot(records):
-    fig = plt.figure()
-    ax = fig.add_subplot(211)
+    fig = plt.figure(figsize=(6*0.8,2.5*0.8))
+    ax = fig.add_subplot(111)
     reveals = [record["reveal_count"] for record in records]
     reveals.sort()
     ax.hist(reveals, bins=max(reveals))
     ax.set_ylabel("count of\nopaque definitions")
 
-    ax = fig.add_subplot(212)
+    a2 = ax.twinx()
     xs = reveals
     ys = [i/(len(reveals)-1.0) for i in range(len(reveals))]
-    ax.plot(xs, ys)
-    ax.set_ylabel("cum. frac. of\nopaque definitions")
+    a2.plot(xs, ys, color="red")
+    a2.set_ylabel("cum. frac. of\nopaque definitions")
     ax.set_xlabel("number of times manually revealed")
-    fig.savefig("../veripapers/osdi2020/figures/automation-figure.pdf")
+    plt.tight_layout()
+    fig.savefig("build/automation-figure.pdf")
 
 def gather_constants(records):
     reveals = [record["reveal_count"] for record in records]
@@ -149,11 +151,11 @@ def totalDefinitions(defs):
     })
     return defs
 
-def emit_constants(defs):
-    with open("../veripapers/osdi2020/data/automation-constants.tex", "w") as fp:
-        fp.write("% generated by tools/automation-study.py\n")
-        for k,v in defs.items():
[…]

### 2020-09-03 tools/line_counter_report_lib.py

@@ -3,6 +3,11 @@ import lib_deps
 import json
 import os
 
+# make build/Impl/Bundle.i.lcreport -j4
+# (incremental report redraw:)
+# tools/line_counter.py --mode report --input Impl/Bundle.i.dfy --output build/Impl/Bundle.i.lcreport
+# cp build/Impl/Bundle.i.lcreport ../veripapers/osdi2020/data/line-counts.tex
+
 def loadReport(iref):
     fp = open(lib_deps.targetName(iref, ".lc"))
     values = json.load(fp)
@@ -19,7 +24,7 @@ def gatherReports(input):
 def accumulate(reports, mapper):
     counters = {}
     def getCounter(report):
-        key = mapper(report)
+        key = mapper.key(report)
         if key not in counters:
             counters[key] = collections.Counter()
         return counters[key]
@@ -29,9 +34,45 @@ def accumulate(reports, mapper):
         del counter["source"]
     return counters
 
+row_labels = {
+    "ManualMapper-map": "Map, 3States\\template{Map}",
+    "ManualMapper-abstractbetree": "Abstract{\\beptree}",
+    "ManualMapper-betree": "{\\beptree}",
+    "ManualMapper-compositeviewmap": "CompositeViewMap",
+    "ManualMapper-betreejournaliosystem": "{\\beptree}{\iosystem}",
+    "ManualMapper-concreteiosystem": "Concrete{\iosystem}",
+    "ManualMapper-impl": "implementation code",
+    "ManualMapper-lib": "libraries",
+        }
+
+def write_tex_table(fp, counters):
+    fp.write("\\begin{tabular}{|l|rrr|}\n")
+    fp.write("\\hline\n")
+    fp.write("Major component & spec & impl & proof \\\\\n")
+    fp.write("\\hline\n")
+
+    def write_row(label, rowdata):
+        fp.write("%s & %d & %d & %d \\\\\n" % (
+                label,
+                rowdata["spec"],
+                rowdata["impl"],
+                rowdata["proof"]
+                ))
+
+    keys = list(counters.keys())
+    keys.sort()
+    for key,label in row_labels.items():
+        if key in row_labels and key in counters:
+            label = row_labels[key]
+            write_row(label, counters[key])
+    fp.write("\\hline\n")
+    write_row("total", counters["AllMapper-all"])
+    fp.write("\\hline\n")
+    fp.write("\\end{tabular}\n")
+
[…]

### 2020-09-03 tools/linear_line_counts.py

@@ -1,0 +1,99 @@
+#!/usr/bin/env python3
+"""
+A hand-tuned script for doing head-to-head line counts of linear component across branches.
+"""
+
+import os
+import json
+import collections
+import subprocess
+import line_count_lib
+import line_counter_report_lib
+
+def do_cmd(cmd):
+    print("CMD", " ".join(cmd))
+    subprocess.call(cmd)
+
+class Package:
+    def __init__(self, label, sources):
+        self.label = label
+        self.sources = sources
+        self.counter = collections.Counter()
+
+    def count(self):
+        for source in self.sources:
+            self.count_one(source)
+
+    def count_one(self, source):
+        counter = line_count_lib.Counter(".")
+        dafnyFile = line_count_lib.DafnyFile(source, 0.0)
+        counter.collect_line_counts([dafnyFile])
+        self.counter["spec"] += dafnyFile.spec
+        self.counter["impl"] += dafnyFile.impl
+        self.counter["proof"] += dafnyFile.proof
+        self.display("intermediate")
+
+    def display(self, msg):
+        print(msg, self.label, self.counter)
+
+    def jsondict(self):
+        d = dict(self.counter)
+        d["label"] = self.label
+        d["sources"] = self.sources
+        return d
+
+packages = [
+    # Package("BTree-linear", branch="osdi20-artifact-linear", sources=
+    #         ["lib/DataStructures/BtreeModel.i.dfy",
+    #             "lib/DataStructures/MutableBtree.i.dfy"]),
+    Package("BTree", sources=
+            ["lib/DataStructures/LinearBtreeModel.i.dfy",
+             "lib/DataStructures/LinearMutableBtree.i.dfy"]),
+    Package("Hashtable", sources=
+            ["lib/DataStructures/LinearMutableMap.i.dfy"]),
+    ]
+
+DATA_FILE = "data/linear_lines.json"
+def collect():
+    accum = []
+    for package in packages:
+        package.count()
+        accum.append(package.jsondict())
+    open(DATA_FILE, "w").write(json.dumps(accum, indent=2))
+
+# Translation table from package labels to tex macro names & table labels.
+tex_names = {
+        "Hashtable": {"macroPrefix": "Hashtable", "tableModule": "Hashtable", "tableMode": "(curr. branch)"},
[…]

### 2020-09-03 tools/mutablebtree-cook-2.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+count=$1
+shift
+logs="$@"
+
+function generate_output_line {
+    operation="$1"
+    if [ -z "$logs" ]; then
+        return
+    fi
+    throughputs=`sed -n "s,^$count[ \t]*$operation[ \t]*\(.*\)$,\1,p" $logs`
+    if [ -z "$throughputs" ]; then
+        return
+    fi
+    throughputs=$(for t in $throughputs; do python3 -c "print($count / $t * 1000)"; done)
+    sorted_throughputs=`echo "$throughputs" | sort -n`
+    min=`echo "$sorted_throughputs" | head -1`
+    max=`echo "$sorted_throughputs" | tail -1`
+    nthroughputs=`echo "$sorted_throughputs" | wc -l`
+    median_index=$[ $nthroughputs / 2 + 1 ]
+    median=`echo "$sorted_throughputs" | head -$median_index | tail -1`
+    echo $operation $median $min $max
+}
+
+function generate_output_file {
+    echo "operation  med min max"
+    operations=(write read)
+    for operation in ${operations[@]}; do
+        generate_output_line $operation 
+    done
+}
+
+generate_output_file

### 2020-09-03 tools/mutablemap-cook.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+logs="$@"
+
+operations=("insert" "readpositive" "readnegative" "remove")
+
+function generate_output_line {
+    operation="$1"
+    if [ -z "$logs" ]; then
+        return
+    fi
+    throughputs=`sed -n "s,^$operation	\(.*\)$,\1,p" $logs`
+    if [ -z "$throughputs" ]; then
+        return
+    fi
+    throughputs=$(for t in $throughputs; do python3 -c "print(64000000 / $t * 1000)"; done)
+    sorted_throughputs=`echo "$throughputs" | sort -n`
+    min=`echo "$sorted_throughputs" | head -1`
+    max=`echo "$sorted_throughputs" | tail -1`
+    nthroughputs=`echo "$sorted_throughputs" | wc -l`
+    median_index=$[ $nthroughputs / 2 + 1 ]
+    median=`echo "$sorted_throughputs" | head -$median_index | tail -1`
+    echo $operation  $median $min $max
+}
+
+function generate_output_file {
+    echo "phase  med min max"
+    for operation in ${operations[@]}; do
+        generate_output_line $operation 
+    done
+}
+
+generate_output_file
+

### 2020-09-03 tools/plot/plot-verification-time.py

@@ -9,8 +9,7 @@ import collections
 import tarfile
 
 #EXPERIMENT="expresults/veri_time_13-*"
-EXPERIMENT="expresults/veri_time_13.tgz"
-SLOW_THRESH = 20
+EXPERIMENT="build/verification-times.tgz"
 
 class Observation:
     def __init__(self, time, worker_name, source_filename):
@@ -32,26 +31,41 @@ class SymbolReport:
     def times(self):
         return [obs.time for obs in self.observations]
 
-    def mean(self):
-        return np.mean(self.times())
+    def avg(self):
+        # called "avg" so I don't feel bad changing from mean to median
+        return np.median(self.times())
 
     def stdev(self):
         return np.std(self.times())
 
     def cv(self):
-        return self.stdev() / self.mean()
+        return self.stdev() / self.avg()
 
 class SymbolPile:
     def __init__(self):
         self.symbols = {}
+        self.avg_times_ = None
 
     def put_replica(self, name, obs):
+        assert not self.avg_times_
         if name not in self.symbols:
             self.symbols[name] = SymbolReport(name)
         self.symbols[name].add(obs)
 
     def reports(self):
         return list(self.symbols.values())
+
+    def avg_times(self):
+        if self.avg_times_ == None:
+            self.avg_times_ = [report.avg() for report in self.reports()]
+            self.avg_times_.sort()
+        return self.avg_times_ 
+
+    def N(self):
+        return len(self.avg_times())
+
+    def cdist(self, i):
+        return i*1.0/self.N()
 
 def parse_one(opener, pile):
     symbol_parts = {}
@@ -182,36 +196,13 @@ def get_declaration(source_filename, symbol):
 def report_slowest_symbols(pile):
     """report on slowest symbols. Are they dynamic-frames-repr disasters?"""
 
-    reports = pile.reports()
-    reports.sort(key = lambda r: -r.mean())
-    repr_set = set()
-    no_repr_set = set()
-    for r in reports:
-        if r.mean()<=SLOW_THRESH: break
-        #print("slowest", r.name, r.mean())
-        source_filename = r.observations[0].source_filename
-        decl = get_declaration(source_filename, r.name)
-        if decl and decl.is_repr():
-            repr_set.add(r)
[…]

### 2020-09-03 tools/run-map-config-experiment.py

@@ -1,0 +1,73 @@
+#!/usr/bin/env python3
+
+from __future__ import print_function
+from __future__ import division
+
+import sys
+import os
+import subprocess
+import time
+import datetime
+import signal
+
+def actuallyprint(msg):
+    print(msg)
+    sys.stdout.flush()
+
+def main():
+  seed=None
+  output=None
+
+  print("arguments", sys.argv)
+
+  for arg in sys.argv[1:]:
+    if arg.startswith("seed="):
+      seed = arg[len("seed=") : ]
+    elif arg.startswith("output="):
+      output = arg[len("output=") : ]
+    else:
+      assert False, "unrecognized argument: " + arg
+
+  value_updates = []
+  for (name, value) in value_updates:
+    print("setting " + name + " to " + value)
+    # splice_value_into_bundle(name, value)
+
+  actuallyprint("Building executable...")
+  sys.stdout.flush()
+
+  dafny_cmd = ".dafny/dafny/Binaries/dafny /noVerify /spillTargetCode:3 /countVerificationErrors:0 /compileTarget:cpp lib/DataStructures/MutableMapImpl.i.dfy framework/NativeArithmetic.h framework/NativeArrays.h framework/LinearCongruentialGenerator.h"
+  actuallyprint(dafny_cmd)
+  ret = os.system(dafny_cmd)
+  assert ret == 0
+
+  dafny_cmd_2 = ".dafny/dafny/Binaries/dafny /noVerify /spillTargetCode:3 /countVerificationErrors:0 /compileTarget:cpp bench/MutableMap.dfy framework/NativeArithmetic.h framework/NativeArrays.h framework/LinearCongruentialGenerator.h"
+  actuallyprint(dafny_cmd_2)
+  ret = os.system(dafny_cmd_2)
+  assert ret == 0
+
+  cmd = "g++ -O3 bench/bench/MutableMap.cpp bench/run-mutable-map.cpp framework/NativeArithmetic.cpp -o MutableMapBench  -I .dafny/dafny/Binaries/ -I lib/DataStructures/ -I bench -Ilib -std=c++17 -I."
+  actuallyprint(cmd)
+  ret = os.system(cmd)
+  assert ret == 0
+
+  # bitmask indicating which CPUs we can use
+  # See https://linux.die.net/man/1/taskset
+  taskset_cmd = "taskset 4 "
+
+
+  with open(output, 'w') as f:
+      f.write("METADATA map perf comparison\n")
[…]

### 2020-09-04 bench/run-mutable-btree.cpp
got microbenchmarks to build

@@ -1,5 +1,5 @@
-#include "lib/Lang/LinearExtern.h"
-#include "lib/DataStructures/MutableBtree.i.h"
+#include "LinearExtern.h"
+#include "lib/DataStructures/LinearMutableBtree.i.h"
 
 int main(int argc, char**argv) {
     if (argc != 4) {

### 2020-09-04 framework/Framework.h

@@ -5,6 +5,7 @@
 #include "NativeArrays.h"
 #include "LinearExtern.h"
 #include "LinearBox.h"
+#include "LinearCongruentialGenerator.h"
 
 #include <map>
 #include <unordered_map>

### 2020-09-04 framework/NativeArithmetic.h

@@ -1,0 +1,7 @@
+#pragma once
+
+#include <cstdint>
+
+namespace NativeArithmetic_Compile {
+  uint64_t u64add(uint64_t a, uint64_t b);
+}

### 2020-09-04 tools/checkout-dafny-commit.sh

@@ -20,4 +20,5 @@ set -x
 
 # https://github.com/secure-foundations/dafny.git
 # cpp branch
-git checkout aa60a3bac2e3aed375af867c49affcf0519e7840
+git checkout 791f0e14f28c25dfb2cc2720f18772a2f22d76b5
+

### 2020-09-04 tools/checkout-dafny-commit.sh
update dafny commit

@@ -18,7 +18,7 @@
 set -e
 set -x
 
-commit=eaa754140b98dd91f8cd5560cd687b625262c5d8
+commit=791f0e14f28c25dfb2cc2720f18772a2f22d76b5
 if [ $1 ]; then
    commit=$1
 fi

### 2020-09-04 tools/checkout-dafny-commit.sh
update dafny commit

@@ -18,7 +18,7 @@
 set -e
 set -x
 
-commit=791f0e14f28c25dfb2cc2720f18772a2f22d76b5
+commit=0c324a76930c7710732dfd271d3fe2ba4d528868
 if [ $1 ]; then
    commit=$1
 fi

### 2020-09-04 tools/checkout-dafny-commit.sh
update dafny commit

@@ -19,6 +19,6 @@ set -e
 set -x
 
 # https://github.com/secure-foundations/dafny.git
-# cpp branch
-git checkout 791f0e14f28c25dfb2cc2720f18772a2f22d76b5
+# osdi20-artifact branch
+git checkout 0c324a76930c7710732dfd271d3fe2ba4d528868
 

### 2020-09-04 tools/linear_line_counts.py
pull in some cleanups from linear branch

@@ -49,13 +49,10 @@ packages = [
     Package("BTree", sources=
             ["lib/DataStructures/BtreeModel.i.dfy",
              "lib/DataStructures/MutableBtree.i.dfy"]),
-    Package("Hashtable-master", sources=
+    Package("Hashtable", sources=
             ["lib/DataStructures/MutableMapModel.i.dfy",
              "lib/DataStructures/MutableMapImpl.i.dfy"]),
     ]
-
-if os.path.exists("lib/DataStructures/LinearMutableMap.i.dfy"):
-    packages = packages + [ Package("Hashtable-linear", sources=["lib/DataStructures/LinearMutableMap.i.dfy"]) ]
 
 DATA_FILE = "data/linear_lines.json"
 def collect():
@@ -67,8 +64,7 @@ def collect():
 
 # Translation table from package labels to tex macro names & table labels.
 tex_names = {
-        "Hashtable-linear": {"macroPrefix": "HashtableLinear", "tableModule": "Hashtable", "tableMode": "linear"},
-        "Hashtable-master": {"macroPrefix": "HashtableRepr", "tableModule": "Hashtable", "tableMode": "dyn. frames"},
+        "Hashtable": {"macroPrefix": "Hashtable", "tableModule": "Hashtable", "tableMode": "(curr. branch)"},
         "BTree": {"macroPrefix": "BTree", "tableModule": "BTree", "tableMode": "(curr. branch)"},
         }
 

### 2020-09-04 tools/checkout-dafny-commit.sh
switch to working dafny version

@@ -20,5 +20,4 @@ set -x
 
 # https://github.com/secure-foundations/dafny.git
 # osdi20-artifact branch
-git checkout 0c324a76930c7710732dfd271d3fe2ba4d528868
-
+git checkout 51e6315a040b3316f00a8542f19be0487c5e2353

### 2020-09-04 tools/checkout-dafny-commit.sh
remove some assumes from MutableMap benchmark, switch to working dafny version

@@ -18,7 +18,7 @@
 set -e
 set -x
 
-commit=0c324a76930c7710732dfd271d3fe2ba4d528868
+commit=51e6315a040b3316f00a8542f19be0487c5e2353
 if [ $1 ]; then
    commit=$1
 fi

### 2020-09-04 examples/tree.rs
remove examples

@@ -1,64 +1,0 @@
-#[allow(dead_code)]
-
-extern crate prusti_contracts;
-use prusti_contracts::*;
-
-pub struct Node {
-    left: Option<Box<Node>>,
-    right: Option<Box<Node>>,
-    value: u64,
-}
-
-#[pure]
-pub fn search_tree(node: &Node, from: u64, to: u64) -> bool {
-    (from <= node.value && node.value < to) &&
-    (match &node.left {
-        Some(ref left) => search_tree(left, from, node.value),
-        None => true,
-    }) &&
-    (match &node.right {
-        Some(ref right) => search_tree(right, node.value, to),
-        None => true,
-    })
-}
-
-#[requires="search_tree(tree, ghost_from, ghost_to)"]
-#[requires="0 <= value && value < 1024"]
-#[requires="ghost_from <= value && value < ghost_to"]
-#[ensures="search_tree(tree, ghost_from, ghost_to)"]
-pub fn insert(tree: &mut Node, value: u64, ghost_from: u64, ghost_to: u64) {
-    if value < tree.value {
-        match &mut tree.left {
-            Some(ref mut left) => {
-                insert(left, value, ghost_from, tree.value);
-            },
-            None => {
-                tree.left = Some(Box::new(Node { left: None, right: None, value: value }));
-            },
-        }
-    } else {
-        match &mut tree.right {
-            Some(ref mut right) => {
-                insert(right, value, tree.value, ghost_to);
-            },
-            None => {
-                tree.right = Some(Box::new(Node { left: None, right: None, value: value }));
-            },
-        }
-    }
-}
-
-#[requires="search_tree(tree, 0, 1024)"]
-#[requires="0 <= value && value < 1024"]
-#[ensures="search_tree(tree, 0, 1024)"]
-pub fn do_insert(tree: &mut Node, value: u64) {
-    insert(tree, value, 0, 1024);
-}
-
-#[requires="search_tree(tree1, 0, 1024)"]
-#[requires="search_tree(tree2, 0, 1024)"]
-#[ensures="search_tree(tree1, 0, 1024)"]
-#[ensures="search_tree(tree2, 0, 1024)"]
-pub fn dont_interfere(tree1: &mut Node, tree2: &mut Node) {
-    do_insert(tree1, 82);
-}

### 2020-09-04 malloc-exp/plot-heap-vs-malloc.py
remove malloc stuff

@@ -1,68 +1,0 @@
-#!/usr/bin/env python3
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-import re
-import sys
-
-def parse(filename):
-    t = 0
-    os_map_total = {}
-    os_map_heap = {}
-    malloc_total = {}
-    for line in open(filename, "r").readlines():
-        fields = line.split()
-        if line.startswith("os-map-total"):
-            os_map_total[t] = int(fields[1])
-            os_map_heap[t] = int(fields[3])
-            malloc_total[t] = int(fields[5])
-            t += 1
-        if line.startswith("ma-coarse-histogram"):
-            size = fields[1]
-            allocations[size][t] = int(fields[5))
-        if line.startswith("ma-total"):
-            allocations[size]["total"] = int(fields[4))
-        if line.startswith("veribetrkv [op] sync"):
-            completed[t] = int(fields[4])
-
-    xs = range(t)
-    fig, axes = plt.subplots(4, 1)
-
-    rate = [float(completed[t])/t for t in completed]
-    line, = axes[1].plot(xs, rate);
-
-#    ax2 = ax1.twinx()
-#
-#    map_ratios = [float(os_map_total[t])/malloc_total[t] for t in xs]
-#    line, = ax2.plot(xs, map_ratios, color="orange", linestyle="--")
-#    line.set_label("map-malloc ratio")
-#    heap_ratios = [float(os_map_heap[t])/malloc_total[t] for t in xs]
-#    line, = ax2.plot(xs, heap_ratios, "r", linestyle="--")
-#    line.set_label("heap-malloc ratio")
-#
-#    ax2.set_ylim(bottom = 0, top=3)
-#    ax2.set_ylabel("x/malloc ratio")
-#    ax2.legend(loc = "lower right")
-#
-#    trunc = 500 if len(map_ratios)>500 else 0 # initial reports are thrown off by weight of the executable
-#    max_ratio = max(map_ratios[trunc:])   
-#    max_t = map_ratios.index(max_ratio)
-#    max_msg = "map-malloc ratio\nmax: %.2f @ %ds" % (max_ratio, max_t)
-#    print(max_msg)
-#    plt.text(max_t, max_ratio, max_msg)
-#
-#    scaleGB = 1e-9;
-#    line, = ax1.plot(xs, [os_map_total[t]*scaleGB for t in xs], color="orange")
[…]

### 2020-09-04 malloc-exp/plot-histogram.py
remove malloc stuff

@@ -1,69 +1,0 @@
-#!/usr/bin/env python3
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-import re
-#import json
-
-def parse_one_histogram(line):
-    #return json.loads(line)
-    assert line[0] == "{"
-    assert line[-2:] == "}\n"
-    line = line[1:-2]
-    pairs = line.split(",")[:-1]
-    histo = {}
-    for pair in pairs:
-        size,count = map(int, pair.split(":"))
-        if count>0:
-            histo[size] = count
-    return histo
-
-def cdf(histo, by_size):
-    sizes = list(histo.keys())
-    sizes.sort()
-    xs = []
-    ys = []
-    accum = 0
-    for size in sizes:
-        count = histo[size]
-        accum += count * size if by_size else count
-        xs.append(size)
-        ys.append(accum)
-    #print(xs)
-    # normalize ys to 0..1
-    ys = [y/float(accum) for y in ys]
-    #print(ys)
-    return xs, ys
-
-def parse():
-    t = 0
-    proc_heap = {}
-    malloc_total = {}
-    histos = {}
-    for line in open("malloc-exp/histograms", "r").readlines():
-        if line.startswith("proc-heap"):
-            fields = line.split()
-            proc_heap[t] = int(fields[1])
-            malloc_total[t] = int(fields[3])
-            t += 1
-        
-        if line.startswith("{"):
-            histos[t] = parse_one_histogram(line)
-
-    max_histo_t = max(histos.keys())
-    print(max_histo_t)
-    max_histo = histos[max_histo_t]
-    print(max_histo)
-
-    # accumulate the CDF
-    line, = plt.plot(*cdf(max_histo, True))
-    line.set_label("by size")
-    line, = plt.plot(*cdf(max_histo, False))
-    line.set_label("by allocation count")
-    plt.xscale("log")
-    plt.legend()
-    
-    plt.savefig("malloc-exp/size-cdf.png")
-    #plt.show()
-    
-parse()

### 2020-09-04 tools/aws/automation.py
resolve tools/aws conflict

@@ -1,193 +1,0 @@
-SSH_ID_PEM="~/.ssh/veribetrfsbastion.pem"
-
-import json
-import subprocess
-import time
-import select
-import sys
-import os
-import fcntl
-import termcolor    # pip3 install termcolor
-
-logfile = None
-
-def set_logfile(path):
-    if os.path.exists(path):
-        sys.stderr.write("logfile %s exists; change experiment label?\n" % path)
-        sys.exit(-1)
-    global logfile
-    logfile = open(path, "w")
-
-def log(msg):
-    sys.stdout.write(msg + "\n")
-    sys.stdout.flush()
-    if logfile:
-        logfile.write(msg + "\n")
-        logfile.flush()
-
-def retrieve_running_workers(ssd=False):
-    workers_pipe = subprocess.Popen("ssh bastion veribetrfs/tools/aws/describe-instances.py --running --json".split() + (["--ssd"] if ssd else []), stdout=subprocess.PIPE)
-    workers_json,_ = workers_pipe.communicate()
-    workers = json.loads(workers_json)
-    return workers
-
-def ssh_preamble():
-    return "ssh -o UserKnownHostsFile=/dev/null -o StrictHostKeyChecking=no -i".split() + [SSH_ID_PEM]
-
-def ssh_target_for_worker(worker):
-    return"ubuntu@%s" % worker["PublicIpAddress"]
-
-def ssh_cmd_for_worker(worker):
-    return ssh_preamble() + [ssh_target_for_worker(worker)]
-
-next_index = 0
-def get_index():
-    global next_index
-    result = next_index
-    next_index = next_index + 1
-    return result
-
-spectrum = ["red", "yellow", "green", "cyan", "blue", "magenta"]
-
-class WorkerMonitor:
-    def __init__(self, worker, cmd):
-        self.worker = worker
-        self.cmd = cmd
-        self.running = True
-        if "index" not in worker:
-            worker["index"] = get_index()
-        self.index = worker["index"]
-
-    def set_pipe(self, pipe):
-        self.pipe = pipe
-        fd = self.pipe.stdout.fileno()
-        flags = fcntl.fcntl(fd, fcntl.F_GETFL)
-        fcntl.fcntl(fd, fcntl.F_SETFL, flags | os.O_NONBLOCK)
-
-    def __repr__(self):
-        return self.worker["Name"]
-
-    def linetag(self):
[…]

### 2020-09-04 tools/aws/kill-jobs.py

@@ -1,27 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--ssd', action='store_true', help="select ssd machines")
-args = parser.parse_args()
-
-def cmd_for_idx(idx, worker):
-    return Command("kill", ssh_cmd_for_worker(worker) + ["killall", "-q", "VeribetrfsYcsb", "RocksYcsb", "make", "z3", "mono", "dafny", "MutableBtreeBench", "MutableBtreeBench"])
-
-def main():
-    workers = retrieve_running_workers(ssd=args.ssd)
-    blacklist = [
-        "veri-worker-b00", 
-        "veri-worker-b01", 
-        "veri-worker-b02", 
-        "veri-worker-b03", 
-        "veri-worker-b04", 
-        "veri-worker-b05", 
-    ]
-    workers = [w for w in workers if w["Name"] not in blacklist]
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/launch-btree.py

@@ -1,58 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-from suite import *
-
-common_vars = [
-    Variable("ops", "run_btree", [
-        Value("2pows", "ops=[1,2,3,4,5,6,7,8]") ]),
-    Variable("seed", "run_btree", [
-        Value("seed0", "seed=348725789"),
-        Value("seed1", "seed=34578699348"),
-        Value("seed2", "seed=23945765203"),
-        Value("seed3", "seed=625478238923"),
-        Value("seed4", "seed=657843290874"),
-        Value("seed5", "seed=9526729840202"),
-        ]),
-    ]
-repr_suite = Suite(
-    "repr",
-    Variable("git_branch", "git_branch", [Value("master", "eval-btree-master")]),
-    *common_vars)
-linear_suite = Suite(
-    "linear",
-    Variable("git_branch", "git_branch", [Value("linear", "eval-btree-linear")]),
-    *common_vars)
-suite = ConcatSuite("silver-btree-millions-cc", linear_suite, repr_suite)
-
-MBTREE_PATH="./tools/run-btree-config-experiment.py"
-
-def cmd_for_idx(idx, worker):
-    variant = suite.variants[idx]
-    cmd = (ssh_cmd_for_worker(worker) + [
-        "cd", "veribetrfs", ";",
-        "git", "clean", "-fd", ".", ";",
-        "sh", "tools/clean-for-build.sh", variant.git_branch(), ";",
-        ] + ["python3", MBTREE_PATH] + ["git_branch=" + variant.git_branch()] + [variant.valmap[var].param_value for var in variant.vars_of_type("run_btree")] + ["output=../"+variant.outfile()]
-        )
-    return Command(str(variant), cmd)
-
-def main():
-    set_logfile(suite.logpath())
-    log("PLOT tools/aws/pull-results.py")
-    log("VARIANTS %s" % suite.variants)
-
-    workers = retrieve_running_workers(ssd=False)
-    blacklist = [
-        "veri-worker-b00",
-        "veri-worker-b01",
-        "veri-worker-b02",
-        "veri-worker-b03",
-        "veri-worker-b04",
-        "veri-worker-b05",
-    ]
-    workers = [w for w in workers if w["Name"] not in blacklist]
-    worker_pipes = launch_worker_pipes(workers, len(suite.variants), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/launch-mutablemap.py

@@ -1,51 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-from suite import *
-
-common_vars = [
-    Variable("seed", "run_map", [
-        Value("seed0", "seed=11348725789"),
-        Value("seed1", "seed=34578699348"),
-        Value("seed2", "seed=23945765203"),
-        Value("seed3", "seed=625478238923"),
-        Value("seed4", "seed=657843290874"),
-        Value("seed5", "seed=9526729840202"),
-        Value("seed6", "seed=95229840202"),
-        ]),
-    ]
-repr_suite = Suite(
-    "repr",
-    Variable("git_branch", "git_branch", [Value("master", "eval-mutablemap-master")]),
-    *common_vars)
-linear_suite = Suite(
-    "linear",
-    Variable("git_branch", "git_branch", [Value("linear", "eval-mutablemap-linear")]),
-    *common_vars)
-suite = ConcatSuite("bold-golden-mutablemap-master-00", repr_suite)
-
-MBTREE_PATH="./tools/run-map-config-experiment.py"
-
-def cmd_for_idx(idx, worker):
-    variant = suite.variants[idx]
-    cmd = (ssh_cmd_for_worker(worker) + [
-        "cd", "veribetrfs", ";",
-        "git", "clean", "-fd", ".", ";",
-        "sh", "tools/clean-for-build.sh", variant.git_branch(), ";",
-        ] + ["python3", MBTREE_PATH] + ["git_branch=" + variant.git_branch()] + [variant.valmap[var].param_value for var in variant.vars_of_type("run_map")] + ["output=../"+variant.outfile()]
-        )
-    return Command(str(variant), cmd)
-
-def main():
-    set_logfile(suite.logpath())
-    log("PLOT tools/aws/pull-results.py")
-    log("VARIANTS %s" % suite.variants)
-
-    workers = retrieve_running_workers()
-    blacklist = [
-    ]
-    workers = [w for w in workers if w["Name"] not in blacklist]
-    worker_pipes = launch_worker_pipes(workers, len(suite.variants), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/launch-ycsb-benchmarks.py

@@ -1,85 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-from suite import *
-
-replica_count = 6
-
-common_vars = [
-    Variable("cgroup",     "run_veri",   [Value("yescgroup", "cgroup=True")]),
-    Variable("ram",        "run_veri",   [Value("2gb",       "ram=2.0gb")]),
-    Variable("device",     "run_veri",   [Value("disk",      "device=disk")]),
-    Variable("workload",   "run_veri",
-             [Value("all", "workload=" + ",".join(
-                    ["ycsb/workloada-onefield.spec", # load
-                    "ycsb/workloadc-uniform.spec"  , # Doesn't modify database, so sneak it in here
-                    "ycsb/workloada-onefield.spec" , # runs...
-                    "ycsb/workloadb-onefield.spec" ,
-                    "ycsb/workloadc-onefield.spec" ,
-                    "ycsb/workloadd-onefield.spec" ,
-                    "ycsb/workloadf-onefield.spec" ,
-                    ])),]),
-    #Variable("duration", "run_veri", [Value("2h", "time_budget=2h")]),
-    Variable("replica",  "silent", [Value("r{}".format(r), "r={}".format(r)) for r in range(replica_count)]),
-    ]
-veri_suite = Suite(
-    "veribetrkv",
-    Variable("git_branch", "git_branch", [
-        Value("master",    "master"),
-        Value("linear",    "linear-disintegration"),
-        ]),
-    Variable("system",     "run_veri", [Value("veri", "veri")]),
-    Variable("rowcache",   "run_veri", [Value("norowcache", "")]),
-                                        # Value("yesrowcache", "use_unverified_row_cache")]),
-    Variable("cachesize", "run_veri", [Value("cachesize200m", "cacheSize={}".format(200*1024*1024))]),
-    Variable("bucketweight", "run_veri", [
-        Value("bucketweight2pow{}".format(p), "bucketWeight={}".format((2**p)*1024)) for p in [7]]),
-    *common_vars)
-
-common_vars_others = common_vars + [
-    Variable("git_branch", "git_branch", [Value("master",    "master")]),
-    ]
-
-rocks_suite = Suite(
-    "rocksdb",
-    Variable("system",  "run_veri", [Value("rocks", "rocks")]),
-    Variable("filters", "run_veri", [Value("nofilters", ""),
[…]

### 2020-09-04 tools/aws/pull-results.py

@@ -1,19 +1,0 @@
-#!/usr/bin/python3
-
-import subprocess
-from automation import *
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--ssd', action='store_true', help="select ssd machines")
-args = parser.parse_args()
-
-def cmd_for_idx(idx, worker):
-    return Command("pull", ["rsync", "-zrav", "-e", " ".join(ssh_preamble()), ssh_target_for_worker(worker)+":expresults/", "expresults/"])
-
-def main():
-    workers = retrieve_running_workers(ssd=args.ssd)
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/run-all.py

@@ -1,31 +1,0 @@
-#!/usr/bin/python3
-
-import sys
-from automation import *
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--ssd', action='store_true', help="select ssd machines")
-parser.add_argument('--dry-run', action='store_true')
-parser.add_argument('command')
-args = parser.parse_args()
-
-def cmd_for_idx(idx, worker):
-    return Command("run", ssh_cmd_for_worker(worker) + [args.command])
-
-def main():
-    workers = retrieve_running_workers(ssd=args.ssd)
-    blacklist = []
-    # blacklist = [
-    #     "veri-worker-b00", 
-    #     "veri-worker-b01", 
-    #     "veri-worker-b02", 
-    #     "veri-worker-b03", 
-    #     "veri-worker-b04", 
-    #     "veri-worker-b05", 
-    # ]
-    workers = [w for w in workers if w["Name"] not in blacklist]
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=args.dry_run)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/suite.py

@@ -1,116 +1,0 @@
-import os
-
-class Value:
-    def __init__(self, label, param_value):
-        self.label = label
-        self.param_value = param_value
-
-    def __repr__(self):
-        return "#(%s)" % self.label
-
-class Variable:
-    def __init__(self, label, type, values):
-        self.label = label
-        self.type = type
-        self.values = values
-        for value in self.values:
-            value.variable = self
-
-    def __repr__(self):
-        return "$(%s)" % self.label
-
-class Variant:
-    def __init__(self, suite, values):
-        self.suite = suite
-        self.values = values
-        self.valmap = dict([(value.variable, value) for value in values])
-        self.make_label()
-
-    def make_label(self):
-        parts = []
-        for val in self.values:
-            val_label_set = self.suite.value_label_set_for_variable_label(val.variable.label)
-            if len(val_label_set)>1:
-                parts.append("%s" % val.label)
-        self.label_ = "-".join(parts)
-
-    def get_label(self):
-        return self.label_
-
-    def __repr__(self):
-        return self.get_label()
-
-    def outfile(self):
-        return "expresults/%s-%s.data" % (self.suite.label, self.get_label())
-
-    def vars_of_type(self, type_label):
-        return [val.variable for val in self.values if val.variable.type == type_label]
-
-    def value_by_name(self, label):
-        match = [val for val in self.values if val.variable.label == label]
-        assert len(match)==1
-        return match[0]
-
-    def git_branch(self):
-        return self.valmap[self.vars_of_type("git_branch")[0]].param_value
-
-    def run_veri_params(self):
-        return [self.valmap[var].param_value for var in self.vars_of_type("run_veri")]
-
-class BaseSuite:
-    def __init__(self, label):
-        self.label = label
-
-    def png_filename(self):
-        return "%s.png" % self.label
-
-    def plot_command(self):
-        return " ".join([
-            "tools/plot/perf-compare.py",
-            "output=%s" % self.png_filename()] + [variant.get_label()+"="+variant.outfile() for variant in self.variants])
-
-    def logpath(self):
-        return os.path.join("logs", self.label+".log")
-
-class Suite(BaseSuite):
-    def __init__(self, label, *variables):
-        super().__init__(label)
-        self.variables = variables
-        self.variables_by_label = dict([(v.label,v) for v in self.variables])
-        self.variants = self.make_variants(self.variables)
-
-    def vars_of_type(self, type):
-        return [var for var in self.variables if var.type==type]
-
[…]

### 2020-09-04 tools/aws/verification-time-experiment.py

@@ -1,66 +1,0 @@
-#!/usr/bin/python3
-import sys
-import os
-sys.path.append(os.path.dirname(__file__)+"/..")
-
-import re
-from automation import *
-from suite import *
-from lib_deps import *
-
-# Consider updating all the workers before running the experiment!
-# tools/aws/run-all.py "cd veribetrfs; git checkout master; git pull"
-# tools/aws/run-all.py "cd veribetrfs; tools/update-submodules.sh; tools/update-dafny.sh"
-ROOT="Impl/Bundle.i.dfy"
-#ROOT="lib/DataStructures/MutableBtree.i.dfy"   # a small test case
-SUITE_NAME="veri_time_16"   # one big parallel build
-N_REPLICAS=1
-
-def listSources():
-    paths = set()
-    for source in depsFromDfySources([ROOT]):
-        # depsFromDfySources seems to produce some duplication; not sure why
-        paths.add(source.normPath)
-
-    values = []
-    for path in paths:
-        clean_path = re.sub("[^A-Za-z0-9-]", "_", path)
-        values.append(Value(clean_path, path))
-    return values
-
-def constructSuite(nReplicas):
-    sourceVariable = Variable("source", "source", listSources()[::-1])
-    replicaVariable = Variable("replica", "silent", [Value("r%d"%i, "r%d"%i) for i in range(nReplicas)])
-    suite = Suite(SUITE_NAME, sourceVariable, replicaVariable)
-    return suite
-
-RUN_VERI_PATH="tools/run-veri-config-experiment.py"
-
-def main():
-    def cmd_for_idx(idx, worker):
-        variant = suite.variants[idx]
-        source_path = variant.value_by_name("source").param_value
-        output_path = "../" + variant.outfile()
-        cmd = (ssh_cmd_for_worker(worker) + [
-            "cd", "veribetrfs", ";",
-            "echo", "WORKER", worker["Name"], ">", output_path, ";",
-    #        "sleep", "1",
-            "tools/local-dafny.sh", source_path, "/compile:0", "/trace", ">>", output_path
-#            "tools/local-dafny.sh", source_path, "/compile:0", "/trace", "/noNLarith", ">>", output_path
-            ]
-            )
-        return Command(str(variant), cmd)
-
-    suite = constructSuite(N_REPLICAS)
-    set_logfile(suite.logpath())
-    #log("VARIANTS %s" % suite.variants)
-
[…]

### 2020-09-04 tools/aws/automation.py
more cleanup

@@ -1,193 +1,0 @@
-SSH_ID_PEM="~/.ssh/veribetrfsbastion.pem"
-
-import json
-import subprocess
-import time
-import select
-import sys
-import os
-import fcntl
-import termcolor    # pip3 install termcolor
-
-logfile = None
-
-def set_logfile(path):
-    if os.path.exists(path):
-        sys.stderr.write("logfile %s exists; change experiment label?\n" % path)
-        sys.exit(-1)
-    global logfile
-    logfile = open(path, "w")
-
-def log(msg):
-    sys.stdout.write(msg + "\n")
-    sys.stdout.flush()
-    if logfile:
-        logfile.write(msg + "\n")
-        logfile.flush()
-
-def retrieve_running_workers():
-    workers_pipe = subprocess.Popen("ssh bastion veribetrfs/tools/aws/describe-instances.py --running --json".split(), stdout=subprocess.PIPE)
-    workers_json,_ = workers_pipe.communicate()
-    workers = json.loads(workers_json)
-    return workers
-
-def ssh_preamble():
-    return "ssh -o UserKnownHostsFile=/dev/null -o StrictHostKeyChecking=no -i".split() + [SSH_ID_PEM]
-
-def ssh_target_for_worker(worker):
-    return"ubuntu@%s" % worker["PublicIpAddress"]
-
-def ssh_cmd_for_worker(worker):
-    return ssh_preamble() + [ssh_target_for_worker(worker)]
-
-next_index = 0
-def get_index():
-    global next_index
-    result = next_index
-    next_index = next_index + 1
-    return result
-
-spectrum = ["red", "yellow", "green", "cyan", "blue", "magenta"]
-
-class WorkerMonitor:
-    def __init__(self, worker, cmd):
-        self.worker = worker
-        self.cmd = cmd
-        self.running = True
-        if "index" not in worker:
-            worker["index"] = get_index()
-        self.index = worker["index"]
-
-    def set_pipe(self, pipe):
-        self.pipe = pipe
-        fd = self.pipe.stdout.fileno()
-        flags = fcntl.fcntl(fd, fcntl.F_GETFL)
-        fcntl.fcntl(fd, fcntl.F_SETFL, flags | os.O_NONBLOCK)
-
-    def __repr__(self):
-        return self.worker["Name"]
-
-    def linetag(self):
-        return termcolor.colored("[%s]" % self, spectrum[self.index % len(spectrum)])
-
[…]

### 2020-09-04 tools/aws/describe-instances.py

@@ -1,60 +1,0 @@
-#!/usr/bin/python3
-import argparse
-import urllib, json, sys
-import boto3
-import pprint
-import json
-from collections import namedtuple
-from botocore.exceptions import ClientError
-
-def describeInstances(showAll=False, showRunning=False, ssd=False):
-    response = ec2_connection.describe_instances()
-    # print(prettyResponse(response))
-
-    def extract_salient(inst):
-        return Instance(
-                Name=[x for x in inst['Tags'] if x['Key'] == 'Name'][0]['Value'],
-                InstanceId=inst['InstanceId'],
-                PublicIpAddress=inst.get('PublicIpAddress'),
-                State=inst['State']['Name'])
-    insts = [extract_salient(x) for reservation in response['Reservations'] for x in reservation['Instances']]
-    insts.sort(key=lambda x: x.Name)
-    if not showAll:
-        insts = [x for x in insts if x.Name.startswith(
-            'veri-ssd' if ssd else 'veri-worker')]
-    if showRunning:
-        insts = [x for x in insts if x.State == 'running']
-    return insts
-
-if __name__=="__main__":
-    parser = argparse.ArgumentParser()
-    parser.add_argument('--all', action='store_true', help="don't filter down to workers, list all instances")
-    parser.add_argument('--ssh', action='store_true', help="print summary with ssh commands")
-    parser.add_argument('--json', action='store_true', help="print output as json")
-    parser.add_argument('--ssd', action='store_true', help="select ssd machines")
-    parser.add_argument('--running', action='store_true', help="only show running instances")
-    args = parser.parse_args()
-
-    Instance = namedtuple('Instance', ['Name', 'InstanceId', 'PublicIpAddress', 'State'])
-
-    ec2_connection = boto3.client('ec2', region_name='us-east-2')
-    try:
-        insts = describeInstances(showAll = args.all, showRunning = args.running, ssd = args.ssd)
-        if args.ssh:
-            for ist in insts:
-                print("\033[1m{}\033[0m \x1b[34m{}\033[0m\tssh ubuntu@{}".format(ist.Name, ist.State, ist.PublicIpAddress))
-        elif args.json:
-            print(json.dumps([
-                {
-                    'Name': inst.Name,
-                    'InstanceId': inst.InstanceId,
-                    'PublicIpAddress': inst.PublicIpAddress,
-                    'State': inst.State
-                } for inst in insts]))
-        else:
[…]

### 2020-09-04 tools/aws/init_again.sh

@@ -1,6 +1,0 @@
-#! /bin/bash
-
-mkfs.ext4 /dev/xvde
-mount /mnt/xvde
-chown ubuntu:ubuntu /mnt/xvde
-cgcreate -a ubuntu:ubuntu -t ubuntu:ubuntu -g memory:VeribetrfsExp

### 2020-09-04 tools/aws/kill-jobs.py

@@ -1,13 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-
-def cmd_for_idx(idx, worker):
-    return Command("kill", ssh_cmd_for_worker(worker) + ["killall", "-q", "VeribetrfsYcsb", "RocksYcsb", "make", "z3", "mono", "dafny"])
-
-def main():
-    workers = retrieve_running_workers()
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/launch-experiments.py

@@ -1,49 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-from suite import *
-
-common_vars = [
-    Variable("ram", "run_veri", [Value("2gb", "ram=2.0gb")]),
-    Variable("device", "run_veri", [Value("disk", "device=disk")]),
-    Variable("workload", "run_veri", [Value("wka6m", "workload=ycsb/wka-uniform-rc6000k.spec")]),
-    Variable("duration", "run_veri", [Value("2h", "time_budget=2h")]),
-    Variable("replica", "silent", [Value("r0", "r=0"), Value("r1", "r=1")]),
-    ]
-veri_suite = Suite(
-    "veri",
-    Variable("git_branch", "git_branch", [Value("suspless", "suspend-less-la2"), Value("la2", "leak-adventure-2")]),
-    Variable("nodeCountFudge", "run_veri", [Value(str(f), "nodeCountFudge="+str(f)) for f in [0.5]]),
-    Variable("system", "run_veri", [Value("veri1m", "config-1mb")]),
-    Variable("max_children", "run_veri", [Value("fanout16", "max_children=16")]),
-    Variable("cgroup", "run_veri", [Value("yescgroup", "cgroup=True")]),
-    *common_vars)
-rocks_suite = Suite(
-    "rocks",
-    Variable("git_branch", "git_branch", [Value("la2", "leak-adventure-2")]),
-    Variable("system", "run_veri", [Value("rocks", "rocks")]),
-    *common_vars)
-suite = ConcatSuite("robj-010", veri_suite, rocks_suite)
-
-RUN_VERI_PATH="tools/run-veri-config-experiment.py"
-
-def cmd_for_idx(idx, worker):
-    variant = suite.variants[idx]
-    cmd = (ssh_cmd_for_worker(worker) + [
-        "cd", "veribetrfs", ";",
-        "sh", "tools/clean-for-build.sh", variant.git_branch(), ";",
-        ]
-        + [RUN_VERI_PATH] + variant.run_veri_params() + ["output=../"+variant.outfile()]
-        )
-    return Command(str(variant), cmd)
-
-def main():
-    set_logfile(suite.logpath())
-    log("PLOT tools/aws/pull-results.py && %s && eog %s" % (suite.plot_command(), suite.png_filename()))
-    log("VARIANTS %s" % suite.variants)
-
-    workers = retrieve_running_workers()
[…]

### 2020-09-04 tools/aws/launch-ycsb-benchmarks.py

@@ -1,81 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-from suite import *
-
-replica_count = 6
-
-common_vars = [
-    Variable("cgroup",     "run_veri",   [Value("yescgroup", "cgroup=True")]),
-    Variable("ram",        "run_veri",   [Value("2gb",       "ram=2.0gb")]),
-    Variable("device",     "run_veri",   [Value("disk",      "device=disk")]),
-    Variable("workload",   "run_veri",
-             [Value("all", "workload=" + ",".join(
-                    ["ycsb/workloada-onefield.spec", # load
-                    "ycsb/workloadc-uniform.spec"  , # Doesn't modify database, so sneak it in here
-                    "ycsb/workloada-onefield.spec" , # runs...
-                    "ycsb/workloadb-onefield.spec" ,
-                    "ycsb/workloadc-onefield.spec" ,
-                    "ycsb/workloadd-onefield.spec" ,
-                    "ycsb/workloadf-onefield.spec" ,
-                    ])),]),
-    #Variable("duration", "run_veri", [Value("2h", "time_budget=2h")]),
-    Variable("replica",  "silent", [Value("r{}".format(r), "r={}".format(r)) for r in range(replica_count)]),
-    ]
-veri_suite = Suite(
-    "veribetrkv",
-    Variable("git_branch", "git_branch", [
-        Value("master",    "master"),
-        Value("linear",    "linear-disintegration"),
-        ]),
-    Variable("system",     "run_veri", [Value("veri", "veri")]),
-    Variable("rowcache",   "run_veri", [Value("norowcache", ""),
-                                        Value("yesrowcache", "use_unverified_row_cache")]),
-    *common_vars)
-
-common_vars_others = common_vars + [
-    Variable("git_branch", "git_branch", [Value("master",    "master")]),
-    ]
-
-rocks_suite = Suite(
-    "rocksdb",
-    Variable("system",  "run_veri", [Value("rocks", "rocks")]),
-    Variable("filters", "run_veri", [Value("nofilters", ""),
-                                     Value("yesfilters", "use_filters")]),
-    *common_vars_others)
-berkeley_suite = Suite(
-    "berkeleydb",
-    Variable("system", "run_veri", [Value("berkeley", "berkeley")]),
-    *common_vars_others)
[…]

### 2020-09-04 tools/aws/ls-jobs.py

@@ -1,13 +1,0 @@
-#!/usr/bin/python3
-
-from automation import *
-
-def cmd_for_idx(idx, worker):
-    return Command("ls", ssh_cmd_for_worker(worker) + ["ps", "-auxww", "|", "grep", "Ycsb", "|", "grep", "-v", "grep"])
-
-def main():
-    workers = retrieve_running_workers()
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/pull-results.py

@@ -1,14 +1,0 @@
-#!/usr/bin/python3
-
-import subprocess
-from automation import *
-
-def cmd_for_idx(idx, worker):
-    return Command("pull", ["rsync", "-zrav", "-e", " ".join(ssh_preamble()), ssh_target_for_worker(worker)+":expresults/", "expresults/"])
-
-def main():
-    workers = retrieve_running_workers()
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/run-all.py

@@ -1,14 +1,0 @@
-#!/usr/bin/python3
-
-import sys
-from automation import *
-
-def cmd_for_idx(idx, worker):
-    return Command("run", ssh_cmd_for_worker(worker) + sys.argv[1:])
-
-def main():
-    workers = retrieve_running_workers()
-    worker_pipes = launch_worker_pipes(workers, len(workers), cmd_for_idx, dry_run=False)
-    monitor_worker_pipes(worker_pipes)
-
-main()

### 2020-09-04 tools/aws/scale-workers.py

@@ -1,72 +1,0 @@
-#!/usr/bin/python3
-import urllib, json, sys
-import argparse
-import boto3
-import pprint
-from collections import namedtuple
-from botocore.exceptions import ClientError
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--scale-up', type=int, help='scale up to target worker count')
-parser.add_argument('--scale-down', type=int, help='scale down to target worker count')
-parser.add_argument('--ssd', action='store_true', help='act on the ssd cluster')
-parser.add_argument('--dry-run', action='store_true')
-
-args = parser.parse_args()
-
-print("Args: ", args)
-
-print("Fetching instances.")
-
-Instance = namedtuple('Instance', ['Name', 'InstanceId', 'PublicIpAddress', 'State'])
-
-def extract_salient(inst):
-    return Instance(
-            Name=[x for x in inst['Tags'] if x['Key'] == 'Name'][0]['Value'],
-            InstanceId=inst['InstanceId'],
-            PublicIpAddress=inst.get('PublicIpAddress'),
-            State=inst['State']['Name'])
-
-try:
-    ec2_connection = boto3.client('ec2', region_name='us-east-2')
-    response = ec2_connection.describe_instances()
-except ClientError as e:
-    print(e)
-    sys.exit(-1)
-
-insts = [extract_salient(x) for reservation in response['Reservations'] for x in reservation['Instances']]
-insts.sort(key=lambda x: x.Name)
-insts = [x for x in insts if x.Name.startswith('veri-ssd' if args.ssd else 'veri-worker')]
-
-print("Found instances:")
-pprint.pprint(insts, indent=2)
-
-instance_ids = [x.InstanceId for x in insts]
-
-if args.scale_up is None and args.scale_down is None:
-    print("either --scale-up or --scale-down is required")
-    sys.exit(-1)
-
-if args.scale_up is not None:
-    to_start = instance_ids[:args.scale_up]
-    print("starting instances:", to_start)
-
-    if not args.dry_run:
-        try:
-            response = ec2_connection.start_instances(InstanceIds=to_start)
-            pprint.pprint(response, indent=2)
-        except ClientError as e:
-            print(e)
-
-elif args.scale_down is not None:
-    to_stop = instance_ids[args.scale_down:]
-    print("stopping instances:", to_stop)
-
-    if not args.dry_run:
-        try:
-            response = ec2_connection.stop_instances(InstanceIds=to_stop)
[…]

### 2020-09-04 tools/aws/shell.py

@@ -1,24 +1,0 @@
-#!/usr/bin/python3
-
-import sys
-from automation import *
-
-
-#>def ssh_cmd_for_worker(worker):
-
-def main():
-    target = sys.argv[1]
-    workers = retrieve_running_workers()
-    matches = [w for w in workers if target in w["Name"]]
-    if len(matches)==0:
-        sys.stderr.write("No running worker matches '%s':\n" % target)
-        sys.stderr.write(" ".join([w["Name"] for w in workers])+"\n")
-        sys.exit(-1)
-    if len(matches)>1:
-        sys.stderr.write("Multiple running workers match '%s':\n" % target)
-        sys.stderr.write(" ".join([w["Name"] for w in matches])+"\n")
-        sys.exit(-1)
-    worker = matches[0]
-    subprocess.call(ssh_cmd_for_worker(worker))
-
-main()

### 2020-09-04 tools/aws/verification-time-experiment.py

@@ -1,65 +1,0 @@
-#!/usr/bin/python3
-import sys
-import os
-sys.path.append(os.path.dirname(__file__)+"/..")
-
-import re
-from automation import *
-from suite import *
-from lib_deps import *
-
-# Consider updating all the workers before running the experiment!
-# tools/aws/run-all.py "cd veribetrfs; git checkout master; git pull"
-# tools/aws/run-all.py "cd veribetrfs; tools/update-submodules.sh; tools/update-dafny.sh"
-ROOT="Impl/Bundle.i.dfy"
-#ROOT="lib/DataStructures/MutableBtree.i.dfy"   # a small test case
-SUITE_NAME="veri_time_13"   # comment branch merged
-N_REPLICAS=5
-
-def listSources():
-    paths = set()
-    for source in depsFromDfySources([ROOT]):
-        # depsFromDfySources seems to produce some duplication; not sure why
-        paths.add(source.normPath)
-
-    values = []
-    for path in paths:
-        clean_path = re.sub("[^A-Za-z0-9-]", "_", path)
-        values.append(Value(clean_path, path))
-    return values
-
-def constructSuite(nReplicas):
-    sourceVariable = Variable("source", "source", listSources())
-    replicaVariable = Variable("replica", "silent", [Value("r%d"%i, "r%d"%i) for i in range(nReplicas)])
-    suite = Suite(SUITE_NAME, sourceVariable, replicaVariable)
-    return suite
-
-RUN_VERI_PATH="tools/run-veri-config-experiment.py"
-
-def main():
-    def cmd_for_idx(idx, worker):
-        variant = suite.variants[idx]
-        source_path = variant.value_by_name("source").param_value
-        output_path = "../" + variant.outfile()
-        cmd = (ssh_cmd_for_worker(worker) + [
-            "cd", "veribetrfs", ";",
-            "echo", "WORKER", worker["Name"], ">", output_path, ";",
-    #        "sleep", "1",
-            "tools/local-dafny.sh", source_path, "/compile:0", "/trace", ">>", output_path
-            ]
-            )
-        return Command(str(variant), cmd)
-
-    suite = constructSuite(N_REPLICAS)
-    set_logfile(suite.logpath())
-    #log("VARIANTS %s" % suite.variants)
-
-    for variant in suite.variants:
-        log("VARIANT %s" % variant)
-    log("NUM_SOURCES %s" % len(listSources()))
[…]

### 2020-09-04 tools/big-rename.sh
cleaning up tools

@@ -1,25 +1,0 @@
-cd disk-betree
-../tools/rename.sh MapSpec.s.dfy
-../tools/rename.sh ThreeStateVersionedMap.s.dfy
-../tools/rename.sh ThreeStateVersioned.s.dfy
-../tools/rename.sh AsyncDiskModel.s.dfy
-../tools/rename.sh UI.s.dfy
-../tools/rename.sh UIStateMachine.s.dfy
-cd ..
-
-cd lib
-../tools/rename.sh NativeTypes.s.dfy
-../tools/rename.sh Maps.s.dfy
-../tools/rename.sh Option.s.dfy
-../tools/rename.sh total_order.s.dfy
-../tools/rename.sh sequences.s.dfy
-cd ..
-
-cd disk-betree
-echo *.dfy| tr ' ' '\n'  | grep -v '\.[is].dfy$' | sed 's#^./##' | sed 's#.dfy$#.i.dfy#' | xargs -n 1 ../tools/rename.sh
-cd ..
-
-cd lib
-echo *.dfy| tr ' ' '\n'  | grep -v '\.[is].dfy$' | sed 's#^./##' | sed 's#.dfy$#.i.dfy#' | xargs -n 1 ../tools/rename.sh
-cd ..
-

### 2020-09-04 tools/clean-for-build.sh

@@ -1,10 +1,0 @@
-#!/bin/sh
-# Args: git-branch to check out
-git fetch
-git checkout $*
-git pull
-git rev-parse --abbrev-ref HEAD
-tools/update-submodules.sh
-tools/update-dafny.sh
-rm -rf build/
-make clean

### 2020-09-04 tools/leakfinder_analyze.py

@@ -1,48 +1,0 @@
-#!/usr/bin/env python3
-import re
-
-class Trace:
-    def __init__(self, count, hash):
-        self.count = count
-        self.hash = hash
-        self.frames = []
-
-    def __lt__(self, other):
-        return self.hash < other.hash
-
-    def character(self):
-        return hash(tuple(self.frames[4:]))
-
-class Capture:
-    def __init__(self):
-        self.load()
-
-    def load(self):
-        self.table = {}
-        for line in open("/tmp/trace-1", "r").readlines():
-            mo = re.search("(trace|expanded): count ([-0-9]*).*hash *(.*)", line)
-            if mo!=None:
-                mode,count,hash = mo.groups()
-                if mode=="trace": continue
-                trace = Trace(int(count), int(hash, 16))
-                self.table[hash] = trace
-            else:
-                pc = int(line.strip(), 16)
-                trace.frames.append(pc)
-
-    def study(self):
-        byChar = []
-        for trace in self.table.values():
-            byChar.append((trace.character(), trace))
-        byChar.sort()
-        lastChar = None
-        total = 0
-        for (character, trace) in byChar:
-            if lastChar and lastChar != character:
-                print()
-            print("%016x %016x %d" % (character&0xffffffffffffffff, trace.hash, trace.count))
-            total += trace.count
-            lastChar = character
-        print("total", total)
-
-Capture().study()

### 2020-09-04 tools/profile.sh

@@ -1,2 +1,0 @@
-#!/bin/bash
-dafny /z3opt:smt.qi.profile=true /z3opt:smt.qi.profile_freq=100 /timeLimit:20 "$@"

### 2020-09-04 tools/rename.sh

@@ -1,9 +1,0 @@
-#!/bin/bash
-# example: rename.sh MapSpec.s.dfy
-newname=$1
-basename=`echo $newname | sed s/\.[is].dfy$/.dfy/`
-find /home/jonh/veribetrfs -type f -name \*.dfy | xargs sed -i "s/$basename/$newname/g"
-mv $basename $newname
-git rm $basename
-git add $newname
-git add --all

### 2020-09-04 tools/renaminator.py

@@ -1,101 +1,0 @@
-#!/usr/bin/env python3
-# Automation for moving dfy files among directories, cleaning up include references.
-
-import os
-import subprocess
-
-EXCLUDED_DIRS = set([".dafny", "build"])
-
-class Renaminator:
-    def __init__(self):
-        self.catalog()
-        # Apply the include path fixes first, since they're expressed relative
-        # to the source path locations.
-        self.mkdirCmds = []
-        self.fixCmds = []
-        self.gitAddCmds = []    # some of these will be pre-move referrers, so do them first
-        self.gitCmds = []
-
-    def catalog(self):
-        paths = []
-        for root, dirs, files in os.walk("."):
-            parts = root.split("/")
-            if len(parts) == 1:
-                # must be "."
-                top = parts[0]
-            else:
-                top = parts[1]
-            if top in EXCLUDED_DIRS or (top[0]=='.' and len(top)>1):
-                continue
-            for file in files:
-                if not file.endswith(".dfy"):
-                    continue
-                fullpath = os.path.join(root, file)
-                paths.append(fullpath)
-        self.paths = paths
-
-    def findSourceDir(self, filename):
-        matchingSourcePaths = [path for path in self.paths if path.endswith("/"+filename)]
-        if len(matchingSourcePaths) == 0:
-            raise Exception("No path matches %s" % filename)
-        if len(matchingSourcePaths) > 1:
-            raise Exception("Multiple paths match %s: %s" % (filename, matchingSourcePaths))
-        path = matchingSourcePaths[0]
-        return path[:-(len(filename)+1)]
-
-    def containsLine(self, filepath, testString):
-        contents = open(filepath).read()
-        return testString in contents
-
-    def fixReferrer(self, referrer, sourceDir, sourceFilename, destDir, destFilename):
-        referrerPath = os.path.split(referrer)[0]
-        sourceRelative = os.path.relpath(os.path.join(sourceDir, sourceFilename), referrerPath)
-        destRelative = os.path.relpath(os.path.join(destDir, destFilename), referrerPath)
-        expectInclude = 'include "%s"' % sourceRelative
-        newInclude = 'include "%s"' % destRelative
-        if self.containsLine(referrer, expectInclude):
[…]

### 2020-09-04 tools/roslyn-csc.sh

@@ -1,14 +1,0 @@
-#! /bin/bash
-
-# if a recent dotnet core is installed, run csc with the first sdk version found
-# by running dotnet --list-sdks
-
-set -e
-
-SDK_VER=`dotnet --list-sdks | head -n 1 | sed 's/\(.*\) .*/\1/'`
-DOTNET_DIR=`dotnet --list-sdks | head -n 1 | sed 's/.* \[\(.*\)\]/\1/'`
-SDK_DIR="$DOTNET_DIR/$SDK_VER"
-
-echo dotnet core $SDK_VER in $SDK_DIR
-
-dotnet $SDK_DIR/Roslyn/bincore/csc.dll /r:$SDK_DIR/ref/netstandard.dll $@

### 2020-09-04 tools/run-perf-docker-csharp.sh

@@ -1,14 +1,0 @@
-#! /bin/bash
-
-# NOTE: pass the sampling rate (Hz) as first argument, followed by parameters to the benchmarker (i.e. ./tools/run-perf-docker.sh 1000 --benchmark=random-inserts)
-# 1000Hz is a good choice for longer experiments (a few minutes), higher rates are recommended for shorter experiments
-# if you don't see enough detail, try and increase the sampling rate
-
-echo "==== capturing at frequency $1 ===="
-
-echo "==== making the cs bundle ===="
-make build/Bundle.cs
-
-echo "==== running docker ===="
-docker run --rm --cap-add SYS_ADMIN -it -v `pwd`:/veribetrfs utaal/dotnet-core-sdk-make-perf /bin/bash /veribetrfs/tools/run-perf-docker-internal-csharp.sh $@
-

### 2020-09-04 tools/build-exe.sh
resolve tools/renaminator.py conflict

@@ -1,4 +1,0 @@
-#!/bin/bash
-
-dafny Mkfs.dfy /spillTargetCode:3 /noVerify
-csc Mkfs.cs Framework.cs /r:System.Numerics.dll /debug /nowarn:0164 /nowarn:0219 /nowarn:1717 /nowarn:0162 /nowarn:0168

### 2020-09-04 tools/plot/perf-compare.py

@@ -1,52 +1,0 @@
-#!/usr/bin/env python3
-import matplotlib
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-import operator
-import bisect
-from parser import Experiment
-from PlotHelper import *
-from TimeSeries import *
-
-output_filename = "compare.png"
-
-def plot_perf_compare(experiments):
-    plotHelper = PlotHelper(12, scale=2)
-
-    try: plotThroughput(plotHelper.nextAxis(depth=2), experiments)
-    except: raise
-    
-    try: plotGrandUnifiedMemory(plotHelper.nextAxis(depth=2), experiments)
-    except: raise
-
-    try: plotSlowIos(plotHelper.nextAxis(depth=2), experiments)
-    except: raise
-
-    try: plotCpuTime(plotHelper.nextAxis(depth=2), experiments)
-    except: raise
-
-    try: plotProcIoBytes(plotHelper.nextAxis(depth=2), experiments)
-    except: raise
-
-#    try: plotIoLatencyCdf(plotHelper.nextAxis(depth=2), experiments)
-#    except: raise
-
-#    try: plotRocksIo(plotHelper.nextAxis(depth=2), experiments)
-#    except: raise
-
-    plotHelper.save(output_filename)
-
-experiments = []
-for arg in sys.argv[1:]:
-    nick,fn = arg.split("=")
-    if nick=="output":
-        output_filename = fn
-    else:
-        try:
-            exp = Experiment(fn, nick)
-            #exp.sortedOpns = exp.sortedOpns[:-5]    # hack: truncate teardown tail of completed exp where memory all goes to 0
-            experiments.append(exp)
-        except (ValueError,FileNotFoundError):
-            print("Can't parse %s; skipping" % nick)
-plot_perf_compare(experiments)

### 2020-09-04 tools/profile-reader.py

@@ -1,64 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-
-class ParsedQILog:
-    def __init__(self, logfilename):
-        self.map = {}
-        for line in open(logfilename).readlines():
-            fields = line.split()
-            if len(fields) < 6 or fields[0] != "Prover":
-                continue
-            loc = fields[3]
-            count = int(fields[5])
-            if loc not in self.map:
-                self.map[loc] = 0
-            if count > self.map[loc]:
-                self.map[loc] = count
-
-    def display(self):
-        pairs = [(count, loc) for loc,count in self.map.items()]
-        pairs.sort()
-        for count, loc in pairs:
-            print("%8d %-1s" % (count, loc))
-
-class CompareRecord:
-    def __init__(self, loc, small_value, big_value):
-        self.loc = loc
-        self.small_value = small_value
-        self.big_value = big_value
-        self.ratio = big_value / float(small_value) if small_value != 0 else 99
-
-    def __repr__(self):
-        return ("%6.1f %8d %8d %-1s" % (self.ratio, self.small_value, self.big_value, self.loc))
-
-class Comparison:
-    def __init__(self, small_filename, big_filename):
-        self.small_parsed = ParsedQILog(small_filename)
-        self.big_parsed = ParsedQILog(big_filename)
-
-        self.recs = []
-        locs = set(self.small_parsed.map.keys()).union(set(self.big_parsed.map.keys()))
-        for loc in locs:
-            small_value = self.small_parsed.map.get(loc, 1)
-            big_value = self.big_parsed.map.get(loc, 1)
-            self.recs.append(CompareRecord(loc, small_value, big_value))
-
-    def display(self):
-        recs = self.recs
-        recs.sort(key = lambda rec: rec.big_value)
-        #recs.sort(key = lambda rec: rec.ratio)
-        for rec in recs:
-            print(rec)
-
-def main():
-    args = sys.argv[1:]
-    if len(args) == 1:
-        ParsedQILog(*args).display()
-    elif len(args) == 2:
-        Comparison(*args).display()
-    else:
-        sys.stderr.write("Usage: %s <qi.log> [qi2.log]\n")
-        sys.stderr.write("With two arguments: sorts references by ratio of counts in second file. Use with different timeouts to see which qis grow most.\n")
-
-main()

### 2020-09-04 tools/renaminator.py

@@ -1,100 +1,0 @@
-#!/usr/bin/env python3
-# Automation for moving dfy files among directories, cleaning up include references.
-
-import os
-import subprocess
-
-EXCLUDED_DIRS = set([".dafny", "build"])
-
-class Renaminator:
-    def __init__(self):
-        self.catalog()
-        # Apply the include path fixes first, since they're expressed relative
-        # to the source path locations.
-        self.mkdirCmds = []
-        self.fixCmds = []
-        self.gitAddCmds = []    # some of these will be pre-move referrers, so do them first
-        self.gitCmds = []
-
-    def catalog(self):
-        paths = []
-        for root, dirs, files in os.walk("."):
-            parts = root.split("/")
-            if len(parts) == 1:
-                # must be "."
-                top = parts[0]
-            else:
-                top = parts[1]
-            if top in EXCLUDED_DIRS or (top[0]=='.' and len(top)>1):
-                continue
-            for file in files:
-                if not file.endswith(".dfy"):
-                    continue
-                fullpath = os.path.join(root, file)
-                paths.append(fullpath)
-        self.paths = paths
-
-    def findSourceDir(self, filename):
-        matchingSourcePaths = [path for path in self.paths if path.endswith("/"+filename)]
-        if len(matchingSourcePaths) == 0:
-            raise Exception("No path matches %s" % filename)
-        if len(matchingSourcePaths) > 1:
-            raise Exception("Multiple paths match %s: %s" % (filename, matchingSourcePaths))
-        path = matchingSourcePaths[0]
-        return path[:-(len(filename)+1)]
-
-    def containsLine(self, filepath, testString):
-        contents = open(filepath).read()
-        return testString in contents
-
-    def fixReferrer(self, referrer, sourceDir, sourceFilename, destDir, destFilename):
-        referrerPath = os.path.split(referrer)[0]
-        sourceRelative = os.path.relpath(os.path.join(sourceDir, sourceFilename), referrerPath)
-        destRelative = os.path.relpath(os.path.join(destDir, destFilename), referrerPath)
-        expectInclude = 'include "%s"' % sourceRelative
-        newInclude = 'include "%s"' % destRelative
-        if self.containsLine(referrer, expectInclude):
[…]

### 2020-09-04 tools/run-perf-docker-internal-csharp.sh

@@ -1,50 +1,0 @@
-#! /bin/bash
-
-frequency=$1
-shift
-
-set -x
-set -e
-
-apt-get update
-apt-get install -y tmux locales
-
-echo "LC_ALL=en_US.UTF-8" >> /etc/environment
-echo "en_US.UTF-8 UTF-8" >> /etc/locale.gen
-echo "LANG=en_US.UTF-8" > /etc/locale.conf
-locale-gen en_US.UTF-8
-
-cd /veribetrfs
-rm -R .veribetrfs-storage || true
-mkdir .veribetrfs-storage
-
-set +e
-rm build/disk-betree/Bundle.i.roslyn.exe
-set -e
-
-cd disk-betree
-make exe-roslyn
-cd ..
-
-echo "==== starting benchmark ===="
-echo "flags: $@"
-
-COMPlus_PerfMapEnabled=1 dotnet build/roslyn-veribetrfs.exe $@ &
-PID=$!
-read -n 1 -s -r -p "Press any key to start recording with perf (preferably before veribetrfs exits)"
-perf record -p $PID -g -F$frequency &
-PERF_PID=$!
-sleep 1
-read -n 1 -s -r -p "Press any key to stop recording"
-
-set +e
-kill $PID
-sleep 2
-
-echo "==== starting benchmark ===="
-
-tmux \
-  new-session  "perf report -g graph,0.1,callee,function,percent" \; \
-  split-window "echo \"The upper pane is perf with callee trees, this is with caller trees. In tmux, use C-b Up and C-b Down to switch panes. If perf is crashing, try with a shorter trace.\"; read -n 1 -s -r -p \"Press any key to start loading this pane\"; perf report -g graph,0.4,caller,function,percent" \; \
-  attach \;
-

### 2020-09-04 tools/run-perf-docker-internal.sh

@@ -1,49 +1,0 @@
-#! /bin/bash
-
-frequency=$1
-shift
-
-set -x
-set -e
-
-apt-get update
-apt-get install -y tmux locales
-
-echo "LC_ALL=en_US.UTF-8" >> /etc/environment
-echo "en_US.UTF-8 UTF-8" >> /etc/locale.gen
-echo "LANG=en_US.UTF-8" > /etc/locale.conf
-locale-gen en_US.UTF-8
-
-cd /veribetrfs
-rm -R .veribetrfs-storage || true
-mkdir .veribetrfs-storage
-
-set +e
-rm build/*.o
-rm build/framework/*.o
-rm build/Veribetrfs
-set -e
-
-make CC=clang++ CCFLAGS="-g -fpermissive" build/Veribetrfs
-
-echo "==== starting benchmark ===="
-echo "flags: $@"
-
-./build/Veribetrfs $@ &
-PID=$!
-perf record -p $PID -g -F$frequency &
-PERF_PID=$!
-sleep 1
-read -n 1 -s -r -p "Press any key to stop recording"
-
-set +e
-kill $PID
-sleep 2
-
-echo "==== starting benchmark ===="
-
-tmux \
-  new-session  "perf report -g graph,0.1,callee,function,percent" \; \
-  split-window "echo \"The upper pane is perf with callee trees, this is with caller trees. In tmux, use C-b Up and C-b Down to switch panes. If perf is crashing, try with a shorter trace.\"; read -n 1 -s -r -p \"Press any key to start loading this pane\"; perf report -g graph,0.4,caller,function,percent" \; \
-  attach \;
-

### 2020-09-04 tools/kube/start-docker.sh
delete kube tools, clean up paper.tex

@@ -1,1 +1,0 @@
-docker container run -t -i --mount src="`pwd`",target=/home/dafnyserver/tutorial,type=bind --workdir /home/dafnyserver/tutorial jonhdotnet/veribetrfs:1.0 /bin/bash

### 2020-09-04 tools/line_counter_report_lib.py
move data/ files to build/

@@ -3,11 +3,6 @@ import lib_deps
 import json
 import os
 
-# make build/Impl/Bundle.i.lcreport -j4
-# (incremental report redraw:)
-# tools/line_counter.py --mode report --input Impl/Bundle.i.dfy --output build/Impl/Bundle.i.lcreport
-# cp build/Impl/Bundle.i.lcreport ../veripapers/osdi2020/data/line-counts.tex
-
 def loadReport(iref):
     fp = open(lib_deps.targetName(iref, ".lc"))
     values = json.load(fp)

### 2020-09-04 tools/linear_line_counts.py

@@ -53,7 +53,7 @@ packages = [
             ["lib/DataStructures/LinearMutableMap.i.dfy"]),
     ]
 
-DATA_FILE = "data/linear_lines.json"
+DATA_FILE = "build/linear_lines.json"
 def collect():
     accum = []
     for package in packages:

### 2020-09-04 tools/automation-study.py
move data/ files to build/

@@ -82,7 +82,7 @@ def opaqueInstances():
         #sys.stdout.flush()
     return records
 
-INTERMEDIATE = "data/auto.data"
+INTERMEDIATE = "build/auto.data"
 
 def scrape():
     records = opaqueInstances()
@@ -150,12 +150,6 @@ def totalDefinitions(defs):
         "autoOpaquePct": "%d\\%%" % (100.0*defs["autoOpaqueCount"] / len(ghost)),
     })
     return defs
-
-# def emit_constants(defs):
-#     with open("../veripapers/osdi2020/data/automation-constants.tex", "w") as fp:
-#         fp.write("% generated by tools/automation-study.py\n")
-#         for k,v in defs.items():
-#             fp.write("\\newcommand{\\%s}{%s}\n" % (k, v))
 
 def find_dead_lemmas():
     lemmaRE = r'\blemma\b(\s|{[^}]+})+(?P<X>\w+)'

### 2020-09-04 tools/linear_line_counts.py

@@ -54,7 +54,7 @@ packages = [
              "lib/DataStructures/MutableMapImpl.i.dfy"]),
     ]
 
-DATA_FILE = "data/linear_lines.json"
+DATA_FILE = "build/linear_lines.json"
 def collect():
     accum = []
     for package in packages:

### 2020-09-04 tools/line_count_lib.py
move some tmp files to build

@@ -25,7 +25,7 @@ class DafnyFile:
     return self.filename.endswith(".s.dfy")
 
 # We'll leave copies of counted files here so we can inspect the counting filters for sanity.
-INSPECT_DIR = "./tmp/inspect"
+INSPECT_DIR = "./build/inspect"
 
 class Counter:
     def __init__(self, iron_base):
@@ -177,7 +177,7 @@ class Counter:
       return int(sloc)
 
     def make_tmp_dir(self):
-      tmp_dir = self.iron_base + "/tmp/linecounts_%x" % random.randint(0, 1<<31)
+      tmp_dir = self.iron_base + "/build/linecounts_%x" % random.randint(0, 1<<31)
       if not os.path.exists(tmp_dir):
         os.makedirs(tmp_dir)
       #print("tmp dir: %s" % tmp_dir)

### 2021-05-11 concurrency/docs/theory/theory.tex
move directory

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{dashbox}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{stackengine}
\usepackage{stmaryrd}
\usepackage{wasysym}
\usepackage[]{stix}
\usepackage{graphicx}

\usepackage{parskip}
\usepackage[tmargin=1in,bmargin=1in,lmargin=0.75in,rmargin=0.75in]{geometry}

\usepackage{proof}

%\usepackage[most]{tcolorbox}

%\newtcbox{\dashedbox}[1][]{
%  math upper,
%  baseline=0.4\baselineskip,
%  equal height group=dashedbox,
%  nobeforeafter,
%  colback=white,
%  boxrule=0pt,
%  enhanced jigsaw,
%  boxsep=0pt,
%  top=2pt,
%  bottom=2pt,
%  left=2pt,
%  right=2pt,
%  borderline horizontal={0.5pt}{0pt}{dashed},
%  borderline vertical={0.5pt}{0pt}{dashed},
%  drop lifted shadow,
%  #1
%}

\newcommand{\einfer}[3]
        {\infer[(\textsc{#1})]{#3}{#2}}
\newcommand{\stacktwo}[2]{\stackanchor{\ensuremath{#1}}{\ensuremath{#2}}}
\newcommand{\stackthree}[3]{\stackanchor{\stacktwo{#1}{#2}}{\ensuremath{#3}}}
\newcommand{\stackfour}[4]{\stackanchor{\stackthree{#1}{#2}{#3}}{\ensuremath{#4}}}
\newcommand{\stackfive}[5]{\stackanchor{\stackfour{#1}{#2}{#3}{#4}}{\ensuremath{#5}}}
\newcommand{\stacksix}[6]{\stackanchor{\stackfive{#1}{#2}{#3}{#4}{#5}}{\ensuremath{#6}}}
\newcommand{\stackseven}[7]{\stackanchor{\stacksix{#1}{#2}{#3}{#4}{#5}{#6}}{\ensuremath{#7}}}

\newcommand\dboxed[1]{\dbox{\vphantom{t}\ensuremath{#1}}}

\newcommand{\hoare}[3]{\ensuremath{\{ #1 \}~ #2 ~\{ #3 \}}}
\newcommand{\viewshift}{\Rrightarrow}
\newcommand{\unit}{\epsilon}
\newcommand{\true}{\textsf{True}}
\newcommand{\false}{\textsf{False}}
\newcommand{\bool}{\textsf{bool}}
\DeclareMathOperator{\rel}{rel}
\DeclareMathOperator{\PCM}{PCM}
\DeclareMathOperator{\MonoidBase}{MonoidBase}
\DeclareMathOperator{\MonoidExt}{MonoidExt}
\DeclareMathOperator{\Frac}{Frac}
\newcommand{\monoid}[1]{\textsf{monoid}~#1}
\renewcommand{\eqdef}{\triangleq}
\newcommand{\borrow}[2]{\textsc{\&}^{#1}#2}
\newcommand{\pointsto}{\hookrightarrow}

%\DeclareFontEncoding{LS1}{}{}
%\DeclareFontSubstitution{LS1}{stix}{m}{n}
%\DeclareSymbolFont{arrows2}       {LS1}{stixsf}   {m}{it}
%\SetSymbolFont{arrows2}     {bold}{LS1}{stixsf}   {b}{it}
%\DeclareSymbolFontAlphabet{\mathsfit}{arrows2}
%\DeclareMathSymbol{\rightarrowgtr}{\mathrel}{arrows2}{"5F}

\newcommand{\refines}{\rightarrowgtr}

\newcommand{\ltIntersect}{\sqcup}
\newcommand{\ltLe}{\sqsubseteq}
\newcommand{\ltGe}{\sqsupseteq}
\DeclareMathOperator{\lifetime}{lifetime}
\DeclareMathOperator{\borrowed}{borrowed}

\let\active\undefined
\DeclareMathOperator{\active}{active}

%\newcommand{\shardProduct}{\rtimes}
%\newcommand{\shardProduct}{\rightmoon}
%\newcommand{\shardProduct}{\rsub}
\newcommand{\shardProduct}{\ogreaterthan}

\begin{document}

Here, we describe a PCM-based program logic for reasoning about concurrent programs,
with a built-in notion of ``PCM refinement'' to provide modularity that is both useful in its own right
and particularly useful for reasoning about ``read-only'' state.

\textbf{Monoid definitions}

%We write $\widehat{M} = (M, \cdot, \mapsto)$ is a Partial Commuative Monoid (PCM) if,
%where $\cdot : M \times M \to M^?, \text{~~comm., assoc.}$
%and $\mapsto : M \times M \to \bool, \text{~~trans., refl., monotonic}$,
%we have
%\begin{align*}
%& a \cdot b = b \cdot a & & \text{(commutativity)}\\
%& (a \cdot b) \cdot c = a \cdot (b \cdot c) & & \text{(associativity)}\\
%& (a \mapsto b) \land (b \mapsto c) \implies a \mapsto c && \text{(transitivity)}\\
%& (a \mapsto a) && \text{(reflexitivity)}\\
%& (a \mapsto b \land a \cdot c \ne \bot \implies b \cdot c \ne \bot \land a \cdot c \mapsto b %\cdot c) && \text{(monotonic)}
%\end{align*}

%\\
%Monoid $\mathcal{M} := M ~|~ \mathcal{M}\otimes\mathcal{M} ~|~ E[\mathcal{M}]$\\
%\\
%\textbf{Base monoid} $M$ equip with $\cdot$ commutative, associative,
%unit $\unit \in M$, bot element $\bot \in M$, and
%$a \mapsto_M b \eqdef (\forall c .~ a\cdot c \ne \bot \implies b\cdot c \ne \bot)$\\
%\\
%\textbf{Product monoid} $M \otimes N$ where $(a\otimes b)\cdot(c \otimes d) = (a\cdot c)\otimes(b\cdot d)$
%and $(a\otimes b) \mapsto (a' \otimes b') \eqdef (a \mapsto a') \land (b \mapsto b')$\\
%\\
%\textbf{Extension monoid} $E[M]$. Monoid $E,\cdot$, predicate $\rel : E \times M \to \bool$,
%and $b \mapsto b' \eqdef
%(\forall p, q .~ \rel(b \cdot p, q) \implies \exists b' .~ \rel(b' \cdot p, q') \land (q \mapsto q'))$
%and
%$\forall b, a, a' .~ \rel(b,a) \land \rel(b,a') \implies a \mapsto a'$\\
%\\
%\\
%Define $\le$ by $a \le b \eqdef \exists b . a \cdot c = b$.\\
%\\

We start by defining what it means to be a PCM with a transition relation $\mapsto$.

Let $\widehat{M} = (M, \cdot, \mapsto)$ where
where $\cdot : M \times M \to M^?$
and $\mapsto : M \times M \to \bool$.
(We use $M^? \eqdef M \cup \{\bot\}$.)

We say that $\mapsto$ is \emph{monotonic} if
\[ \forall a, b, c.~ (a \mapsto b) \land (a \cdot c \ne \bot) \implies (b \cdot c \ne \bot) \land (a \cdot c \mapsto b \cdot c) \]
This allows us to define what it means to be a PCM,
\[
  \einfer{PCM}{
    \stackthree{
      \widehat{M} = (M, \cdot, \mapsto)
    }{
      \cdot : M \times M \to M^?, \text{~~comm., assoc.}
    }{
      \mapsto : M \times M \to \bool, \text{~~trans., refl., monotonic}
    }
  }{
    \PCM(\widehat{M})
  }
\]
Note that we can always construct a PCM by picking $\cdot$ to be commutative and associative,
and then by definition setting $\mapsto$ to be,
\[
  a \mapsto b = \forall c .~ a \cdot c \ne \bot \implies b \cdot c \ne \bot
\]
However, we can also set $\mapsto$ to be more restrictive, which will be useful later.

\textbf{PCM Example}

Consider a \emph{Sharded State Machine} (SSM). (This is similar to the STS common in the literature but framed a little differently.)

An SSM is given by \textbf{(i)} some PCM state $S, \cdot$, \textbf{(ii)} an invariant $inv : S \to \bool$ and a sharded transition $\rightsquigarrow : S \times S$. For a sharded state
machine, the invariant inductivity predicate is given by,
\[
  inv(a \cdot c) \land (a \rightsquigarrow b) \implies inv(b \cdot c)
\]
(Note: $\rightsquigarrow$ is neither transitive nor required to be monotonic here. However, an alternative definition would require $\rightsquigarrow$ to be monotonic, in which case we could
simplify the inductivity predicate to $inv(a) \land (a \rightsquigarrow b) \implies inv(b)$.)

We can define a PCM to correspond to this SSM by,
\[M \eqdef \{m \in S : \exists c .~ inv(m \cdot c) \}\]
\[a \cdot_M b \eqdef 
  \begin{cases}
      a \cdot b & \text{if } a \cdot b \in M\\
      \bot,              & \text{otherwise}
  \end{cases}\]
and finally we can define $\mapsto$ as the transitive closure of $\mapsto_\text{pre}$ defined as,
\[
a \mapsto_{\text{pre}} b \eqdef
  \exists a_0, b_0, c .~ (a = a_0 \cdot c) \land (b = b_0 \cdot c) \land a_0 \rightsquigarrow b_0
\]

\textbf{PCM refinements (motivation)}

There are a handful of reasons to want to modularize PCM definitions:
\begin{itemize}
  \item The ``base language'' may provide some ``points to'' assertion $\ell \pointsto v$,
    and then the user wants to build something else on top of that (like a fractional permission logic or similar) without needing to inject it into the native definition.
  \item You have some PCM $N$ (custom built for your application), and you want to build a reader-writer lock to manage
    access to the state of $N$, without encoding the reader-writer lock protocol into $N$.
  \item You have some PCM $N$, and you want to manage part of it via some hash table implementation. You want to verify the hash table implementation independently (e.g., via some SSM strategy) and then use it modularly to manage some part of the $N$-state.
  \item You have some PCM $N$ with a counter field, and you want to write the counter-manipulation code as its own independent module.
\end{itemize}
These are all things that have come up in VeriBetrKV.

We'll develop the concept of a PCM refinement, denoted $\widehat{M} \refines \widehat{N}$.
In a modular design, $\widehat{M}$ will be the ``sub-component'', while $\widehat{N}$
is the client.

For example, a hypothetical cache architecture might have a ``main'' ghost state \textsf{Cache}
represented as an SSM.
We might have a sub-component $\textsf{RwLock} \refines \textsf{Cache}$,
another sub-component $\textsf{HashTable} \refines \textsf{Cache}$,
and even further a sub-component $\textsf{Counter} \refines \textsf{HashTable}$.

\begin{center}
\includegraphics[scale=0.3]{hypothetical-cache-architecture.png}
\end{center}

\textbf{PCM refinements}

We'll use the symbol $\refines$ to denote refinement. Yeah, you can tell I scoured the \LaTeX~ documentation for this thing, but I actually like it: it suggests that the thing on the left is more ``broken up'' (e.g., like in a fractional logic).

We use $\widehat{M} \refines_{\rel} \widehat{N}$ to denote that $\widehat{M}$ \emph{refines} $\widehat{N}$ via a relation $\rel : M \times N \to bool$. Of course, it must satisfy some conditions 
(and this is why we allow for flexibility in the definition of $\mapsto$ above).

\[
  \einfer{PCM-Refinement}{
    \stackfive{
      \PCM(\widehat{M}) \quad \PCM(\widehat{N})
    }{
      \rel : M \times N \to \bool
    }{
      \rel(\unit, \unit)
    }{
       (\forall b, b', q .~ b \mapsto b' \land \rel(b, q) \implies \exists q' .~ \rel(b', q') \land (q \mapsto q'))
    }{
      \forall b, a, a' .~ \rel(b,a) \land \rel(b,a') \implies a \mapsto a'
    }
  }{
    \widehat{M} \refines_{\rel} \widehat{N}
  }
\]

\textbf{Using ghost state in programs}

We instantiate instances of monoidal ghost state at \emph{locations}.
The notation $\dboxed{a}^\gamma$ means that state $a$ is at location $\gamma$.
Typically, locations are considered arbitrary values; here, we have to do a little
more to handle the relationships between different ghost states.

For example, suppose we have a refinement $\widehat{M} \refines_{\rel} \widehat{N}$.
Perhaps $\widehat{M}$ represents a reader-writer protocol or something similar.
Furthermore, let's say we're tracking ghost state $n \in N$ at location $\gamma$:
we will use the notation $\dboxed{n}^\gamma$.

Now, we want to use the $\widehat{M}$ protocol to help manage that state.
We'll need some rules for ``exchanging'' the $\widehat{M}$-state with $\widehat{N}$-state.
Furthermore, when we manage the $\widehat{M}$-state, we'll need to somehow ``remember''
that it corresponds to the $\widehat{N}$-state at $\gamma$, via the refinement $\rel$.
We will track this information in the location value.

Thus, we'll define the location values recursively, reusing the $\refines$ notation:

$\gamma ~~:=~~ \alpha ~~|~~ \alpha \refines_{\rel} \gamma$.

Here, $\alpha$ denotes any ``base location.'' We will define rules that let us
convert $\dboxed{n}^\gamma$ state to and from $\dboxed{m}^{\alpha \refines_{\rel} \gamma}$
state.

Here are our rules for manipulating PCM ghost state, which are mostly standard except
the ones involving $\refines$.

\[
  \true \implies \dboxed{\unit}^\gamma
\]

\[
  \dboxed{\bot}^\gamma \implies \false
\]

\[
  \dboxed{t\cdot u}^\gamma\iff\dboxed{t}^\gamma * \dboxed{u}^\gamma
\]

\[
  \einfer{Ghost-Alloc}{
    \PCM(\widehat{M}) \quad a\in M
  }{
    \true \viewshift \exists\alpha.~ \dboxed{a}^\alpha
  }
\]

\[
  \einfer{Ghost-Refinement-Alloc}{
    \stacktwo{
      \widehat{M} \refines_{\rel} \widehat{N}
    }{
      \rel(m, n) \quad b \ne \bot
    }
  }{
    \dboxed{n}^\gamma \viewshift \exists \alpha.~ \dboxed{m}^{\alpha \refines_{\rel} \gamma}
  }
\]

\[
  \einfer{Ghost-ViewShift}{
    a \mapsto b
  }{
    \dboxed{a}^\gamma \viewshift \dboxed{b}^\gamma
  }
\]

\[
  \einfer{Ghost-Refinement-Exchange}{
    \stacktwo{
      \widehat{M} \refines_{\rel} \widehat{N}
    }{
      \forall p,q.~ \rel(b\cdot p, q) \implies \exists q' .~ \rel(b' \cdot p, q') \land (a \cdot q \mapsto a' \cdot q')
    }
  }{
    \dboxed{b}^{\alpha \refines_{\rel} \gamma} * \dboxed{a}^\gamma
      \viewshift
    \dboxed{b'}^{\alpha \refines_{\rel} \gamma} * \dboxed{a'}^\gamma
  }
\]

\textbf{Justification.}

To justify these rules, we can roll up all the $\dboxed{m}^\gamma$ values into a mega-monoid (or resource algebra)
one defined as a mapping from locations to monoid elements.
Then we show that all those rules above follow from frame-preserving updates.

\textbf{Notes about abstraction.}

When using a refinement
$\widehat{M} \refines \widehat{N}$
to abstract out some subcomponent, we observe that $\widehat{M}$ (which corresponds to the subcomponent) needs to be defined with respect to the base state $\widehat{N}$,
which initially seems a little strange. Shouldn't we define the subcomponent \emph{first}, and then build the larger components out of those?

And yes, if for example $\widehat{M}$ is a reader-writer lock, then $\widehat{M}$ and the refinement $\rel$ will need to be defined in terms of $\widehat{N}$. However, this is not really so different than using a templated type, e.g., writing \texttt{RwLock<T>} to create a mutex storing type \texttt{T}.

\textbf{Borrows and read-only state.}

I think we can use a fairly simple system for representing (shared) borrows. 
We use $\kappa$ to represent a lifetime variable.
$\borrow{\kappa}{A}$ represents some borrowed proposition;
$\active(\kappa)$ indicates that lifetime $\kappa$ is active,
and $\borrowed(\kappa, A)$ indicates you can get back exclusive access to $A$
once the lifetime $\kappa$ expires.

We use $\kappa_1 \ltIntersect \kappa_2$ to represent lifetime intersection
and $\kappa_1 \ltLe \kappa_2$ to represent lifetime inclusion.

Of course, the intent is that the verification framework will use some sort of borrow-checker
to identify the lifetimes and perform these automatically.

(Note: RustBelt uses a much more complex system - need to understand why; am I missing something or is it for some feature I don't need?)

\[
  \einfer{Borrow-begin}{
  }{
    A \viewshift \exists
        \kappa .~ \active(\kappa) * \borrowed(\kappa, A) * \borrow{\kappa}{A}
  }
\]

\[
  \einfer{Borrow-end}{
  }{
    \active(\kappa) * \borrowed(\kappa, A) \viewshift A
  }
\]

\[
  \einfer{Borrow-dupe}{
  }{
    \borrow{\kappa}{A} \viewshift \borrow{\kappa}{A} * \borrow{\kappa}{A}
  }
\]

\[
  \einfer{Reborrow}{
    \kappa_1 \ltGe \kappa_2
  }{
    \borrow{\kappa_1}{A} \viewshift \borrow{\kappa_2}{A}
  }
\]

\[
  \active(\kappa_1 \ltIntersect \kappa_2) \iff \active(\kappa_1) * \active(\kappa_2)
\]

Next, let's add some rules for making use of borrowed PCM state.

\[
  \true \implies \borrow{\kappa}{\dboxed{\unit}^\gamma}
\]

\[
  \active(\kappa) * \borrow{\kappa}{\dboxed{\bot}^\gamma} \implies \false
\]

\[
  \einfer{Borrow-Combine}{
    \forall r .~ m \le r \land n \le r \implies k \le r
  }{
    \borrow{\kappa}{\dboxed{m}^\gamma} * \borrow{\kappa}{\dboxed{n}^\gamma} \viewshift \borrow{\kappa}{\dboxed{k}^\gamma}
  }
\]

\[
  \einfer{Borrow-Ghost-ViewShift}{
    c \cdot a \mapsto c \cdot b
  }{
    \active(\kappa) * \borrow{\kappa}{\dboxed{c}^\gamma} * \dboxed{a}^\gamma
    \viewshift
    \active(\kappa) * \dboxed{b}^\gamma
  }
\]

Finally, rules to allow borrows to interact with PCM refinements:

\[
  \einfer{Borrow-Back}{
    \stacktwo{
      \widehat{M} \refines_{\rel} \widehat{N}
    }{
      \forall p, b .~ \rel(m \cdot p, b) \implies a \le b
    }
  }{
    \borrow{\kappa}{\dboxed{m}^{\delta \refines \gamma}} \viewshift \borrow{\kappa}{\dboxed{a}^\gamma}
  }
\]

\[
  \einfer{Borrow-Ghost-Refinement-Exchange}{
    \stacktwo{
      \widehat{M} \refines_{\rel} \widehat{N}
    }{
      \forall p,q.~ \rel(c\cdot b\cdot p, q) \implies \exists q' .~ \rel(c \cdot b' \cdot p, q') \land (a \cdot q \mapsto a' \cdot q')
    }
  }{
    \active(\kappa) *
    \borrow{\kappa}{\dboxed{c}^{\delta\refines\gamma}} *
    \dboxed{b}^{\delta\refines\gamma} * \dboxed{a}^\gamma
      \viewshift
    \active(\kappa) *
    \dboxed{b'}^{\delta\refines\gamma} * \dboxed{a'}^\gamma
  }
\]

Note that Borrow-Ghost-Refinement-Exchange allows us to utilize a borrow in the extension world $\delta \refines \gamma$ but \emph{not} a borrow in the base world $\gamma$. This is important as we would have an unsoundness otherwise. In general, we can't have any rule that allows a borrow in the $\gamma$ world to flow back to the $\delta \refines \gamma$ world, as then that borrow could ``interfere''

\textbf{Justification}

As before, our justification will be to roll all the ghost state, the $\borrow{\kappa}{\dboxed{a}^\gamma}$ tokens, the $\active$ tokens, and the $\borrowed$
tokens into a single mega-monoid, and derive the above rules as frame-preserving updates.

There's one trick worth pointing out. Intuitively, we're going to have some invariant
that says the following:

Suppose we have $\active(\kappa) * \borrow{\kappa}{\dboxed{a}^\gamma}$ with $a \in M$. Then, consider:
\begin{itemize}
  \item All ``reserved'' $\gamma$-state: $\borrowed(\kappa', \dboxed{b}^\gamma)$
  \item All state in an extension $\delta$ of $\gamma$, both \emph{borrowed and non-borrowed}.
        By ``extension'' $\delta$, we mean any location $\alpha \refines \gamma$,
        $\alpha \refines (\beta \refines \gamma)$, and so on, (and \emph{not} including $\gamma$ itself).
        Project all this state down to $M$ via the refinement relations.
\end{itemize}
\emph{Borrow state invariant}: Then the monoidal sum of all this state gives some value $t$ where $a \preceq t$.

Essentially, this says that for active borrow state $\borrow{\kappa}{\dboxed{a}^\gamma}$,
that state is actually ``backed'' by some state which is either ``borrowed out'' or
in some ``refinement world.''

Note again in the second bullet that within the ``refinement worlds'' we allow both
\emph{borrowed and non-borrowed} state. This is the ``trick'' that lets the
PCM refinement be so effective for dealing with borrowed state.

For example, suppose we derive the fractional logic
$\Frac(\widehat{N}) \refines \widehat{N}$.

Now, as a user of the fractional logic, we can derive $a \viewshift \Frac(1, a)$,
and then $\Frac(1, a) \viewshift \Frac(1/2, a) * \Frac(1/2, a)$.
Then, we could perform a borrow of one of the halves and obtain
$\borrow{\kappa}{\Frac(1/2, a)}$,
then obtain $\borrow{\kappa}{a}$.

We are able to do this, even though we have ``only'' borrowed half, e.g.,
we have $\borrowed(\kappa, \Frac(1/2, a)) * \Frac(1/2, a)$, rather than
$\borrowed(\kappa, \Frac(1, a))$.

\end{document}

### 2021-05-11 concurrency/docs/theory/theory.tex
fix typos

@@
 
-\newcommand{\ltIntersect}{\sqcup}
+\newcommand{\ltIntersect}{\sqcap}
 \newcommand{\ltLe}{\sqsubseteq}
@@
 
-An SSM is given by \textbf{(i)} some PCM state $S, \cdot$, \textbf{(ii)} an invariant $inv : S \to \bool$ and a sharded transition $\rightsquigarrow : S \times S$. For a sharded state
+An SSM is given by \textbf{(i)} some PCM state $S, \cdot$, \textbf{(ii)} an invariant $inv : S \to \bool$ and a sharded transition $\rightsquigarrow : S \times S \to \bool$. For a sharded state
 machine, the invariant inductivity predicate is given by,

### 2021-05-11 concurrency/docs/theory/theory.tex
fix another typo

@@
 This is important as we would have an unsoundness otherwise.
-In general, we can't have any rule that allows a borrow in the $\gamma$ world to flow back to the $\delta \refines \gamma$ world, as then that borrow could ``interfere''
+In general, we can't have any rule that allows a borrow in the $\gamma$ world to flow back to the $\delta \refines \gamma$ world, as then that borrow could ``interfere'' with the $\delta[\gamma]$-world's state which is being held ``in reserve'' for that borrow.
 

### 2021-05-11 concurrency/docs/theory/theory.tex
typo

--- 2021-05-11 concurrency/docs/theory/theory.tex
@@
 This is important as we would have an unsoundness otherwise.
-In general, we can't have any rule that allows a borrow in the $\gamma$ world to flow back to the $\delta \refines \gamma$ world, as then that borrow could ``interfere''
+In general, we can't have any rule that allows a borrow in the $\gamma$ world to flow back to the $\delta \refines \gamma$ world, as then that borrow could ``interfere'' with the $\delta\refines\gamma$-world's state which is being held ``in reserve'' for that borrow.
 

### 2021-05-19 tools/checkout-dafny-commit.sh
update dafny to latest betr-concurrent-merge

@@ -19,7 +19,7 @@ set -e
 set -x
 
 # set to betr-concurrent-merge branch
-commit=76f6c6f97b091ad4d2db587945fca8d4d828e48d
+commit=0bec10367bcba81a30457c71bb72292c72dfc4ed
 if [ $1 ]; then
    commit=$1
 fi

### 2021-05-19 tools/checkout-dafny-commit.sh
update dafny commit to latest betr

@@ -23,7 +23,7 @@ set -e
 set -x
 
 # currently set to 'betr' branch
-commit=46714c9d87fdffe1e7050bb3bc4e304c74f696aa
+commit=2c810aa9cfa1ff413a7eb34730e71b722e6443d7
 if [ $1 ]; then
    commit=$1
 fi

### 2021-05-20 concurrency/docs/theory/theory.tex
fix typo
write a more formal proof of the soundness of the TPCM-refinement borrowing concept

@@
 \DeclareMathOperator{\rel}{rel}
-\DeclareMathOperator{\PCM}{PCM}
+\DeclareMathOperator{\TPCM}{TPCM}
 \DeclareMathOperator{\MonoidBase}{MonoidBase}
@@
 \DeclareMathOperator{\borrowed}{borrowed}
+
+\newcommand\Bigcircle{\raisebox{-0.5mm}{\scalebox{1.7}{$\bigcirc$}}}
 
@@
 We start by defining what it means to be a PCM with a transition relation $\mapsto$.
+We call this a \emph{Transitional PCM} (TPCM).
 
@@
 \[ \forall a, b, c.~ (a \mapsto b) \land (a \cdot c \ne \bot) \implies (b \cdot c \ne \bot) \land (a \cdot c \mapsto b \cdot c) \]
-This allows us to define what it means to be a PCM,
+This allows us to define what it means to be a TPCM,
 \[
-  \einfer{PCM}{
+  \einfer{TPCM}{
     \stackthree{
@@
   }{
-    \PCM(\widehat{M})
+    \TPCM(\widehat{M})
   }
 \]
-Note that we can always construct a PCM by picking $\cdot$ to be commutative and associative,
+Note that we can always construct a TPCM by picking $\cdot$ to be commutative and associative,
 and then by definition setting $\mapsto$ to be,
@@
 
-\textbf{PCM Example}
+\textbf{TPCM Example}
 
@@
 
-We can define a PCM to correspond to this SSM by,
+We can define a TPCM to correspond to this SSM by,
 \[M \eqdef \{m \in S : \exists c .~ inv(m \cdot c) \}\]
@@
 
-\textbf{PCM refinements (motivation)}
+\textbf{TPCM refinements (motivation)}
 
@@
 
-We'll develop the concept of a PCM refinement, denoted $\widehat{M} \refines \widehat{N}$.
+We'll develop the concept of a TPCM refinement, denoted $\widehat{M} \refines \widehat{N}$.
 In a modular design, $\widehat{M}$ will be the ``sub-component'', while $\widehat{N}$
@@
 
-\textbf{PCM refinements}
+\textbf{TPCM refinements}
 
@@
 \[
-  \einfer{PCM-Refinement}{
+  \einfer{TPCM-Refinement}{
     \stackfive{
-      \PCM(\widehat{M}) \quad \PCM(\widehat{N})
+      \TPCM(\widehat{M}) \quad \TPCM(\widehat{N})
     }{
@@
 
-Here are our rules for manipulating PCM ghost state, which are mostly standard except
+Here are our rules for manipulating TPCM ghost state, which are mostly standard except
 the ones involving $\refines$.
@@
   \einfer{Ghost-Alloc}{
-    \PCM(\widehat{M}) \quad a\in M
+    \TPCM(\widehat{M}) \quad a\in M
   }{
@@
     }{
-      \rel(m, n) \quad b \ne \bot
+      \rel(m, n)
     }
@@
 
-Next, let's add some rules for making use of borrowed PCM state.
+Next, let's add some rules for making use of borrowed TPCM state.
 
@@
 
-Finally, rules to allow borrows to interact with PCM refinements:
+Finally, rules to allow borrows to interact with TPCM refinements:
 
@@
 \emph{borrowed and non-borrowed} state. This is the ``trick'' that lets the
-PCM refinement be so effective for dealing with borrowed state.
+TPCM refinement be so effective for dealing with borrowed state.
 
@@
 
+\textbf{Justification (more formally)}
+
+We can define a resource algebra (RA) as follows. Let $\Gamma$ be the set of locations
+$\gamma$ and $K$ the set of lifetimes $\kappa$.
+\begin{itemize}
+  \item \emph{live} state at each location $\gamma$ given by $live : \Gamma \to m$
+  \item \emph{reserved} state at each location $\gamma$ and lifetime $\kappa$,
+      given by $reserved : K \to \Gamma \to m$.
+  \item A set of \emph{borrows} where $borrows \subseteq K \times \Gamma \times M$.
+  \item An active set of lifetimes, given by $active : K$.
+\end{itemize}
+
+The live state is additive (that is, $live(\gamma \to m) \cdot live(\gamma \to p) \eqdef live(\gamma \to p \cdot q)$.
+Reserved state is similarly additive.
+
+Finally, borrows are freely duplicable, and add \emph{freely} (i.e., take the union
+of the set of borrow objects).
+
+We'll describe the invariants on this resource algebra below, but first, I'll explain how
+all the above concepts are embedded into this resource algebra:
+\begin{align*}
+ \dboxed{m}^\gamma &\eqdef live(\gamma \to m)\\
+ \borrowed(\kappa, \dboxed{m}^\gamma) &\eqdef reserved(\kappa,\gamma \to m)\\
+ \borrow{\kappa}{\dboxed{m}^\gamma} &\eqdef borrow(\kappa, \gamma \to m)\\
+ \active(\kappa) &\eqdef active(\kappa)
+\end{align*}
+
+Now we define some functions on this state, defined recursively.
+First, for any refinement location $\delta \refines_{\rel} \gamma$,
+where $\rel : M \times N \to \bool$,
+we let $Project_\delta : M \to N$ be a ``functionalization'' of $\rel$. In particular,
+\[
+  \forall m .~ domain(\rel, m) \implies \rel(m, Project_\delta(m)).
+\]
+where $domain(\rel, m) \eqdef \exists n .~ \rel(m, n)$.
+
+Now we define $Total_\kappa(\gamma)$ to be the ``total'' state of a location $\gamma$:
+that is, all the live and reserved state (for the active livetime $\kappa$)
+\emph{plus} all the total state of locations which depend on $\gamma$. Thus we use
+this recursive definition,
+
+\[
+  Total_\kappa(\gamma) \eqdef live(\gamma) \cdot
+      \left(
+        \Bigcircle_{\kappa' \ltGe \kappa} reserved(\kappa',\gamma)
+      \right) \cdot
+      \left(
+        \Bigcircle_{\delta} Project_\delta\left(Total_\kappa(\delta \refines_{\rel} \gamma)\right)
+      \right)
+\]
+
+Now, this definition is only sensible when 
+$Total_\kappa(\delta \refines_{\rel} \gamma)$ is in the ``domain'' of $\rel$.
+
+Thus we will have an invariant, only applying to the active lifetime $active$,
+\begin{align}
+  domain(rel, Total_{active}(\delta \refines_{\rel} \gamma)) \label{invTotal}
+\end{align}
+
+Next, we define a similar recursive function called $View$, which will be defined
+only in terms of the $reserved$ state. However, it will be used to explain what the ``possible''
+values of the $Total$ (minus the live state). A view will be a function $v : M \to \bool$.
+These views will also be closed \emph{upwards}, i.e., $v(a) \implies v(a \cdot b)$.
+
+Again, we define a ``projection'' function from a view on a refinement TPCM to a base TPCM.
+\[
+  ProjectView_\delta(v) \eqdef \lambda a .~
+      \exists x, y .~ v(x) \land rel(x, y) \land y \le a.
+\]
+For a given $m$, we define $ViewOver(m)$ to be the view of all elements that are at least $m$:
+\[
+  ViewOver(m) \eqdef \lambda n .~ n \ge m
+\]
+And we can add two views in the natural way:
+\[
+  v \otimes w \eqdef \lambda m .~ \exists x, y .~ v(x) \land w(y) \land x \cdot y = m
+\]
+Finally, we're ready to define the $View_\kappa$ function for a given location, again
+recursively, like the $Total_\kappa$ function:
+\[
+  View_\kappa(\gamma) \eqdef 
+      \left(
+        \bigotimes_{\kappa' \ltGe \kappa} ViewOver(reserved(\kappa,\gamma))
+      \right) \otimes
+      \left(
+        \bigotimes_{\delta} ProjectView_\delta\left(View_\kappa(\delta \refines_{\rel} \gamma)\right)
+      \right)
+\]
+Now the $View_\kappa$ denotes the ``possible'' states that the total unlive state can be in
+for a given $\gamma$, without actually depending on any live state even in the refining locations.
+The $View_\kappa$ is in some sense used to justify the borrows at $\kappa$, so we will have an invariant,
+\begin{align}
+  & \forall \kappa, \gamma, m .~
+  (\kappa, \gamma, m) \in borrows \implies \forall y .~ View_\kappa(\gamma)(m)(y) \implies m \le y
+  \label{invView}
+\end{align}
+    %\left(
+    %    \Bigcircle_{\kappa' \ltGe \kappa} reserved(\kappa,\gamma)
+    %  \right) \cdot
+    %  \left(
+    %    \Bigcircle_{\delta} Project_\delta\left(Total_\kappa(\delta \refines_{\rel} \gamma)\right)
+    %  \right)
+   %\right)
+%\end{align}
+(We don't need a condition that $\kappa \ltGe active$, although we won't be able to make use
+of the invariant otherwise.)
+
+Now, we define $Inv$ to be both (1) and (2) together, and define an RA with a validity predicate $\mathcal{V}(x) \eqdef \exists y.~ Inv(x \cdot y)$. Then we can derive all the rules
+claimed above, by simply showing that they are frame-preserving within this RA.
+
+First, the borrowing rules: we can check that they preserve (\ref{invView}).
+Also, all of the borrowing rules will preserve $Total_{active}$,
+so they preserve (\ref{invTotal}) trivially.
+
+Now, for the frame update rules (Borrow-Ghost-ViewShift and Borrow-Ghost-Refinement-Exchange).
+First, these don't affect $View_{\kappa}$ at all, so it is easy to show that they preserve
+(\ref{invView}).
+
+We can also check that (\ref{invTotal}) is preserved. There are two key points here.
+First, the rules about TPCM refinements imply that a frame-preserving update on a
+$Total_\kappa(\delta \refines_{\rel} \gamma)$ implies a frame-preserving update
+on the projection,
+$Project_\delta(Total_\kappa(\delta \refines_{\rel} \gamma))$.
+
+Second, we have to show why it is allowed to use borrowed state to justify the transitions.
+Luckily, from (\ref{invTotal}) we can derive this law,
+\[
+  \kappa \ltGe active \implies View_\kappa(\gamma)(y)\left(
+      \left(
+        \Bigcircle_{\kappa' \ltGe active} reserved(\kappa,\gamma)
+      \right) \cdot
+      \left(
+        \Bigcircle_{\delta} Project_\delta\left(Total_{active}(\delta \refines_{\rel} \gamma)\right)
+      \right)
+    \right).
+    \label{invDerived}
+\]
+This large expression inside the $View_\kappa(\gamma)$ is just the part of
+$Total_{active}(\gamma)$ except the part \emph{which is live at $\gamma$}.
+Thus, combined with \ref{invView}, we can see that any borrow will be covered by the reserved 
+state at $\gamma$ plus all state in the refined locations.
+
+Just to reiterate the ``trick'': $View_\kappa(\gamma)$ depends only on reserved state:
+however we are able to show that the value \[\left(
+        \Bigcircle_{\kappa' \ltGe active} reserved(\kappa',\gamma)
+      \right) \cdot
+      \left(
+        \Bigcircle_{\delta} Project_\delta\left(Total_{active}(\delta \refines_{\rel} \gamma)\right)
+      \right)\] is in the active view, even though it includes the \emph{total} state at locations refining $\gamma$.
+
 \end{document}

### 2021-05-20 tools/local-dafny.sh
fix verification errors in MsgSeq

@@ -5,6 +5,7 @@
 
 
 echo "NOTE: this will use the /noNLarith flag! If you're working on a file where nonlinear arith is allowed use tools/local-dafny-nonlinear.sh"
+echo "NOTE: since it's okay to park in the middle of the street with my blinkers on, jonh added /compile:0"
 DIR="$( cd "$( dirname "${BASH_SOURCE[0]}" )" >/dev/null 2>&1 && pwd )"
 set -x
-$DIR/../.dafny/bin/dafny "$@" /induction:1 /noNLarith
+$DIR/../.dafny/bin/dafny "$@" /induction:1 /noNLarith /compile:0

### 2021-05-21 concurrency/docs/theory/theory.tex
typo
another typo
make something more clear
start drafting the formalized rwlock example

@@
 \[
-  \forall m .~ domain(\rel, m) \implies \rel(m, Project_\delta(m)).
+  \forall m .~ m \in domain(\rel) \implies \rel(m, Project_\delta(m)).
 \]
-where $domain(\rel, m) \eqdef \exists n .~ \rel(m, n)$.
+where $domain(\rel) \eqdef \{m : \exists n .~ \rel(m, n)\}$.
 
@@
 \begin{align}
-  domain(rel, Total_{active}(\delta \refines_{\rel} \gamma)) \label{invTotal}
+  Total_{active}(\delta \refines_{\rel} \gamma) \in domain(\rel) \label{invTotal}
 \end{align}
@@
   ProjectView_\delta(v) \eqdef \lambda a .~
-      \exists x, y .~ v(x) \land rel(x, y) \land y \le a.
+      \exists x, y .~ v(x) \land y = Project_\delta(x) \land (x \in domain(rel)) \land y \le a.
 \]
@@
   & \forall \kappa, \gamma, m .~
-  (\kappa, \gamma, m) \in borrows \implies \forall y .~ View_\kappa(\gamma)(m)(y) \implies m \le y
+  (\kappa, \gamma, m) \in borrows \implies \forall y .~ View_\kappa(\gamma)(y) \implies m \le y
   \label{invView}
@@
 \[
-  \kappa \ltGe active \implies View_\kappa(\gamma)(y)\left(
+  \kappa \ltGe active \implies View_\kappa(\gamma)\left(
       \left(
@@
 
+\textbf{Worked RWLock example}
+
+\newcommand{\Central}{\textsf{Central}}
+\newcommand{\ExcPending}{\textsf{ExcPending}}
+\newcommand{\SharedPending}{\textsf{SharedPending}}
+\newcommand{\ExcGuard}{\textsf{ExcGuard}}
+\newcommand{\SharedGuard}{\textsf{SharedGuard}}
+
+Given a TPCM $M$, we can construct a TPCM $RwLock_M$ with the following laws. (Here, we have $exc : \bool$, $rc \in \mathbb{N}$, and $x \in M$.
+
+These laws are for exclusive locks:
+\begin{align*}
+\dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma} &\viewshift \dboxed{\Central\left(\true, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcPending}^{\alpha \refines \gamma}\\
+%\Central\left(exc, 0, x\right) \cdot \ExcPending &\mapsto \Central\left(exc, 0, \bot\right) \cdot \ExcGuard
+\dboxed{\Central\left(exc, 0, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcPending}^{\alpha \refines \gamma}
+    &\viewshift
+    \dboxed{\Central\left(exc, 0, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma\\
+\dboxed{\Central\left(exc, rc, y\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma
+    &\viewshift
+        \dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma}\\
+\end{align*}
+The first two are for acquisition, while the third is for release. Meanwhile,
+the first one is derived from Ghost-ViewShift, while the last two are derived from
+Ghost-Refinement-Exchange.
+
+These laws are for shared locks:
+\begin{align*}
+\dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma} &\viewshift
+  \dboxed{\Central\left(exc, rc+1, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma}\\
+\dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma}
+    &\viewshift
+      \dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}\\
+\dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}
+    &\viewshift
+        \dboxed{\Central\left(exc, rc - 1, x\right)}^{\alpha \refines \gamma}
+\end{align*}
+All of these are derived from Ghost-ViewShift.
+
+We don't use Ghost-Refinement-Exchange, because in the `shared' case, our mechanism for getting back $M$-state is to use the Borrow-Back rule. In particular, we use it to derive this:
+\begin{align*}
+\borrow{\kappa}{\dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}}
+  &\viewshift
+    \borrow{\kappa}{\dboxed{x}^{\gamma}}
+\end{align*}
+This logic is all we need to verify a reader-writer lock.
+
 \end{document}

### 2021-05-24 concurrency/docs/theory/theory.tex
add frac & counting logic examples
rwlock hoare proofs

@@
 \usepackage{graphicx}
+\usepackage{xcolor}
 
@@
 \newcommand{\viewshift}{\Rrightarrow}
+\newcommand{\viewshiftBack}{\Lleftarrow}
 \newcommand{\unit}{\epsilon}
@@
 \]
-Finally, we're ready to define the $View_\kappa$ function for a given location, again
-recursively, like the $Total_\kappa$ function:
+Finally, we're ready to define the $View_\kappa$ function for a given location, again recursively, like the $Total_\kappa$ function:
 \[
@@
       \left(
-        \bigotimes_{\kappa' \ltGe \kappa} ViewOver(reserved(\kappa,\gamma))
+        \bigotimes_{\kappa' \ltGe \kappa} ViewOver(reserved(\kappa',\gamma))
       \right) \otimes
@@
 
-\textbf{Worked RWLock example}
+\textbf{Fractional logic}
+
+Given a TPCM $M$, we can construct a TPCM $Frac_M$ with the following laws.
+
+\newcommand{\FFrac}{\textsf{Frac}}
+
+\begin{align*}
+  \dboxed{x}^\gamma &\viewshiftBack \viewshift \dboxed{\FFrac(x, 1)}^{\alpha \refines \gamma}\\
+  \dboxed{\FFrac(x, r)}^{\alpha \refines \gamma}
+    *
+  \dboxed{\FFrac(x, q)}^{\alpha \refines \gamma}
+    &\viewshiftBack \viewshift
+  \dboxed{\FFrac(x, r+q)}^{\alpha \refines \gamma}
+\end{align*}
+And,
+\begin{align*}
+  \borrow{\kappa}{\dboxed{\FFrac(x, r)}^{\alpha \refines \gamma}} &\viewshift
+  \borrow{\kappa}{\dboxed{x}^\gamma}
+\end{align*}
+
+\textbf{Counting logic}
+
+\newcommand{\Counter}{\textsf{Counter}}
+\newcommand{\SharedRef}{\textsf{SharedRef}}
+
+\begin{align*}
+  \dboxed{x}^\gamma &\viewshiftBack \viewshift \dboxed{\Counter(x, 0)}^{\alpha \refines \gamma}\\
+  \dboxed{\Counter(x, n)}^{\alpha \refines \gamma}
+    &\viewshiftBack \viewshift
+  \dboxed{\Counter(x, n+1)}^{\alpha \refines \gamma}
+      * \dboxed{\SharedRef(x)}^{\alpha \refines \gamma}
+\end{align*}
+And,
+\begin{align*}
+  \borrow{\kappa}{\dboxed{\SharedRef(x)}^{\alpha \refines \gamma}} &\viewshift
+  \borrow{\kappa}{\dboxed{x}^\gamma}
+\end{align*}
+
+\textbf{RwLock Logic}
 
@@
 
+\textbf{RwLock Verified Example}
+
+\newcommand{\rwlock}{rwlock}
+\newcommand{\exc}{exc}
+\newcommand{\rc}{rc}
+\newcommand{\Pointer}{Pointer}
+
+\newcommand{\AcquireExc}{\text{AcquireExc}}
+\newcommand{\AcquireShared}{\text{AcquireShared}}
+\newcommand{\AcquireExcSetBit}{\text{AcquireExcSetBit}}
+\newcommand{\AcquireExcCheckRc}{\text{AcquireExcCheckRc}}
+\newcommand{\AcquireSharedIncRc}{\text{AcquireSharedIncRc}}
+\newcommand{\AcquireSharedCheckExc}{\text{AcquireSharedCheckExc}}
+\newcommand{\AcquireSharedAbort}{\text{AcquireSharedAbort}}
+\newcommand{\ReleaseExc}{\text{ReleaseExc}}
+\newcommand{\ReleaseShared}{\text{ReleaseShared}}
+
+\newcommand{\langloop}{\textbf{loop}}
+\newcommand{\until}{\textbf{until}}
+\newcommand{\langif}{\textbf{if}}
+\newcommand{\langthen}{\textbf{then}}
+\newcommand{\langelse}{\textbf{else}}
+
+\newcommand{\CAS}{\textbf{CAS}}
+\newcommand{\AtomicAdd}{\textbf{AtomicAdd}}
+
+\newcommand{\Inv}{Inv}
+\newcommand{\inv}{inv}
+\newcommand{\inlinehoare}[1]{{\color{blue}\left\{#1\right\}}}
+
+Let's use a struct $\rwlock := \{\exc: \Pointer, \rc: \Pointer\}$. Our implementation will be straightforward:
+
+\begin{align*}
+  \AcquireExc(\rwlock) ~\eqdef~
+      & \langloop~\until~\AcquireExcSetBit(\rwlock);\\
+      & \langloop~\until~\AcquireExcCheckRc(\rwlock)\\
+      \\
+  \AcquireShared(\rwlock) ~\eqdef~
+      & \langloop~\until~(\\
+        & ~~~~ (\langloop~\until~!\rwlock.\exc);\\
+        & ~~~~ \AcquireSharedIncRc(\rwlock);\\ %\AtomicAdd(\rwlock.\rc, 1);
+        & ~~~~ \langif~ \AcquireSharedCheckExc(\rwlock) ~\langthen\\
+        & ~~~~~~~~ \true\\
+        & ~~~~ \langelse\\
+        & ~~~~~~~~ \AcquireSharedAbort(\rwlock); \false\\ %\AtomicAdd(\rwlock.\rc, -1); \false
+      )\\
+      \\
+  \AcquireExcSetBit(\rwlock) ~\eqdef~
+      & \CAS(\rwlock.\exc,\false,\true)\\
+      \\
+  \AcquireExcCheckRc(\rwlock) ~\eqdef~
+      & (!\rwlock.\rc) = 0\\
+      \\
+  \AcquireSharedIncRc(\rwlock) ~\eqdef~
+      & \AtomicAdd(\rwlock.\rc, 1)\\
+      \\
+  \AcquireSharedAbort(\rwlock) ~\eqdef~
+      & \AtomicAdd(\rwlock.\rc, -1)\\
+      \\
+  \AcquireSharedCheckExc(\rwlock) ~\eqdef~
+      & (!\rwlock.\exc) = \false\\
+      \\
+  \ReleaseExc(\rwlock) ~\eqdef~
+      & \rwlock.\exc := \false\\
+      \\
+  \ReleaseShared(\rwlock) ~\eqdef~
+      & AtomicAdd(\rwlock.\rc, -1)
+\end{align*}
+We would like to prove the following specifications, for some predicate $\Inv(\rwlock,\alpha,\gamma,\inv)$. (Here, $\inv$ is the user-defined predicate on the ghost state stored by the $\rwlock$.)
+
+$\AcquireExc$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }\\
+& \AcquireExc(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \exists x.~ \dboxed{x}^\gamma \land \inv(x) }
+\end{align*}
+$\ReleaseExc$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma \land \inv(x) }\\
+& \ReleaseExc(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
+\end{align*}
+$\AcquireShared$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }\\
+& \AcquireShared(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \exists x.~ \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} \land \inv(x) }
+\end{align*}
+$\ReleaseShared$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} \land \inv(x) }\\
+& \ReleaseShared(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
+\end{align*}
+To prove these specifications, we will first define $\Inv$:
+\begin{align*}
+  \Inv(\rwlock) ~\eqdef~
+      \exists \iota, \boxed{
+        \exists exc,rc,x .~
+            \dboxed{\Central(exc,rc,x)}^{\alpha \refines \gamma}
+            * (\rwlock.\exc \pointsto exc)
+            * (\rwlock.\rc \pointsto rc)
+      }^{~\iota}
+\end{align*}
+Now, we can work out how to verify the above programs.
+These will follow from the below:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }\\
+& \AcquireExcSetBit(\rwlock)\\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) *
+    \dboxed{\ExcPending}^{\alpha\refines\gamma} }
+\end{align*}
+
 \end{document}

### 2021-05-24 concurrency/docs/theory/theory.tex
continue working out rwlock hoare proofs

--- 2021-05-21 concurrency/docs/theory/theory.tex
@@
 $Total_{active}(\gamma)$ except the part \emph{which is live at $\gamma$}.
-Thus, combined with \ref{invView}, we can see that any borrow will be covered by the reserved 
+Thus, combined with (\ref{invView}), we can see that any borrow will be covered by the reserved 
 state at $\gamma$ plus all state in the refined locations.
@@
 \begin{align*}
-\dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma} &\viewshift \dboxed{\Central\left(\true, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcPending}^{\alpha \refines \gamma}\\
+\dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma} &\viewshift \dboxed{\Central\left(\true, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcPending}^{\alpha \refines \gamma} && \text{(exc-begin)} \\
 %\Central\left(exc, 0, x\right) \cdot \ExcPending &\mapsto \Central\left(exc, 0, \bot\right) \cdot \ExcGuard
@@
     &\viewshift
-    \dboxed{\Central\left(exc, 0, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma\\
+    \dboxed{\Central\left(exc, 0, x\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma && \text{(exc-finish)} \\
 \dboxed{\Central\left(exc, rc, y\right)}^{\alpha \refines \gamma} * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma
     &\viewshift
-        \dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma}\\
+        \dboxed{\Central\left(\false, rc, x\right)}^{\alpha \refines \gamma} && \text{(exc-return)} \\
 \end{align*}
@@
 \dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma} &\viewshift
-  \dboxed{\Central\left(exc, rc+1, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma}\\
+  \dboxed{\Central\left(exc, rc+1, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma} && \text{(shared-begin)} \\
+\dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma}  * \dboxed{\SharedPending}^{\alpha \refines \gamma} &\viewshift
+  \dboxed{\Central\left(exc, rc-1, x\right)}^{\alpha \refines \gamma} && \text{(shared-abort)} \\
 \dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedPending}^{\alpha \refines \gamma}
     &\viewshift
-      \dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}\\
+      \dboxed{\Central\left(false, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} && \text{(shared-return)}\\
 \dboxed{\Central\left(exc, rc, x\right)}^{\alpha \refines \gamma} * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma}
     &\viewshift
-        \dboxed{\Central\left(exc, rc - 1, x\right)}^{\alpha \refines \gamma}
+        \dboxed{\Central\left(exc, rc - 1, x\right)}^{\alpha \refines \gamma} && \text{(shared-abort)}
 \end{align*}
@@
 
+\textbf{RwLock Verified Example}
+
+\newcommand{\rwlock}{rwlock}
+\newcommand{\exc}{exc}
+\newcommand{\rc}{rc}
+\newcommand{\Pointer}{Pointer}
+
+\newcommand{\AcquireExc}{\text{AcquireExc}}
+\newcommand{\AcquireShared}{\text{AcquireShared}}
+%\newcommand{\AcquireExcSetBit}{\text{AcquireExcSetBit}}
+%\newcommand{\AcquireExcCheckRc}{\text{AcquireExcCheckRc}}
+%\newcommand{\AcquireSharedIncRc}{\text{AcquireSharedIncRc}}
+%\newcommand{\AcquireSharedCheckExc}{\text{AcquireSharedCheckExc}}
+%\newcommand{\AcquireSharedAbort}{\text{AcquireSharedAbort}}
+\newcommand{\ReleaseExc}{\text{ReleaseExc}}
+\newcommand{\ReleaseShared}{\text{ReleaseShared}}
+
+\newcommand{\langloop}{\textbf{loop}}
+\newcommand{\until}{\textbf{until}}
+\newcommand{\langif}{\textbf{if}}
+\newcommand{\langthen}{\textbf{then}}
+\newcommand{\langelse}{\textbf{else}}
+
+\newcommand{\CAS}{\textbf{CAS}}
+\newcommand{\AtomicAdd}{\textbf{AtomicAdd}}
+
+\newcommand{\Inv}{Inv}
+\newcommand{\inv}{inv}
+\newcommand{\inlinehoare}[1]{{\color{blue}\left\{#1\right\}}}
+
+Let's use a struct $\rwlock := \{\exc: \Pointer, \rc: \Pointer\}$. Our implementation will be straightforward:
+
+\begin{align*}
+  \AcquireExc(\rwlock) ~\eqdef~
+      & \langloop~\until~\CAS(\rwlock.\exc,0,1);\\
+      & \langloop~\until~(!\rwlock.\rc) = 0\\
+      \\
+  \AcquireShared(\rwlock) ~\eqdef~
+      & \langloop~\until~(\\
+        & ~~~~ (\langloop~\until~!\rwlock.\exc);\\
+        & ~~~~ \AtomicAdd(\rwlock.\rc, 1);\\
+        & ~~~~ \langif~ (!\rwlock.\exc) = 0 ~\langthen\\
+        & ~~~~~~~~ \true\\
+        & ~~~~ \langelse\\
+        & ~~~~~~~~ \AtomicAdd(\rwlock.\rc, -1); \false\\
+      )\\
+      \\
+  %\AcquireExcSetBit(\rwlock) ~\eqdef~
+  %    & \CAS(\rwlock.\exc,\false,\true)\\
+  %    \\
+  %\AcquireExcCheckRc(\rwlock) ~\eqdef~
+  %    & (!\rwlock.\rc) = 0\\
+  %    \\
+  %\AcquireSharedIncRc(\rwlock) ~\eqdef~
+  %    & \AtomicAdd(\rwlock.\rc, 1)\\
+  %    \\
+  %\AcquireSharedAbort(\rwlock) ~\eqdef~
+  %    & \AtomicAdd(\rwlock.\rc, -1)\\
+  %    \\
+  %\AcquireSharedCheckExc(\rwlock) ~\eqdef~
+  %    & (!\rwlock.\exc) = \false\\
+  %    \\
+  \ReleaseExc(\rwlock) ~\eqdef~
+      & \rwlock.\exc := 0\\
+      \\
+  \ReleaseShared(\rwlock) ~\eqdef~
+      & \AtomicAdd(\rwlock.\rc, -1)
+\end{align*}
+We would like to prove the following specifications, for some predicate $\Inv(\rwlock,\alpha,\gamma,\inv)$. (Here, $\inv$ is the user-defined predicate on the ghost state stored by the $\rwlock$.)
+
+$\AcquireExc$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }\\
+& \AcquireExc(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \exists x.~ \dboxed{x}^\gamma \land \inv(x) }
+\end{align*}
+$\ReleaseExc$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma \land \inv(x) }\\
+& \ReleaseExc(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
+\end{align*}
+$\AcquireShared$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }\\
+& \AcquireShared(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \exists x.~ \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} \land \inv(x) }
+\end{align*}
+$\ReleaseShared$:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} \land \inv(x) }\\
+& \ReleaseShared(\rwlock) \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
+\end{align*}
+To prove these specifications, we will first define $\Inv$:
+\begin{align*}
+  \Inv(\rwlock) ~\eqdef~
+      \exists \iota, \boxed{
+        \exists exc,rc,x .~
+            \dboxed{\Central(exc,rc,x)}^{\alpha \refines \gamma}
+            * (\rwlock.\exc \pointsto exc)
+            * (\rwlock.\rc \pointsto rc)
+      }^{~\iota}
+\end{align*}
+Now, we can work out how to verify the above programs.
+$\AcquireExc$ will follow from the below:
+\begin{align}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) } \nonumber \\
+& \CAS(\rwlock.\exc,0,1) \nonumber \\ %\label{hoare-cas-exc-pending} \\
+& \inlinehoare{v.~
+  \Inv(\rwlock,\alpha,\gamma,\inv) *
+      \left(v \implies \dboxed{\ExcPending}^{\alpha\refines\gamma}\right) } \nonumber \\
+   \nonumber \\
+& \inlinehoare{
+    \Inv(\rwlock,\alpha,\gamma,\inv) *
+    \dboxed{\ExcPending}^{\alpha\refines\gamma} } \nonumber \\
+& (!\rwlock.\rc) \nonumber \\
+& \inlinehoare{v .~
+    \Inv(\rwlock,\alpha,\gamma,\inv) *
+    \left(v\ne 0 \implies \dboxed{\ExcPending}^{\alpha\refines\gamma}\right) *
+    \left(v=0 \implies \dboxed{\ExcGuard}^{\alpha\refines\gamma} *
+        \exists x.~ \dboxed{x}^{\gamma} \land \inv(x)\right)
+    } \nonumber
+\end{align}
+$\AcquireShared$ will follow from:
+\begin{align}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) } \nonumber \\
+& \AtomicAdd(\rwlock.\rc, 1) \label{hoare-atomic-add-shared-pending} \\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \SharedPending } \nonumber\\
+ \nonumber\\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \SharedPending }\nonumber\\
+& (!\rwlock.\exc)\nonumber\\
+& \inlinehoare{ v.~
+    \Inv(\rwlock,\alpha,\gamma,\inv) * \left(v \ne 0 \implies \SharedPending\right)
+        * \left(v = 0 \implies \exists x.~
+          \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} \land \inv(x)\right)
+   }\nonumber\\
+   \nonumber\\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \SharedPending }\nonumber\\
+& \AtomicAdd(\rwlock.\rc, -1)\nonumber\\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }\nonumber
+\end{align}
+$\ReleaseExc$ will follow from:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma \land \inv(x) }\\
+& \rwlock.\exc := 0\\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
+\end{align*}
+And finally, $\ReleaseShared$ will follow from:
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\SharedGuard(x)}^{\alpha \refines \gamma} \land \inv(x) }\\
+& \AtomicAdd(\rwlock.\rc, -1)\\
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
+\end{align*}
+Let's work out (\ref{hoare-atomic-add-shared-pending}) explicitly:
+%\begin{align*}
+%& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) } \\
+%& \inlinehoare{
+%        \exists exc,rc,x .~
+%            \dboxed{\Central(exc,rc,x)}^{\alpha \refines \gamma}
+%            * (\rwlock.\exc \pointsto exc)
+%            * (\rwlock.\rc \pointsto rc)
+%    }\\
+%& \CAS(\rwlock.\exc,0,1) \\
+%& \inlinehoare{v.~
+%  \Inv(\rwlock,\alpha,\gamma,\inv) *
+%      \left(v \implies \dboxed{\ExcPending}^{\alpha\refines\gamma}\right) }
+%\end{align*}
+\begin{align*}
+& \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) } \\
+\text{(open Inv)} ~~~~ & \inlinehoare{
+        \exists exc,rc,x .~
+            \dboxed{\Central(exc,rc,x)}^{\alpha \refines \gamma}
+            * (\rwlock.\exc \pointsto exc)
+            * (\rwlock.\rc \pointsto rc)
+    }\\
+& \AtomicAdd(\rwlock.\rc, 1) \\
+& \inlinehoare{
+        \exists exc,rc,x .~
+            \dboxed{\Central(exc,rc,x)}^{\alpha \refines \gamma}
+            * (\rwlock.\exc \pointsto exc)
+            * (\rwlock.\rc \pointsto rc + 1)
+    }\\
+\text{(shared-begin)} ~~~~ & \inlinehoare{
+        \exists exc,rc,x .~
+            \dboxed{\Central(exc,rc + 1,x)}^{\alpha \refines \gamma}
+            * \SharedPending
+            * (\rwlock.\exc \pointsto exc)
+            * (\rwlock.\rc \pointsto rc + 1)
+    }\\
+(rc'=rc+1) ~~~~ & \inlinehoare{
+        \exists exc,rc',x .~
+            \dboxed{\Central(exc,rc',x)}^{\alpha \refines \gamma}
+            * \SharedPending
+            * (\rwlock.\exc \pointsto exc)
+            * (\rwlock.\rc \pointsto rc')
+    }\\
+\text{(close Inv)} ~~~~ & \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \SharedPending } \nonumber\\
+\end{align*}
+
 \end{document}

### 2021-05-24 concurrency/docs/theory/theory.tex
minor edits
work on hash table writeup

@@
 \usepackage{xcolor}
+
+\newcommand{\BigAsterisk}{\mathop{\scalebox{3}{\raisebox{-0.2ex}{$\ast$}}}}%
 
@@
     }{
-      \forall p,q.~ \rel(b\cdot p, q) \implies \exists q' .~ \rel(b' \cdot p, q') \land (a \cdot q \mapsto a' \cdot q')
+      \forall p,q.~ \rel(b\cdot p) = Some(q) \implies \exists q' .~ \rel(b' \cdot p) = Some(q') \land (a \cdot q \mapsto a' \cdot q')
     }
@@
 
+\newpage
+
 \textbf{Fractional logic}
@@
 \end{align*}
+
+\newpage
 
@@
 
+\newpage
+
 \textbf{RwLock Verified Example}
@@
 
+\newcommand{\langassign}{\leftarrow}
 \newcommand{\langloop}{\textbf{loop}}
@@
 \newcommand{\langelse}{\textbf{else}}
+\newcommand{\langmatch}{\textbf{match}}
+\newcommand{\langwith}{\textbf{with}}
+\newcommand{\langend}{\textbf{end}}
+\newcommand{\langlet}{\textbf{let}~}
+\newcommand{\langref}{\textbf{ref}}
+\newcommand{\langin}{\textbf{in}}
 
@@
 \newcommand{\inv}{inv}
+\newcommand{\HTInv}{HTInv}
+\newcommand{\Heap}{Heap}
+\newcommand{\htinv}{htinv}
 \newcommand{\inlinehoare}[1]{{\color{blue}\left\{#1\right\}}}
 
-Let's use a struct $\rwlock := \{\exc: \Pointer, \rc: \Pointer\}$. Our implementation will be straightforward:
+Let's use a struct $\rwlock : \{\exc: \Pointer, \rc: \Pointer\}$. Our implementation will be straightforward:
 
@@
   \ReleaseExc(\rwlock) ~\eqdef~
-      & \rwlock.\exc := 0\\
+      & \rwlock.\exc \langassign 0\\
       \\
@@
 \begin{align*}
-  \Inv(\rwlock) ~\eqdef~
+  \Inv(\rwlock,\alpha,\gamma,\inv) ~\eqdef~
       \exists \iota, \boxed{
@@
 & \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) * \dboxed{\ExcGuard}^{\alpha \refines \gamma} * \dboxed{x}^\gamma \land \inv(x) }\\
-& \rwlock.\exc := 0\\
+& \rwlock.\exc \langassign 0\\
 & \inlinehoare{ \Inv(\rwlock,\alpha,\gamma,\inv) }
@@
 
+\newpage
+
+\textbf{A hash-table logic}
+
+\newcommand{\hashto}{\stackrel{\text{hash}}{\hookrightarrow}}
+\newcommand{\entryto}{\stackrel{\text{entry}}{\hookrightarrow}}
+\newcommand{\Some}{\text{Some}}
+\newcommand{\None}{\text{None}}
+
+\newcommand{\Query}{\text{Query}}
+\newcommand{\QueryIter}{\text{QueryIter}}
+\newcommand{\Update}{\text{Update}}
+\newcommand{\UpdateIter}{\text{UpdateIter}}
+\newcommand{\Remove}{\text{Remove}}
+\newcommand{\RemoveIter}{\text{RemoveIter}}
+\newcommand{\NotInRange}{\text{NotInRange}}
+
+\newcommand{\hash}{\ensuremath{hash}}
+\newcommand{\rec}{\textbf{rec}}
+
+%We will derive a TPCM (a ``normal'' one, not a refinement) to represent a hash table.
+We'll want to support the following specs for a linear-probing hash table.
+
+\begin{align*}
+\inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma} &~ \Query(ht,k) ~ \inlinehoare{v.~ \HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma \land v = v_0}\\
+\inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma} &~ \Update(ht,k, v) ~ \inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto \Some~v}^\gamma}\\
+\inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma} &~ \Remove(ht,k) ~ \inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto \None}^\gamma}
+\end{align*}
+
+We assume a function $\hash : Keys \to \mathbb{N}$. We'll let $ht : \{rwlocks: rwlock~list, table: ((Key,Value)~option)~list\}$.
+
+\begin{align*}
+  \Query(ht,k) ~\eqdef~ & \QueryIter(ht,k,\hash(k))\\
+  \QueryIter(ht,k,i) ~\eqdef~ & 
+      \AcquireShared(ht.rwlocks[i]);\\
+      &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+      &~~~~ |~ \None \Rightarrow \None\\
+      &~~~~ |~ \Some~(k_0,v_0) \Rightarrow
+          \langif~ k = k_0 ~\langthen~ Some(v_0) ~\langelse~ \QueryIter(ht,k,i+1)\\
+      &\langend) ~\langin\\
+      &\ReleaseShared(ht.rwlocks[i]);\\
+      &r\\
+      \\
+  \Update(ht,k,v) ~\eqdef~ & \UpdateIter(ht,k,v,\hash(k))\\
+  \UpdateIter(ht,k,v,i) ~\eqdef~ & 
+      \AcquireExc(ht.rwlocks[i]);\\
+      &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+      &~~~~ |~ \None \Rightarrow ht.table[i] \langassign (k,v)\\
+      &~~~~ |~ \Some~(k_0,v_0) \Rightarrow
+          \langif~ k = k_0 ~\langthen~ ht.table[i] \langassign (k,v) ~\langelse~ \UpdateIter(ht,k,v,i+1)\\
+      &\langend) ~\langin\\
+      &\ReleaseExc(ht.rwlocks[i]);\\
+      &r
+\end{align*}
+
+\newpage
+
+We'll define a TPCM resource for our hash table (that is, a ``normal'' TPCM, not a refinement). It will have two kinds of elements, $k \hashto v$ (where $v : Value~option$)
+and $i \entryto c$, where $c : (Key,Value)~option$.
+
+Allowable operations will be encoded into this logic. For example, we will have all the following:
+
+\emph{Update}.
+
+\begin{align*}
+  &\hash(k) \le u ~\land~ \left(\forall i.~ \hash(k) \le i < u \implies i \in entries \land \exists k_1,v_1 .~ entries[i] = \Some (k_1,v_1) \land k_1 \ne k\right)\\
+  &~~~~\land~ \left(entries[u] = \None ~\lor~ \exists v_1 .~ entries[u] = \Some (k, v) \right)\\
+  &~~~~\implies entries \cdot (k \hashto v_0) ~\mapsto~ entries[u \langassign \Some~(k,v)]  \cdot (k \hashto v)
+\end{align*}
+
+\emph{Query (found)}.
+
+The found case is easy: a single entry $(k,v)$ proves that the hash table maps $k$ to $v$.
+
+\[
+  \left((i \entryto (k,v_0)) \cdot (k \hashto v)\right) \ne \bot \implies v = \Some(v_0)
+\]
+
+The not-found case requires another quantifed predicate:
+
+\begin{align*}
+&\left(\forall i.~ \hash(k) \le i < u \implies i \in entries \land \exists k_1,v_1 .~ entries[i] = \Some (k_1,v_1) \land k_1 \ne k\right)\\
+  &~~~~\land~ \left(entries[u] = \None\right)\\
+  &~~~~\implies entries \cdot (k \hashto v) \ne \bot \implies v = \None
+\end{align*}
+
+We'll define $\HTInv$ as follows:
+\begin{align*}
+\HTInv(ht, \gamma) ~\eqdef~& \exists\alpha.~ \BigAsterisk_i \Inv\left(ht.rwlocks[i], \alpha, \gamma\times\Heap, \htinv(i, ht.entries[i])\right)\\
+\htinv(i, \ell) ~\eqdef~& \lambda x .~ \exists c .~ x = (i \entryto c, \ell \pointsto c)
+\end{align*}
+In other words, the RwLocks will be arranged so that we can use the lock at index $i$ to obtain
+some ghost state $(i \entryto c, ht.entries[i] \pointsto c)$.
+
+\newpage
+
+Now, we prove $\QueryIter$. (To avoid clutter, we silently skolemize existential
+variables as they are introduced.)
+
+\begin{align*}
+  &\inlinehoare{\active(\kappa) * \borrow{\kappa}{\dboxed{ht.entries}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land KeyNotInRange(ht.entries, k, i)}\\
+  &\AcquireShared(ht.rwlocks[i]);\\
+  &\inlinehoare{\active(\kappa) * \borrow{\kappa}{\dboxed{ht.entries}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land KeyNotInRange(ht.entries, k, i)
+      * \SharedGuard\left( (i \entryto c, ht.entries[i] \pointsto c) \right) }\\
+  &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+  &~~~~ |~ \None \Rightarrow \None\\
+  &~~~~ |~ \Some~(k_0,v_0) \Rightarrow
+      \langif~ k = k_0 ~\langthen~ Some(v_0) ~\langelse~ \QueryIter(ht,k,i+1)\\
+  &\langend) ~\langin\\
+  &\ReleaseShared(ht.rwlocks[i]);\\
+  &r\\
+  &\inlinehoare{\active(\kappa) * (k \hashto v_0) \land v = v_0}
+\end{align*}
+
 \end{document}

### 2021-05-24 concurrency/docs/theory/theory.tex
add a few more lines

@@
 \newcommand{\inlinehoare}[1]{{\color{blue}\left\{#1\right\}}}
+\newcommand{\inlinehoarel}[1]{{\color{blue}\left\{#1\right.}}
+\newcommand{\inlinehoarer}[1]{{\color{blue}\left.#1\right\}}}
 
@@
       * \SharedGuard\left( (i \entryto c, ht.entries[i] \pointsto c) \right) }\\
+  &\inlinehoarel{\active(\kappa) * \borrow{\kappa}{\dboxed{ht.entries}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land KeyNotInRange(ht.entries, k, i)}\\
+  &\inlinehoarer{~~~~
+      * \active(\kappa_0) * \borrowed\left(\kappa_0, \SharedGuard\left( (i \entryto c, ht.entries[i] \pointsto c) \right) \right)
+      * \borrow{\kappa_0}{\SharedGuard\left( (i \entryto c, ht.entries[i] \pointsto c) \right) }}\\
   &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\

### 2021-05-25 concurrency/docs/theory/theory.tex
finish the proof sketch of QueryIter

--- 2021-05-24 concurrency/docs/theory/theory.tex
@@
 \usepackage{xcolor}
+
+\newcommand{\BigAsterisk}{\mathop{\scalebox{3}{\raisebox{-0.2ex}{$\ast$}}}}%
+
+\DeclareMathOperator{\RangeFilledWithOtherKeys}{RangeFilledWithOtherKeys}
+\DeclareMathOperator{\SlotEmpty}{SlotEmpty}
+\DeclareMathOperator{\SlotHasKey}{SlotHasKey}
 
@@
 \newcommand{\inv}{inv}
+\newcommand{\HTInv}{HTInv}
+\newcommand{\Heap}{Heap}
+\newcommand{\htinv}{htinv}
 \newcommand{\inlinehoare}[1]{{\color{blue}\left\{#1\right\}}}
-
-Let's use a struct $\rwlock := \{\exc: \Pointer, \rc: \Pointer\}$. Our implementation will be straightforward:
+\newcommand{\inlinehoarel}[1]{{\color{blue}\left\{#1\right.}}
+\newcommand{\inlinehoarer}[1]{{\color{blue}\left.#1\right\}}}
+
+Let's use a struct $\rwlock : \{\exc: \Pointer, \rc: \Pointer\}$. Our implementation will be straightforward:
 
@@
 
+\newpage
+
+\textbf{A hash-table logic}
+
+\newcommand{\hashto}{\stackrel{\text{map}}{\hookrightarrow}}
+\newcommand{\entryto}{\stackrel{\text{linear}}{\hookrightarrow}}
+\newcommand{\Some}{\text{Some}}
+\newcommand{\None}{\text{None}}
+
+\newcommand{\Query}{\text{Query}}
+\newcommand{\QueryIter}{\text{QueryIter}}
+\newcommand{\Update}{\text{Update}}
+\newcommand{\UpdateIter}{\text{UpdateIter}}
+\newcommand{\Remove}{\text{Remove}}
+\newcommand{\RemoveIter}{\text{RemoveIter}}
+\newcommand{\NotInRange}{\text{NotInRange}}
+
+\newcommand{\hash}{\ensuremath{hash}}
+\newcommand{\rec}{\textbf{rec}}
+
+%We will derive a TPCM (a ``normal'' one, not a refinement) to represent a hash table.
+We'll want to support the following specs for a linear-probing hash table.
+
+\begin{align*}
+\inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma} &~ \Query(ht,k) ~ \inlinehoare{v.~ \HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma \land v = v_0}\\
+\inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma} &~ \Update(ht,k, v) ~ \inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto \Some~v}^\gamma}\\
+\inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto v_0}^\gamma} &~ \Remove(ht,k) ~ \inlinehoare{\HTInv(ht,\gamma) * \dboxed{k \hashto \None}^\gamma}
+\end{align*}
+
+We assume a function $\hash : Keys \to \mathbb{N}$. We'll let $ht : \{rwlocks: rwlock~list, table: ((Key,Value)~option)~list\}$.
+
+\begin{align*}
+  \Query(ht,k) ~\eqdef~ & \QueryIter(ht,k,\hash(k))\\
+  \QueryIter(ht,k,i) ~\eqdef~ & 
+      \AcquireShared(ht.rwlocks[i]);\\
+      &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+      &~~~~ |~ \None \Rightarrow \None\\
+      &~~~~ |~ \Some~(k_1,v_1) \Rightarrow
+          \langif~ k = k_1 ~\langthen~ Some(v_1) ~\langelse~ \QueryIter(ht,k,i+1)\\
+      &\langend) ~\langin\\
+      &\ReleaseShared(ht.rwlocks[i]);\\
+      &r\\
+      \\
+  \Update(ht,k,v) ~\eqdef~ & \UpdateIter(ht,k,v,\hash(k))\\
+  \UpdateIter(ht,k,v,i) ~\eqdef~ & 
+      \AcquireExc(ht.rwlocks[i]);\\
+      &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+      &~~~~ |~ \None \Rightarrow ht.table[i] \langassign (k,v)\\
+      &~~~~ |~ \Some~(k_1,v_1) \Rightarrow
+          \langif~ k = k_1 ~\langthen~ ht.table[i] \langassign (k,v) ~\langelse~ \UpdateIter(ht,k,v,i+1)\\
+      &\langend) ~\langin\\
+      &\ReleaseExc(ht.rwlocks[i]);\\
+      &r
+\end{align*}
+
+\newpage
+
+We'll define a TPCM resource for our hash table (that is, a ``normal'' TPCM, not a refinement). It will have two kinds of elements, $k \hashto v$ (where $v : Value~option$)
+and $i \entryto c$, where $c : (Key,Value)~option$.
+
+More precisely, we'll let an element of $HT$ be a pair of maps $(map, linear)$ where
+\begin{align*}
+k \hashto v ~\eqdef~ ([k := v], [])\\
+i \entryto c ~\eqdef~ ([], [i := c])
+\end{align*}
+$a \cdot b$ is given by pairwise map union, and is undefined $(\bot)$ should either map intersect. For example, we always have $(k \hashto v_1) \cdot (k \hashto v_2) = \bot$
+and $(i \entryto c_1) \cdot (i \entryto c_2) = \bot$. (Thus these are very similar to the usual points-to assertions $\pointsto$.
+
+Allowable operations will be encoded into this logic. For example, we will have all the following.
+\begin{align*}
+  \RangeFilledWithOtherKeys(e, k, a, b) &~\eqdef~ \forall i .~ a \le i < b \implies i \in e.linear \land \exists k_1,v_1 .~ entries[i] = \Some (k_1,v_1) \land k_1 \ne k\\
+  %\SlotEmpty(e, a) &~\eqdef~ e.entries[a] = \None\\
+  %\SlotHasKey(e, a, k) &~\eqdef~ \exists v.~ e.entries[a] = \Some(k,v)\\
+\end{align*}
+\emph{Update}.
+\begin{align*}
+  %&\hash(k) \le u ~\land~ \left(\forall i.~ \hash(k) \le i < u \implies i \in entries \land \exists k_1,v_1 .~ entries[i] = \Some (k_1,v_1) \land k_1 \ne k\right)\\
+  %&~~~~\land~ \left(entries[u] = \None ~\lor~ \exists v_1 .~ entries[u] = \Some (k, v) \right)\\
+  %&~~~~\implies entries \cdot (k \hashto v_0) ~\mapsto~ entries[u \langassign \Some~(k,v)]  \cdot (k \hashto v)
+  &\hash(k) \le u ~\land~ \RangeFilledWithOtherKeys(e, k, \hash(k), u)\\
+      %~\land~ \left(\SlotEmpty(e, u) \lor \SlotHasKey(e, a, k)\right)\\
+      &~~~~\implies e \cdot (u \entryto \None) \cdot (k \hashto v_0) ~\mapsto~
+                    e \cdot (u \entryto \Some~(k, v)) \cdot (k \hashto v)\\
+  \\
+  &\hash(k) \le u ~\land~ \RangeFilledWithOtherKeys(e, k, \hash(k), u)\\
+      %~\land~ \left(\SlotEmpty(e, u) \lor \SlotHasKey(e, a, k)\right)\\
+      &~~~~\implies e \cdot (u \entryto \Some~(k, v_1)) \cdot (k \hashto v_0) ~\mapsto~
+                    e \cdot (u \entryto \Some~(k, v)) \cdot (k \hashto v)
+\end{align*}
+
+\emph{Query}.
+
+The found case is easy: a single entry $(k,v)$ proves that the hash table maps $k$ to $v$.
+\[
+  \left((i \entryto (k,v_0)) \cdot (k \hashto v)\right) \ne \bot \implies v = \Some(v_0)
+\]
+The not-found case requires, like the Update case, a check that the key isn't in any of the cells:
+\begin{align*}
+%&\left(\forall i.~ \hash(k) \le i < u \implies i \in entries \land \exists k_1,v_1 .~ entries[i] = \Some (k_1,v_1) \land k_1 \ne k\right)\\
+%  &~~~~\land~ \left(entries[u] = \None\right)\\
+%  &~~~~\implies entries \cdot (k \hashto v) \ne \bot \implies v = \None
+&\hash(k) \le u \land \RangeFilledWithOtherKeys(e, k, \hash(k), u)
+  \implies e \cdot (u \entryto \None) \cdot (k \hashto v) \ne \bot \implies v = \None
+\end{align*}
+
+We'll define $\HTInv$ as follows:
+\begin{align*}
+\HTInv(ht, \gamma) ~\eqdef~& \exists\alpha.~ \BigAsterisk_i \Inv\left(ht.rwlocks[i], \alpha, \gamma\times\Heap, \htinv(i, ht.table[i])\right)\\
+\htinv(i, \ell) ~\eqdef~& \lambda x .~ \exists c .~ x = (i \entryto c, \ell \pointsto c)
+\end{align*}
+In other words, the RwLocks will be arranged so that we can use the lock at index $i$ to obtain, possibly in read-only ``shared'' mode,
+some ghost state $(i \entryto c, ht.table[i] \pointsto c)$.
+
+\newpage
+
+Now, we prove $\QueryIter$. (To avoid clutter, we silently skolemize existential
+variables as they are introduced.)
+The program begins with an established borrow at lifetime $\kappa$.
+We then acquire a shared lock, obtaining the $\SharedGuard$.
+We then borrow that, and from that, obtain borrowed state for the points-to assertion to read a cell from memory,
+along with the corresponding $\entryto$ state.
+The new borrow has lifetime $\kappa_0$. We intersect them to get $\kappa_1 = \kappa \ltIntersect \kappa_0$. We downgrade
+all the borrows to $\kappa_1$, which we will use for the recursive call. The sketch below shows step-by-step the process
+of introducing the new borrow, expiring it, and returning the lock (we handle the $\langmatch$ statement on the next page).
+We leave $\HTInv(ht,\gamma)$ implicit.
+\begin{align*}
+  &\inlinehoare{\active(\kappa) * \borrow{\kappa}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i)}\\
+  &\AcquireShared(ht.rwlocks[i]);\\
+  &\inlinehoare{\active(\kappa) * \borrow{\kappa}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i)
+      * \SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) }\\
+  &\inlinehoarel{\active(\kappa) * \borrow{\kappa}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i)}\\
+  &\inlinehoarer{~~~~
+      * \active(\kappa_0) * \borrowed\left(\kappa_0, \SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) \right)
+      * \borrow{\kappa_0}{\SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) }}\\
+  &\inlinehoarel{\active(\kappa) * \borrow{\kappa}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i) * \borrow{\kappa_0}{\dboxed{i\entryto c}^\gamma} * \borrow{\kappa_0}{ht.table[i] \pointsto c}}\\
+  &\inlinehoarer{~~~~
+      * \active(\kappa_0) * \borrowed\left(\kappa_0, \SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) \right) }\\
+  &\inlinehoarel{\active(\kappa_1) * \borrow{\kappa_1}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i) * \borrow{\kappa_1}{\dboxed{i\entryto c}^\gamma} * \borrow{\kappa_1}{ht.table[i] \pointsto c}}\\
+  &\inlinehoarer{~~~~
+      * \borrowed\left(\kappa_0, \SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) \right) }\\
+  &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+  &~~~~ |~ \None \Rightarrow \None\\
+  &~~~~ |~ \Some~(k_1,v_1) \Rightarrow
+      \langif~ k = k_1 ~\langthen~ \Some(v_1) ~\langelse~ \QueryIter(ht,k,i+1)\\
+  &\langend) ~\langin\\
+  &\inlinehoare{\active(\kappa_1)
+      * \dboxed{k \hashto v_0}^\gamma \land r = v_0
+      * \borrowed\left(\kappa_0, \SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) \right) }\\
+  &\inlinehoare{\active(\kappa)
+      * \dboxed{k \hashto v_0}^\gamma \land r = v_0
+      * \SharedGuard\left( (i \entryto c, ht.table[i] \pointsto c) \right) }\\
+  &\ReleaseShared(ht.rwlocks[i]);\\
+  &\inlinehoare{\active(\kappa) * \dboxed{k \hashto v_0}^\gamma \land r = v_0}\\
+  &r\\
+  &\inlinehoare{v.~ \active(\kappa) * \dboxed{k \hashto v_0}^\gamma \land v = v_0}
+\end{align*}
+
+\newpage
+
+The ``interesting'' logic is all in the match statement. Depending on the value of $c$, we either use our borrowed $\dboxed{i \entryto c}^\gamma$ to justify the result,
+or we combine it with $e$ for the recursive call (extending the range in the $\RangeFilledWithOtherKeys$ predicate to $i+1$).
+We let $e' = e \cdot (i \entryto \Some~(k_1,v_1))$.
+
+\begin{align*}
+  &\inlinehoare{\active(\kappa_1) * \borrow{\kappa_1}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i) * \borrow{\kappa_1}{\dboxed{i\entryto c}^\gamma} * \borrow{\kappa_1}{ht.table[i] \pointsto c}}\\
+  &\langlet r = (\langmatch~ !ht.table[i] ~\langwith\\
+  &~~~~ |~ \None \Rightarrow \\
+  &~~~~~~~~\inlinehoare{\active(\kappa_1) * \borrow{\kappa_1}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i) * \borrow{\kappa_1}{\dboxed{i\entryto \None}^\gamma}}\\
+  &~~~~~~~~\inlinehoare{\active(\kappa_1)
+      * \dboxed{k \hashto v_0}^\gamma \land v_0 = \None}\\
+  &~~~~~~~~\None\\
+  &~~~~ |~ \Some~(k_1,v_1) \Rightarrow \\
+  &~~~~~~~~\langif~ k = k_1 ~\langthen~ \\
+  &~~~~~~~~~~~~\inlinehoare{\active(\kappa_1) * \borrow{\kappa_1}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i) * \borrow{\kappa_1}{\dboxed{i\entryto \Some~(k,v_1)}^\gamma}}\\
+  &~~~~~~~~~~~~\inlinehoare{\active(\kappa_1)
+      * \dboxed{k \hashto v_0}^\gamma \land v_0 = \Some(v_1)}\\
+  &~~~~~~~~~~~~\Some(v_1)\\
+  &~~~~~~~~~\langelse~\\
+  &~~~~~~~~~~~~\inlinehoare{\active(\kappa_1) * \borrow{\kappa_1}{\dboxed{e}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e, k, \hash(k), i) * \borrow{\kappa_1}{\dboxed{i\entryto \Some~(k_1,v_1)}^\gamma} \land k\ne k_1}\\
+  &~~~~~~~~~~~~\inlinehoare{\active(\kappa_1) * \borrow{\kappa_1}{\dboxed{e'}^\gamma}
+      * \dboxed{k \hashto v_0}^\gamma \land \RangeFilledWithOtherKeys(e', k, \hash(k), i+1)}\\
+  &~~~~~~~~~~~~~~~~~~~~~~~~~~~~\text{\color{blue} (where $e' = e \cdot (i \entryto \Some~(k_1,v_1))$)}\\
+  &~~~~~~~~~~~~\QueryIter(ht,k,i+1)\\
+  &~~~~~~~~~~~~\inlinehoare{v.~ \active(\kappa_1) * \dboxed{k \hashto v_0}^\gamma \land v = v_0}\\
+  &\langend) ~\langin\\
+  &\inlinehoare{\active(\kappa_1)
+      * \dboxed{k \hashto v_0}^\gamma \land r = v_0}
+\end{align*}
 \end{document}
